
## IBTrACS Full Analysis

```{r global_ibtracs, echo=FALSE}
```

### Summary Statistics (overall)

```{r, echo=FALSE}
ibtracs_global_results$summary_stat |>
  knitr::kable()
```

```{r, echo=FALSE}
ibtracs_global_results$plot_data
```


### Regression Methods

#### Poisson Regression on non-Shorties

```{r, echo=FALSE}
(ibtracs_global_results$pois_fit)
drop1(ibtracs_global_results$pois_fit, test="LRT")
```


Using this model, we would predict `r predict(ibtracs_global_results$pois_fit, type="response")[1]` storms in `r min(ibtracs_global_results$data$SEASON)` and `r predict(ibtracs_global_results$pois_fit, type="response")[num_years_ibtracs]` storms in `r max(ibtracs_global_results$data$SEASON)`.

That is a change of `r (predict(ibtracs_global_results$pois_fit, type="response")[num_years_ibtracs]-predict(ibtracs_global_results$pois_fit, type="response")[1])/num_years_ibtracs*10` storms per decade.

#### Poisson Regression on Hurricanes

```{r, echo=FALSE}
(ibtracs_global_results$hur_fit)
drop1(ibtracs_global_results$hur_fit, test="LRT")
```

Using this model the number of a hurricanes in `r min(ibtracs_global_results$data$SEASON)` was `r predict(ibtracs_global_results$hur_fit, type="response")[1]` and `r predict(ibtracs_global_results$hur_fit, type="response")[num_years_ibtracs]` in `r max(ibtracs_global_results$data$SEASON)`.

That is a change of `r (predict(ibtracs_global_results$hur_fit, type="response")[num_years_ibtracs]-predict(ibtracs_global_results$hur_fit, type="response")[1])/num_years_ibtracs*10` storms per decade.

#### Poisson Regression on Major Storms

```{r, echo=FALSE}
(ibtracs_global_results$major_fit)
drop1(ibtracs_global_results$major_fit, test="LRT")
```

Using this model the number of a major storm in `r min(ibtracs_global_results$data$SEASON)` was `r predict(ibtracs_global_results$major_fit, type="response")[1]` and `r predict(ibtracs_global_results$major_fit, type="response")[num_years_ibtracs]` in `r max(ibtracs_global_results$data$SEASON)`.

That is a change of `r (predict(ibtracs_global_results$major_fit, type="response")[num_years_ibtracs]-predict(ibtracs_global_results$major_fit, type="response")[1])/num_years_ibtracs*10` storms per decade.

#### Poisson Regression on Intense Storms

```{r, echo=FALSE}
(ibtracs_global_results$intense_fit)
drop1(ibtracs_global_results$intense_fit, test="LRT")
```


Using this model the number of a intense storms in `r min(ibtracs_global_results$data$SEASON)` was `r predict(ibtracs_global_results$intense_fit, type="response")[1]` and `r predict(ibtracs_global_results$intense_fit, type="response")[num_years_ibtracs]` in `r max(ibtracs_global_results$data$SEASON)`.

That is a change of `r (predict(ibtracs_global_results$intense_fit, type="response")[num_years_ibtracs]-predict(ibtracs_global_results$intense_fit, type="response")[1])/num_years_ibtracs*10` storms per decade.

#### Logistic Regression on Proportions of Major

```{r, echo=FALSE}
(ibtracs_global_results$major_logit_fit)
drop1(ibtracs_global_results$major_logit_fit, test="LRT")
```



Using this model the probability of a major storm in `r min(ibtracs_global_results$data$SEASON)` was `r predict(ibtracs_global_results$major_logit_fit, type="response")[1]` and `r predict(ibtracs_global_results$major_logit_fit, type="response")[num_years_ibtracs]` in `r max(ibtracs_global_results$data$SEASON)`.

That is a change of `r (predict(ibtracs_global_results$major_logit_fit, type="response")[num_years_ibtracs]-predict(ibtracs_global_results$major_logit_fit, type="response")[1])/num_years_ibtracs*10*100` percent per decade.

#### Logistic Regression on Proportions of Intense

```{r, echo=FALSE}
(ibtracs_global_results$intense_logit_fit)
drop1(ibtracs_global_results$intense_logit_fit, test="LRT")
```


Using this model the probability of an intense storm in `r min(ibtracs_global_results$data$SEASON)` was `r predict(ibtracs_global_results$intense_logit_fit, type="response")[1]` and `r predict(ibtracs_global_results$intense_logit_fit, type="response")[num_years_ibtracs]` in `r max(ibtracs_global_results$data$SEASON)`.

That is a change of `r (predict(ibtracs_global_results$intense_logit_fit, type="response")[num_years_ibtracs]-predict(ibtracs_global_results$intense_logit_fit, type="response")[1])/num_years_ibtracs*10*100` percent per decade.

### Changepoint Methods

#### Change in Non-Shorty Counts

```{r, echo=FALSE}
ibtracs_global_results$pois_chpt_test
```

The dates suggest the changepoint would occur from `r 1980+ibtracs_global_results$pois_chpt_test[2]` to `r 1980+ibtracs_global_results$pois_chpt_test[2]+1` with $p$-value `r ibtracs_global_results$pois_chpt_test[3]`.

In regime one, the mean number of total storms is `r mean(ibtracs_global_results$data$Total_Storms[1:ibtracs_global_results$pois_chpt_test[2]])` and in regime two, the mean number of storms is `r mean(ibtracs_global_results$data$Total_Storms[(ibtracs_global_results$pois_chpt_test[2]+1):num_years_ibtracs])`; a change of `r mean(ibtracs_global_results$data$Total_Storms[(ibtracs_global_results$pois_chpt_test[2]+1):num_years_ibtracs]) - mean(ibtracs_global_results$data$Total_Storms[1:ibtracs_global_results$pois_chpt_test[2]])`.


#### Change in counts of hurricanes

```{r, echo=FALSE}
ibtracs_global_results$hur_pois_chpt_test
```

The dates suggest the changepoint would occur from `r 1980+ibtracs_global_results$hur_pois_chpt_test[2]` to `r 1980+ibtracs_global_results$hur_pois_chpt_test[2]+1` with $p$-value `r ibtracs_global_results$hur_pois_chpt_test[3]`.

In regime one, the mean number of hurricanes is `r mean(ibtracs_global_results$data$Hurricanes[1:ibtracs_global_results$hur_pois_chpt_test[2]])` and in regime two, the mean number of storms is `r mean(ibtracs_global_results$data$Hurricanes[(ibtracs_global_results$hur_pois_chpt_test[2]+1):num_years_ibtracs])`; a change of `r mean(ibtracs_global_results$data$Hurricanes[(ibtracs_global_results$hur_pois_chpt_test[2]+1):num_years_ibtracs]) - mean(ibtracs_global_results$data$Hurricanes[1:ibtracs_global_results$hur_pois_chpt_test[2]])`.


#### Change in counts of major storms

```{r, echo=FALSE}
ibtracs_global_results$major_pois_chpt_test
```

The dates suggest the changepoint would occur from `r 1980+ibtracs_global_results$major_pois_chpt_test[2]` to `r 1980+ibtracs_global_results$major_pois_chpt_test[2]+1` with $p$-value `r ibtracs_global_results$major_pois_chpt_test[3]`.

In regime one, the mean number of major storms is `r mean(ibtracs_global_results$data$Major_Storms[1:ibtracs_global_results$major_pois_chpt_test[2]])` and in regime two, the mean number of storms is `r mean(ibtracs_global_results$data$Major_Storms[(ibtracs_global_results$major_pois_chpt_test[2]+1):num_years_ibtracs])`; a change of `r mean(ibtracs_global_results$data$Major_Storms[(ibtracs_global_results$major_pois_chpt_test[2]+1):num_years_ibtracs]) - mean(ibtracs_global_results$data$Major_Storms[1:ibtracs_global_results$major_pois_chpt_test[2]])`.

#### Change in counts of intense storms

```{r, echo=FALSE}
ibtracs_global_results$intense_pois_chpt_test
```


The dates suggest the changepoint would occur from `r 1980+ibtracs_global_results$intense_pois_chpt_test[2]` to `r 1980+ibtracs_global_results$intense_pois_chpt_test[2]+1` with $p$-value `r ibtracs_global_results$intense_pois_chpt_test[3]`.

In regime one, the mean number of intense storms is `r mean(ibtracs_global_results$data$Intense_Storms[1:ibtracs_global_results$intense_pois_chpt_test[2]])` and in regime two, the mean number of storms is `r mean(ibtracs_global_results$data$Intense_Storms[(ibtracs_global_results$intense_pois_chpt_test[2]+1):num_years_ibtracs])`; a change of `r mean(ibtracs_global_results$data$Intense_Storms[(ibtracs_global_results$intense_pois_chpt_test[2]+1):num_years_ibtracs]) - mean(ibtracs_global_results$data$Intense_Storms[1:ibtracs_global_results$intense_pois_chpt_test[2]])`.

#### Proportion of major storms

A change in the proportion of major storms?
  
```{r, echo=FALSE}
ibtracs_global_results$prop_major_chpt_test
```

The change in proportions occurs `r 1980+ibtracs_global_results$prop_major_chpt_test[2]` to `r 1980+ibtracs_global_results$prop_major_chpt_test[2]+1` with $p$-value `r ibtracs_global_results$prop_major_chpt_test[3]`. 

In regime one, the proportion of major storms is `r sum(ibtracs_global_results$data$Major_Storms[1:ibtracs_global_results$prop_major_chpt_test[2]])/sum(ibtracs_global_results$data$Total_Storms[1:ibtracs_global_results$prop_major_chpt_test[2]])` and in regime two, the proportion of major storms is `r sum(ibtracs_global_results$data$Major_Storms[(ibtracs_global_results$prop_major_chpt_test[2]+1):num_years_ibtracs])/sum(ibtracs_global_results$data$Total_Storms[(ibtracs_global_results$prop_major_chpt_test[2]+1):num_years_ibtracs])`; a change of `r sum(ibtracs_global_results$data$Major_Storms[(ibtracs_global_results$prop_major_chpt_test[2]+1):num_years_ibtracs])/sum(ibtracs_global_results$data$Total_Storms[(ibtracs_global_results$prop_major_chpt_test[2]+1):num_years_ibtracs]) - sum(ibtracs_global_results$data$Major_Storms[1:ibtracs_global_results$prop_major_chpt_test[2]])/sum(ibtracs_global_results$data$Total_Storms[1:ibtracs_global_results$prop_major_chpt_test[2]])`.

#### Proportion of intense storms

A change in the proportion of major storms?
  
```{r, echo=FALSE}
ibtracs_global_results$prop_intense_chpt_test
```

The change in proportions occurs `r 1980+ibtracs_global_results$prop_intense_chpt_test[2]` to `r 1980+ibtracs_global_results$prop_intense_chpt_test[2]+1` with $p$-value `r ibtracs_global_results$prop_intense_chpt_test[3]`.


In regime one, the proportion of intense storms is `r sum(ibtracs_global_results$data$Intense_Storms[1:ibtracs_global_results$prop_intense_chpt_test[2]])/sum(ibtracs_global_results$data$Total_Storms[1:ibtracs_global_results$prop_intense_chpt_test[2]])` and in regime two, the proportion of major storms is `r sum(ibtracs_global_results$data$Intense_Storms[(ibtracs_global_results$prop_intense_chpt_test[2]+1):num_years_ibtracs])/sum(ibtracs_global_results$data$Total_Storms[(ibtracs_global_results$prop_intense_chpt_test[2]+1):num_years_ibtracs])`; a change of `r sum(ibtracs_global_results$data$Intense_Storms[(ibtracs_global_results$prop_intense_chpt_test[2]+1):num_years_ibtracs])/sum(ibtracs_global_results$data$Total_Storms[(ibtracs_global_results$prop_intense_chpt_test[2]+1):num_years_ibtracs]) - sum(ibtracs_global_results$data$Intense_Storms[1:ibtracs_global_results$prop_intense_chpt_test[2]])/sum(ibtracs_global_results$data$Total_Storms[1:ibtracs_global_results$prop_intense_chpt_test[2]])`.



## IBTrACS Test Results



**Generalized Linear Model Approaches**

Note:

* we report the Likelihood Ratio Test for the generalized linear models, this is better to use than the default $Z$-test provided in the `summary()` output.
* the Change per decade is found by comparing the predicted value at the start time (`r min(ibtracs_global_results$data$SEASON)`) and end time (`r max(ibtracs_global_results$data$SEASON)`) and scaling by the number of years $\times 10$.

```{r, echo=FALSE}
df <- data.frame(
  Response = c("Cyclone Freq", "Hurricanes", "Major Freq", "Prop Major", "Intense Freq", "Prop Freq"),
  Coefficent = c(coef(ibtracs_global_results$pois_fit)[2], 
                 coef(ibtracs_global_results$hur_fit)[2],
                 coef(ibtracs_global_results$major_fit)[2], 
                 coef(ibtracs_global_results$major_logit_fit)[2], 
                 coef(ibtracs_global_results$intense_fit)[2], 
                 coef(ibtracs_global_results$intense_logit_fit)[2] ),
  `p-value` = c(anova(ibtracs_global_results$pois_fit)$`Pr(>Chi)`[2], 
                anova(ibtracs_global_results$hur_fit)$`Pr(>Chi)`[2], 
                anova(ibtracs_global_results$major_fit)$`Pr(>Chi)`[2], 
                anova(ibtracs_global_results$major_logit_fit)$`Pr(>Chi)`[2], 
                anova(ibtracs_global_results$intense_fit)$`Pr(>Chi)`[2], 
                anova(ibtracs_global_results$intense_logit_fit)$`Pr(>Chi)`[2]),
  `Start` = c(predict(ibtracs_global_results$pois_fit, type="response")[1], 
              predict(ibtracs_global_results$hur_fit, type="response")[1],
              predict(ibtracs_global_results$major_fit, type="response")[1], 
              predict(ibtracs_global_results$major_logit_fit, type="response")[1]*100, 
              predict(ibtracs_global_results$intense_fit, type="response")[1],
              predict(ibtracs_global_results$intense_logit_fit, type="response")[1]*100 ),
  `End` = c(predict(ibtracs_global_results$pois_fit, type="response")[num_years_ibtracs],
            predict(ibtracs_global_results$hur_fit, type="response")[num_years_ibtracs],
            predict(ibtracs_global_results$major_fit, type="response")[num_years_ibtracs],
            predict(ibtracs_global_results$major_logit_fit, type="response")[num_years_ibtracs]*100,
            predict(ibtracs_global_results$intense_fit, type="response")[num_years_ibtracs],
            predict(ibtracs_global_results$intense_logit_fit, type="response")[num_years_ibtracs]*100)
)

df |>
  mutate(`Change per Decade` = (End- Start)/num_years_ibtracs*10) |>
  knitr::kable(digits=5)
```

**Changepoint Methods**

```{r, echo=FALSE}
df <- data.frame(
  Response = c("Cyclone Freq", "Hurricanes", "Major Freq", "Prop Major", "Intense Freq", "Prop Freq"),
  `Chpt Location` = c(ibtracs_global_results$pois_chpt_test[2]+1981,
                      ibtracs_global_results$hur_pois_chpt_test[2]+1981,
                      ibtracs_global_results$major_pois_chpt_test[2]+1981,
                      ibtracs_global_results$prop_major_chpt_test[2]+1981,
                      ibtracs_global_results$intense_pois_chpt_test[2]+1981,
                      ibtracs_global_results$prop_intense_chpt_test[2]+1981),
  `p-value` =c(ibtracs_global_results$pois_chpt_test[3],
               ibtracs_global_results$hur_pois_chpt_test[3],
               ibtracs_global_results$major_pois_chpt_test[3],
               ibtracs_global_results$prop_major_chpt_test[3],
               ibtracs_global_results$intense_pois_chpt_test[3],
               ibtracs_global_results$prop_intense_chpt_test[3] ),
  `Regime1` = c(mean(ibtracs_global_results$data$Total_Storms[1:ibtracs_global_results$pois_chpt_test[2]]),
                mean(ibtracs_global_results$data$Hurricanes[1:ibtracs_global_results$hur_pois_chpt_test[2]]),
                mean(ibtracs_global_results$data$Major_Storms[1:ibtracs_global_results$major_pois_chpt_test[2]]),
                sum(ibtracs_global_results$data$Major_Storms[1:ibtracs_global_results$prop_major_chpt_test[2]])/sum(ibtracs_global_results$data$Total_Storms[1:ibtracs_global_results$prop_major_chpt_test[2]])*100,
                mean(ibtracs_global_results$data$Intense_Storms[1:ibtracs_global_results$intense_pois_chpt_test[2]]),
                sum(ibtracs_global_results$data$Intense_Storms[1:ibtracs_global_results$prop_intense_chpt_test[2]])/sum(ibtracs_global_results$data$Total_Storms[1:ibtracs_global_results$prop_intense_chpt_test[2]])*100 ),
  `Regime2` = c(mean(ibtracs_global_results$data$Total_Storms[(ibtracs_global_results$pois_chpt_test[2]+1):num_years_ibtracs]),
                mean(ibtracs_global_results$data$Hurricanes[(ibtracs_global_results$hur_pois_chpt_test[2]+1):num_years_ibtracs]),
                mean(ibtracs_global_results$data$Major_Storms[(ibtracs_global_results$major_pois_chpt_test[2]+1):num_years_ibtracs]),
                sum(ibtracs_global_results$data$Major_Storms[(ibtracs_global_results$prop_major_chpt_test[2]+1):num_years_ibtracs])/sum(ibtracs_global_results$data$Total_Storms[(ibtracs_global_results$prop_major_chpt_test[2]+1):num_years_ibtracs])*100,
                mean(ibtracs_global_results$data$Intense_Storms[(ibtracs_global_results$intense_pois_chpt_test[2]+1):num_years_ibtracs]),
                sum(ibtracs_global_results$data$Intense_Storms[(ibtracs_global_results$prop_intense_chpt_test[2]+1):num_years_ibtracs])/sum(ibtracs_global_results$data$Total_Storms[(ibtracs_global_results$prop_intense_chpt_test[2]+1):num_years_ibtracs])*100)
  )

df |>
  mutate(Change = Regime2-Regime1,
         Perc_Change = Change/Regime1*100) |>
  knitr::kable(digits=5)
  
```

