
# Northern Indian

An analysis of the Northern Indian basin - the least active of all the basins.

## Full IBTracs Record

First, the full IBTracs data.

```{r, echo=FALSE}
ibtracs_ni_results <- get_report_basin(basin="Northern Indian", all_basins=ibtracs_all_basins)
```

```{r, echo=FALSE}
ibtracs_ni_results$plot_data
```

### Summary Statistics (overall)

```{r, echo=FALSE}
ibtracs_ni_results$summary_stat |>
  knitr::kable()
```

### Regression Methods

#### Poisson Regression on non-Shorties

```{r, echo=FALSE}
summary(ibtracs_ni_results$pois_fit)
```



Note, this says a decrease of `r abs(coef(ibtracs_ni_results$pois_fit)[2])` log-storms per year. This corresponds to a `r (1-exp(coef(ibtracs_ni_results$pois_fit)[2]))*100` percent decrease per season. So over 45 years, that is a decrease of `r (1-exp(coef(ibtracs_ni_results$pois_fit)[2])^45)*100` percent.

Using this model, we would predict `r predict(ibtracs_ni_results$pois_fit, type="response")[1]` storms in 1980 and `r predict(ibtracs_ni_results$pois_fit, type="response")[45]` storms in 2024.

That is a change of `r (predict(ibtracs_ni_results$pois_fit, type="response")[45]-predict(ibtracs_ni_results$pois_fit, type="response")[1])/45*10` storms per decade.

#### Poisson Regression on Major Storms

```{r, echo=FALSE}
summary(ibtracs_ni_results$major_fit)
```

Using this model the number of a major storm in 1980 was `r predict(ibtracs_ni_results$major_fit, type="response")[1]` and `r predict(ibtracs_ni_results$major_fit, type="response")[45]` in 2024.

That is a change of `r (predict(ibtracs_ni_results$major_fit, type="response")[45]-predict(ibtracs_ni_results$major_fit, type="response")[1])/45*10` storms per decade.

#### Poisson Regression on Intense Storms

```{r, echo=FALSE}
summary(ibtracs_ni_results$intense_fit)
```


Using this model the number of a major storm in 1980 was `r predict(ibtracs_ni_results$intense_fit, type="response")[1]` and `r predict(ibtracs_ni_results$intense_fit, type="response")[45]` in 2024.

That is a change of `r (predict(ibtracs_ni_results$intense_fit, type="response")[45]-predict(ibtracs_ni_results$intense_fit, type="response")[1])/45*10` storms per decade.

#### Logistic Regression on Proportions of Major

```{r, echo=FALSE}
summary(ibtracs_ni_results$major_logit_fit)
```

We see a significant result with a positive coefficient, so the log-odds of major storms has increased. This tells us the proportion of major storms has increased.

Specifically, using this model the probability of a major storm in 1980 was `r predict(ibtracs_ni_results$major_logit_fit, type="response")[1]` and `r predict(ibtracs_ni_results$major_logit_fit, type="response")[45]` in 2024.

That is a change of `r (predict(ibtracs_ni_results$major_logit_fit, type="response")[45]-predict(ibtracs_ni_results$major_logit_fit, type="response")[1])/45*10*100` percent per decade.

#### Logistic Regression on Proportions of Intense

```{r, echo=FALSE}
summary(ibtracs_ni_results$intense_logit_fit)
```


Specifically, using this model the probability of a major storm in 1980 was `r predict(ibtracs_ni_results$intense_logit_fit, type="response")[1]` and `r predict(ibtracs_ni_results$intense_logit_fit, type="response")[45]` in 2024.

That is a change of `r (predict(ibtracs_ni_results$intense_logit_fit, type="response")[45]-predict(ibtracs_ni_results$intense_logit_fit, type="response")[1])/45*10*100` percent per decade.

### Change Point Methods

#### Change in Non-Shorty Counts

```{r, echo=FALSE}
ibtracs_ni_results$pois_chpt_test
```

The dates suggest the changepoint would occur from `r 1979+ibtracs_ni_results$pois_chpt_test[2]` to `r 1979+ibtracs_ni_results$pois_chpt_test[2]+1` with $p$-value `r ibtracs_ni_results$pois_chpt_test[3]`.

In regime one, the mean number of total storms is `r mean(ibtracs_ni_results$data$Total_Storms[1:ibtracs_ni_results$pois_chpt_test[2]])` and in regime two, the mean number of storms is `r mean(ibtracs_ni_results$data$Total_Storms[(ibtracs_ni_results$pois_chpt_test[2]+1):45])`; a change of `r mean(ibtracs_ni_results$data$Total_Storms[(ibtracs_ni_results$pois_chpt_test[2]+1):45]) - mean(ibtracs_ni_results$data$Total_Storms[1:ibtracs_ni_results$pois_chpt_test[2]])`.

#### Change in counts of major storms

```{r, echo=FALSE}
ibtracs_ni_results$major_pois_chpt_test
```

The dates suggest the changepoint would occur from `r 1979+ibtracs_ni_results$major_pois_chpt_test[2]` to `r 1979+ibtracs_ni_results$major_pois_chpt_test[2]+1` with $p$-value `r ibtracs_ni_results$major_pois_chpt_test[3]`.

In regime one, the mean number of major storms is `r mean(ibtracs_ni_results$data$Major_Storms[1:ibtracs_ni_results$major_pois_chpt_test[2]])` and in regime two, the mean number of storms is `r mean(ibtracs_ni_results$data$Major_Storms[(ibtracs_ni_results$major_pois_chpt_test[2]+1):45])`; a change of `r mean(ibtracs_ni_results$data$Major_Storms[(ibtracs_ni_results$major_pois_chpt_test[2]+1):45]) - mean(ibtracs_ni_results$data$Major_Storms[1:ibtracs_ni_results$major_pois_chpt_test[2]])`.

#### Change in counts of intense storms

```{r, echo=FALSE}
ibtracs_ni_results$intense_pois_chpt_test
```


The dates suggest the changepoint would occur from `r 1979+ibtracs_ni_results$intense_pois_chpt_test[2]` to `r 1979+ibtracs_ni_results$intense_pois_chpt_test[2]+1` with $p$-value `r ibtracs_ni_results$intense_pois_chpt_test[3]`.

In regime one, the mean number of intense storms is `r mean(ibtracs_ni_results$data$Intense_Storms[1:ibtracs_ni_results$intense_pois_chpt_test[2]])` and in regime two, the mean number of storms is `r mean(ibtracs_ni_results$data$Intense_Storms[(ibtracs_ni_results$intense_pois_chpt_test[2]+1):45])`; a change of `r mean(ibtracs_ni_results$data$Intense_Storms[(ibtracs_ni_results$intense_pois_chpt_test[2]+1):45]) - mean(ibtracs_ni_results$data$Intense_Storms[1:ibtracs_ni_results$intense_pois_chpt_test[2]])`.

#### Proportion of major storms

A change in the proportion of major storms?
  
```{r, echo=FALSE}
ibtracs_ni_results$prop_major_chpt_test
```

The change in proportions occurs `r 1979+ibtracs_ni_results$prop_major_chpt_test[2]` to `r 1979+ibtracs_ni_results$prop_major_chpt_test[2]+1` with $p$-value `r ibtracs_ni_results$prop_major_chpt_test[3]`. 

In regime one, the proportion of major storms is `r sum(ibtracs_ni_results$data$Major_Storms[1:ibtracs_ni_results$prop_major_chpt_test[2]])/sum(ibtracs_ni_results$data$Total_Storms[1:ibtracs_ni_results$prop_major_chpt_test[2]])` and in regime two, the proportion of major storms is `r sum(ibtracs_ni_results$data$Major_Storms[(ibtracs_ni_results$prop_major_chpt_test[2]+1):45])/sum(ibtracs_ni_results$data$Total_Storms[(ibtracs_ni_results$prop_major_chpt_test[2]+1):45])`; a change of `r sum(ibtracs_ni_results$data$Major_Storms[(ibtracs_ni_results$prop_major_chpt_test[2]+1):45])/sum(ibtracs_ni_results$data$Total_Storms[(ibtracs_ni_results$prop_major_chpt_test[2]+1):45]) - sum(ibtracs_ni_results$data$Major_Storms[1:ibtracs_ni_results$prop_major_chpt_test[2]])/sum(ibtracs_ni_results$data$Total_Storms[1:ibtracs_ni_results$prop_major_chpt_test[2]])`.

#### Proportion of intense storms

A change in the proportion of major storms?
  
```{r, echo=FALSE}
ibtracs_ni_results$prop_intense_chpt_test
```

The change in proportions occurs `r 1979+ibtracs_ni_results$prop_intense_chpt_test[2]` to `r 1979+ibtracs_ni_results$prop_intense_chpt_test[2]+1` with $p$-value `r ibtracs_ni_results$prop_intense_chpt_test[3]`.


In regime one, the proportion of major storms is `r sum(ibtracs_ni_results$data$Intense_Storms[1:ibtracs_ni_results$prop_intense_chpt_test[2]])/sum(ibtracs_ni_results$data$Total_Storms[1:ibtracs_ni_results$prop_intense_chpt_test[2]])` and in regime two, the proportion of major storms is `r sum(ibtracs_ni_results$data$Intense_Storms[(ibtracs_ni_results$prop_intense_chpt_test[2]+1):45])/sum(ibtracs_ni_results$data$Total_Storms[(ibtracs_ni_results$prop_intense_chpt_test[2]+1):45])`; a change of `r sum(ibtracs_ni_results$data$Intense_Storms[(ibtracs_ni_results$prop_intense_chpt_test[2]+1):45])/sum(ibtracs_ni_results$data$Total_Storms[(ibtracs_ni_results$prop_intense_chpt_test[2]+1):45]) - sum(ibtracs_ni_results$data$Intense_Storms[1:ibtracs_ni_results$prop_intense_chpt_test[2]])/sum(ibtracs_ni_results$data$Total_Storms[1:ibtracs_ni_results$prop_intense_chpt_test[2]])`.


### Overall Change Point Findings

Basically, the total number of storms is not changing but there may be an increase in the number of major storms. Here we look at everything with the 2012-2013 changepoint location determined by the test for changes in major storms.


```{r, echo=FALSE}
bind_rows(
  ibtracs_all_basins |>
    dplyr::filter(BASIN=="Northern Indian") |>
    summarize(`Total Storms (avg)` = mean(Total_Storms),
              `Total Storms (stddev)` = sd(Total_Storms) ) |>
    mutate(Regime = "Entire Record"),
  ibtracs_all_basins |>
    dplyr::filter(BASIN=="Northern Indian") |>
    mutate(Regime = ifelse(SEASON < 2013, "1980-2012", "2013-2024") ) |>
    group_by(Regime) |>
    summarize(`Total Storms (avg)` = mean(Total_Storms),
              `Total Storms (stddev)` = sd(Total_Storms) ) 
)|>
  dplyr::select(Regime, `Total Storms (avg)`, `Total Storms (stddev)`) |>
  knitr::kable()
```

Not much happening here, a change of one storm per year which is why we did not detect anything statistically.

We do see a statistically meaningful change in proportion of major storms.

```{r, echo=FALSE}
bind_rows(
  ibtracs_all_basins |>
    dplyr::filter(BASIN=="Northern Indian") |>
    summarize(`Total Storms (avg)` = mean(Total_Storms),
              `Major Storms (avg)` = mean(Major_Storms),
              `Proportion Major` = sum(Major_Storms)/sum(Total_Storms)) |>
    mutate(Regime = "Entire Record"),
  ibtracs_all_basins |>
    dplyr::filter(BASIN=="Northern Indian") |>
    mutate(Regime = ifelse(SEASON < 2013, "1980-2012", "2013-2024") ) |>
    group_by(Regime) |>
    summarize(`Total Storms (avg)` = mean(Total_Storms),
              `Major Storms (avg)` = mean(Major_Storms),
              `Proportion Major` = sum(Major_Storms)/sum(Total_Storms)) ) |>
  dplyr::select(Regime, `Total Storms (avg)`, `Major Storms (avg)`, `Proportion Major`) |>
  knitr::kable()
```

Here we see a little bit of a story happening. About one storm per year increase overall, but that one storm has likely been a major hurricane. So maybe???  Nothing too revealing in this basin though given how small all the numbers are.

```{r, echo=FALSE}
ibtracs_ni_results$plot_seg_fits
```



## Full HURSAT Record
  
We now repeat the analysis with the full HURSAT record. Note, this data runs from 1981 through 2017.

```{r, echo=FALSE}
hursat_ni_results <- get_report_basin(basin="Northern Indian", all_basins=hursat_all_basins)
```

```{r, echo=FALSE}
hursat_ni_results$plot_data
```

### Summary Statistics (overall)

```{r, echo=FALSE}
hursat_ni_results$summary_stat |>
  knitr::kable()
```

### Regression Methods

#### Poisson Regression on non-Shorties

```{r, echo=FALSE}
summary(hursat_ni_results$pois_fit)
```

Note, this says a decrease of `r abs(coef(hursat_ni_results$pois_fit)[2])` log-storms per year. This corresponds to a `r (1-exp(coef(hursat_ni_results$pois_fit)[2]))*100` percent decrease per season. So over 37 years, that is a decrease of `r (1-exp(coef(hursat_ni_results$pois_fit)[2])^37)*100` percent.

Using this model, we would predict `r predict(hursat_ni_results$pois_fit, type="response")[1]` storms in 1981 and `r predict(hursat_ni_results$pois_fit, type="response")[37]` storms in 2017.

That is a change of `r (predict(hursat_ni_results$pois_fit, type="response")[37]-predict(hursat_ni_results$pois_fit, type="response")[1])/37*10` storms per decade.


#### Poisson Regression on Major Storms

```{r, echo=FALSE}
summary(hursat_ni_results$major_fit)
```


Using this model the number of a major storm in 1981 was `r predict(hursat_ni_results$major_fit, type="response")[1]` and `r predict(hursat_ni_results$major_fit, type="response")[37]` in 2017.

That is a change of `r (predict(hursat_ni_results$major_fit, type="response")[37]-predict(hursat_ni_results$major_fit, type="response")[1])/37*10` storms per decade.

#### Poisson Regression on Intense Storms

```{r, echo=FALSE}
summary(hursat_ni_results$intense_fit)
```

We see a significant increasing trend in the number of intense storms.

Using this model the number of a major storm in 1981 was `r predict(hursat_ni_results$intense_fit, type="response")[1]` and `r predict(hursat_ni_results$intense_fit, type="response")[37]` in 2017.

That is a change of `r (predict(hursat_ni_results$intense_fit, type="response")[37]-predict(hursat_ni_results$intense_fit, type="response")[1])/37*10` storms per decade.


#### Logistic Regression on Proportions

```{r, echo=FALSE}
summary(hursat_ni_results$major_logit_fit)
```


Specifically, using this model the probability of a major storm in 1981 was `r predict(hursat_ni_results$major_logit_fit, type="response")[1]` and `r predict(hursat_ni_results$major_logit_fit, type="response")[37]` in 2017.

That is a change of `r (predict(hursat_ni_results$major_logit_fit, type="response")[37]-predict(hursat_ni_results$major_logit_fit, type="response")[1])/37*10*100` percent per decade.


#### Logistic Regression on Proportions of Intense

```{r, echo=FALSE}
summary(hursat_ni_results$intense_logit_fit)
```


Specifically, using this model the probability of a major storm in 1981 was `r predict(hursat_ni_results$intense_logit_fit, type="response")[1]` and `r predict(hursat_ni_results$intense_logit_fit, type="response")[37]` in 2017.

That is a change of `r (predict(hursat_ni_results$intense_logit_fit, type="response")[37]-predict(hursat_ni_results$intense_logit_fit, type="response")[1])/37*10*100` percent per decade.


### Change Point Methods

#### Change in Non-Shorty Counts

```{r, echo=FALSE}
hursat_ni_results$pois_chpt_test
```

The dates suggest the changepoint would occur from `r 1980+hursat_ni_results$pois_chpt_test[2]` to `r 1980+hursat_ni_results$pois_chpt_test[2]+1` with $p$-value `r hursat_ni_results$pois_chpt_test[3]`.

In regime one, the mean number of total storms is `r mean(hursat_ni_results$data$Total_Storms[1:hursat_ni_results$pois_chpt_test[2]])` and in regime two, the mean number of storms is `r mean(hursat_ni_results$data$Total_Storms[(hursat_ni_results$pois_chpt_test[2]+1):37])`; a change of `r mean(hursat_ni_results$data$Total_Storms[(hursat_ni_results$pois_chpt_test[2]+1):37]) - mean(hursat_ni_results$data$Total_Storms[1:hursat_ni_results$pois_chpt_test[2]])`.

#### Change in counts of major storms

```{r, echo=FALSE}
hursat_ni_results$major_pois_chpt_test
```

The dates suggest the changepoint would occur from `r 1980+hursat_ni_results$major_pois_chpt_test[2]` to `r 1980+hursat_ni_results$major_pois_chpt_test[2]+1` with $p$-value `r hursat_ni_results$major_pois_chpt_test[3]`.

In regime one, the mean number of total storms is `r mean(hursat_ni_results$data$Major_Storms[1:hursat_ni_results$major_pois_chpt_test[2]])` and in regime two, the mean number of storms is `r mean(hursat_ni_results$data$Major_Storms[(hursat_ni_results$major_pois_chpt_test[2]+1):37])`; a change of `r mean(hursat_ni_results$data$Major_Storms[(hursat_ni_results$major_pois_chpt_test[2]+1):37]) - mean(hursat_ni_results$data$Major_Storms[1:hursat_ni_results$major_pois_chpt_test[2]])`.

#### Change in counts of intense storms

```{r, echo=FALSE}
hursat_ni_results$intense_pois_chpt_test
```

The dates suggest the changepoint would occur from `r 1980+hursat_ni_results$intense_pois_chpt_test[2]` to `r 1980+hursat_ni_results$intense_pois_chpt_test[2]+1` with $p$-value `r hursat_ni_results$intense_pois_chpt_test[3]`.

In regime one, the mean number of total storms is `r mean(hursat_ni_results$data$Intense_Storms[1:hursat_ni_results$intense_pois_chpt_test[2]])` and in regime two, the mean number of storms is `r mean(hursat_ni_results$data$Intense_Storms[(hursat_ni_results$intense_pois_chpt_test[2]+1):37])`; a change of `r mean(hursat_ni_results$data$Intense_Storms[(hursat_ni_results$intense_pois_chpt_test[2]+1):37]) - mean(hursat_ni_results$data$Intense_Storms[1:hursat_ni_results$intense_pois_chpt_test[2]])`.


#### Proportion of major storms

A change in the proportion of major storms?
  
```{r, echo=FALSE}
hursat_ni_results$prop_major_chpt_test
```

The change in proportions occurs `r 1979+hursat_ni_results$prop_major_chpt_test[2]` to `r 1979+hursat_ni_results$prop_major_chpt_test[2]+1` with $p$-value `r hursat_ni_results$prop_major_chpt_test[3]`. 

In regime one, the proportion of major storms is `r sum(hursat_ni_results$data$Major_Storms[1:hursat_ni_results$prop_major_chpt_test[2]])/sum(hursat_ni_results$data$Total_Storms[1:hursat_ni_results$prop_major_chpt_test[2]])` and in regime two, the proportion of major storms is `r sum(hursat_ni_results$data$Major_Storms[(hursat_ni_results$prop_major_chpt_test[2]+1):37])/sum(hursat_ni_results$data$Total_Storms[(hursat_ni_results$prop_major_chpt_test[2]+1):37])`; a change of `r sum(hursat_ni_results$data$Major_Storms[(hursat_ni_results$prop_major_chpt_test[2]+1):37])/sum(hursat_ni_results$data$Total_Storms[(hursat_ni_results$prop_major_chpt_test[2]+1):37]) - sum(hursat_ni_results$data$Major_Storms[1:hursat_ni_results$prop_major_chpt_test[2]])/sum(hursat_ni_results$data$Total_Storms[1:hursat_ni_results$prop_major_chpt_test[2]])`.



#### Proportion of intense storms

A change in the proportion of major storms?
  
```{r, echo=FALSE}
hursat_ni_results$prop_intense_chpt_test
```

The change in proportions occurs `r 1979+hursat_ni_results$prop_intense_chpt_test[2]` to `r 1979+hursat_ni_results$prop_intense_chpt_test[2]+1` with $p$-value `r hursat_ni_results$prop_intense_chpt_test[3]`.


In regime one, the proportion of major storms is `r sum(hursat_ni_results$data$Intense_Storms[1:hursat_ni_results$prop_intense_chpt_test[2]])/sum(hursat_ni_results$data$Total_Storms[1:hursat_ni_results$prop_intense_chpt_test[2]])` and in regime two, the proportion of major storms is `r sum(hursat_ni_results$data$Intense_Storms[(hursat_ni_results$prop_intense_chpt_test[2]+1):37])/sum(hursat_ni_results$data$Total_Storms[(hursat_ni_results$prop_intense_chpt_test[2]+1):37])`; a change of `r sum(hursat_ni_results$data$Intense_Storms[(hursat_ni_results$prop_intense_chpt_test[2]+1):37])/sum(hursat_ni_results$data$Total_Storms[(hursat_ni_results$prop_intense_chpt_test[2]+1):37]) - sum(hursat_ni_results$data$Intense_Storms[1:hursat_ni_results$prop_intense_chpt_test[2]])/sum(hursat_ni_results$data$Total_Storms[1:hursat_ni_results$prop_intense_chpt_test[2]])`.


### Overall Change Point Findings

The results here are nothing has changed.  This shouldn't be too surprising given the low counts and the results of IBTRacs -- the suggested change using that data was in 2012-2013.  The HURSAT data ends in 2017, so not a lot of data post 2013 to detect a change (even if there was one).




A table to compare the proportion of major storms.

```{r, echo=FALSE}
bind_rows(
  hursat_all_basins |>
    dplyr::filter(BASIN=="Northern Indian") |>
    summarize(`Total Storms (avg)` = mean(Total_Storms),
              `Major Storms (avg)` = mean(Major_Storms),
              `Proportion Major` = sum(Major_Storms)/sum(Total_Storms)) |>
    mutate(Regime = "Entire Record") ) |>
  dplyr::select(Regime, `Total Storms (avg)`, `Major Storms (avg)`, `Proportion Major`) |>
  knitr::kable()
```

This is a summary of what has happened in time - with no change.

```{r, echo=FALSE}
hursat_ni_results$plot_seg_fits
```


## Trimmed IBTRacs

We now repeat the analysis with the a version of the IBTRacs that has been trimmed to match HURSAT record. So this data runs from 1981 through 2017.

```{r, echo=FALSE}
trimmed_ni_results <- get_report_basin(basin="Northern Indian", all_basins=ibtracs_trimmed)
```


```{r, echo=FALSE}
trimmed_ni_results$plot_data
```


### Summary Statistics (overall)

```{r, echo=FALSE}
trimmed_ni_results$summary_stat |>
  knitr::kable()
```

### Regression Methods

#### Poisson Regression

```{r, echo=FALSE}
summary(trimmed_ni_results$pois_fit)
```

Note, this says a decrease of `r abs(coef(trimmed_ni_results$pois_fit)[2])` log-storms per year. This corresponds to a `r (1-exp(coef(trimmed_ni_results$pois_fit)[2]))*100` percent decrease per season. So over 37 years, that is a decrease of `r (1-exp(coef(trimmed_ni_results$pois_fit)[2])^37)*100` percent.

Using this model, we would predict `r predict(trimmed_ni_results$pois_fit, type="response")[1]` storms in 1981 and `r predict(trimmed_ni_results$pois_fit, type="response")[37]` storms in 2017

That is a change of `r (predict(trimmed_ni_results$pois_fit, type="response")[37]-predict(trimmed_ni_results$pois_fit, type="response")[1])/37*10` storms per decade.

#### Poisson Regression on Major Storms

```{r, echo=FALSE}
summary(trimmed_ni_results$major_fit)
```

We see a non-significant trend in the number of major storms.

Using this model the number of a major storm in 1981 was `r predict(trimmed_ni_results$major_fit, type="response")[1]` and `r predict(trimmed_ni_results$major_fit, type="response")[37]` in 2017.

That is a change of `r (predict(trimmed_ni_results$major_fit, type="response")[37]-predict(trimmed_ni_results$major_fit, type="response")[1])/37*10` storms per decade.


#### Poisson Regression on Intense Storms

```{r, echo=FALSE}
summary(trimmed_ni_results$intense_fit)
```

Using this model the number of a major storm in 1981 was `r predict(trimmed_ni_results$intense_fit, type="response")[1]` and `r predict(trimmed_ni_results$intense_fit, type="response")[37]` in 2017

That is a change of `r (predict(trimmed_ni_results$intense_fit, type="response")[37]-predict(trimmed_ni_results$intense_fit, type="response")[1])/37*10` storms per decade.


#### Logistic Regression on Proportions

```{r, echo=FALSE}
summary(trimmed_ni_results$major_logit_fit)
```


Specifically, using this model the probability of a major storm in 1981 was `r predict(trimmed_ni_results$major_logit_fit, type="response")[1]` and `r predict(trimmed_ni_results$major_logit_fit, type="response")[37]` in 2017.


That is a change of `r (predict(trimmed_ni_results$major_logit_fit, type="response")[37]-predict(trimmed_ni_results$major_logit_fit, type="response")[1])/37*10*100` percent per decade.



#### Logistic Regression on Proportions of Intense

```{r, echo=FALSE}
summary(trimmed_ni_results$intense_logit_fit)
```


Specifically, using this model the probability of a major storm in 1981 was `r predict(trimmed_ni_results$intense_logit_fit, type="response")[1]` and `r predict(trimmed_ni_results$intense_logit_fit, type="response")[37]` in 2017.

That is a change of `r (predict(trimmed_ni_results$intense_logit_fit, type="response")[37]-predict(trimmed_ni_results$intense_logit_fit, type="response")[1])/37*10*100` percent per decade.


### Change Point Methods

#### Change in Non-Shorty Counts

```{r, echo=FALSE}
trimmed_ni_results$pois_chpt_test
```

We have evidence the number of non-shorties has changed. The dates suggest the changepoint would occur from `r 1980+trimmed_ni_results$pois_chpt_test[2]` to `r 1980+trimmed_ni_results$pois_chpt_test[2]+1` with $p$-value `r hursat_ni_results$pois_chpt_test[3]`.

In regime one, the mean number of total storms is `r mean(trimmed_ni_results$data$Total_Storms[1:trimmed_ni_results$pois_chpt_test[2]])` and in regime two, the mean number of storms is `r mean(trimmed_ni_results$data$Total_Storms[(trimmed_ni_results$pois_chpt_test[2]+1):37])`; a change of `r mean(trimmed_ni_results$data$Total_Storms[(trimmed_ni_results$pois_chpt_test[2]+1):37]) - mean(trimmed_ni_results$data$Total_Storms[1:trimmed_ni_results$pois_chpt_test[2]])`.

#### Change in counts of major storms

```{r, echo=FALSE}
trimmed_ni_results$major_pois_chpt_test
```


The dates suggest the changepoint would occur from `r 1980+trimmed_ni_results$major_pois_chpt_test[2]` to `r 1980+trimmed_ni_results$major_pois_chpt_test[2]+1` with $p$-value `r trimmed_ni_results$major_pois_chpt_test[3]`.

In regime one, the mean number of total storms is `r mean(trimmed_ni_results$data$Major_Storms[1:trimmed_ni_results$major_pois_chpt_test[2]])` and in regime two, the mean number of storms is `r mean(trimmed_ni_results$data$Major_Storms[(trimmed_ni_results$major_pois_chpt_test[2]+1):37])`; a change of `r mean(trimmed_ni_results$data$Major_Storms[(trimmed_ni_results$major_pois_chpt_test[2]+1):37]) - mean(trimmed_ni_results$data$Major_Storms[1:trimmed_ni_results$major_pois_chpt_test[2]])`.


#### Change in counts of intense storms

```{r, echo=FALSE}
trimmed_ni_results$intense_pois_chpt_test
```


The dates suggest the changepoint would occur from `r 1980+trimmed_ni_results$intense_pois_chpt_test[2]` to `r 1980+trimmed_ni_results$intense_pois_chpt_test[2]+1` with $p$-value `r trimmed_ni_results$intense_pois_chpt_test[3]`.

In regime one, the mean number of total storms is `r mean(trimmed_ni_results$data$Intense_Storms[1:trimmed_ni_results$intense_pois_chpt_test[2]])` and in regime two, the mean number of storms is `r mean(trimmed_ni_results$data$Intense_Storms[(trimmed_ni_results$intense_pois_chpt_test[2]+1):37])`; a change of `r mean(trimmed_ni_results$data$Intense_Storms[(trimmed_ni_results$intense_pois_chpt_test[2]+1):37]) - mean(trimmed_ni_results$data$Intense_Storms[1:trimmed_ni_results$intense_pois_chpt_test[2]])`.




#### Proportion of major storms

A change in the proportion of major storms?
  
```{r, echo=FALSE}
trimmed_ni_results$prop_major_chpt_test
```


The change in proportions occurs `r 1979+trimmed_ni_results$prop_major_chpt_test[2]` to `r 1979+trimmed_ni_results$prop_major_chpt_test[2]+1` with $p$-value `r trimmed_ni_results$prop_major_chpt_test[3]`. 

In regime one, the proportion of major storms is `r sum(trimmed_ni_results$data$Major_Storms[1:trimmed_ni_results$prop_major_chpt_test[2]])/sum(trimmed_ni_results$data$Total_Storms[1:trimmed_ni_results$prop_major_chpt_test[2]])` and in regime two, the proportion of major storms is `r sum(trimmed_ni_results$data$Major_Storms[(trimmed_ni_results$prop_major_chpt_test[2]+1):37])/sum(trimmed_ni_results$data$Total_Storms[(trimmed_ni_results$prop_major_chpt_test[2]+1):37])`; a change of `r sum(trimmed_ni_results$data$Major_Storms[(trimmed_ni_results$prop_major_chpt_test[2]+1):37])/sum(trimmed_ni_results$data$Total_Storms[(trimmed_ni_results$prop_major_chpt_test[2]+1):37]) - sum(trimmed_ni_results$data$Major_Storms[1:trimmed_ni_results$prop_major_chpt_test[2]])/sum(trimmed_ni_results$data$Total_Storms[1:trimmed_ni_results$prop_major_chpt_test[2]])`.


#### Proportion of intense storms

A change in the proportion of major storms?
  
```{r, echo=FALSE}
trimmed_ni_results$prop_intense_chpt_test
```

The change in proportions occurs `r 1979+trimmed_ni_results$prop_intense_chpt_test[2]` to `r 1979+trimmed_ni_results$prop_intense_chpt_test[2]+1` with $p$-value `r trimmed_ni_results$prop_intense_chpt_test[3]`.


In regime one, the proportion of major storms is `r sum(trimmed_ni_results$data$Intense_Storms[1:trimmed_ni_results$prop_intense_chpt_test[2]])/sum(trimmed_ni_results$data$Total_Storms[1:trimmed_ni_results$prop_intense_chpt_test[2]])` and in regime two, the proportion of major storms is `r sum(trimmed_ni_results$data$Intense_Storms[(trimmed_ni_results$prop_intense_chpt_test[2]+1):37])/sum(trimmed_ni_results$data$Total_Storms[(trimmed_ni_results$prop_intense_chpt_test[2]+1):37])`; a change of `r sum(trimmed_ni_results$data$Intense_Storms[(trimmed_ni_results$prop_intense_chpt_test[2]+1):37])/sum(trimmed_ni_results$data$Total_Storms[(trimmed_ni_results$prop_intense_chpt_test[2]+1):37]) - sum(trimmed_ni_results$data$Intense_Storms[1:trimmed_ni_results$prop_intense_chpt_test[2]])/sum(trimmed_ni_results$data$Total_Storms[1:trimmed_ni_results$prop_intense_chpt_test[2]])`.




### Overall Change Point Findings


Consistent results as HURSAT -- no changes or trends. 


A table to compare the proportion of major storms.

```{r, echo=FALSE}
bind_rows(
  ibtracs_trimmed |>
    dplyr::filter(BASIN=="Northern Indian") |>
    summarize(`Total Storms (avg)` = mean(Total_Storms),
              `Major Storms (avg)` = mean(Major_Storms),
              `Proportion Major` = sum(Major_Storms)/sum(Total_Storms)) |>
    mutate(Regime = "Entire Record") ) |>
  dplyr::select(Regime, `Total Storms (avg)`, `Major Storms (avg)`, `Proportion Major`) |>
  knitr::kable()
```

```{r, echo=FALSE}
trimmed_ni_results$plot_seg_fits
```


## Overall Hypothesis Test Results

### Total Number of storms (non-shorties)
  
**Poisson Regression**

```{r, echo=FALSE}
data.frame(Data=c("IBTracs 1980-2024", "HURSAT 1981-2017", "IBTracs 1981-2017"),
           Coefficent = c(coef(ibtracs_ni_results$pois_fit)[2],coef(hursat_ni_results$pois_fit)[2],coef(trimmed_ni_results$pois_fit)[2]),
           `p-value` = c(anova(ibtracs_ni_results$pois_fit)$`Pr(>Chi)`[2], anova(hursat_ni_results$pois_fit)$`Pr(>Chi)`[2], anova(trimmed_ni_results$pois_fit)$`Pr(>Chi)`[2]),
           `Start` = c(predict(ibtracs_ni_results$pois_fit, type="response")[1], predict(hursat_ni_results$pois_fit, type="response")[1], predict(trimmed_ni_results$pois_fit, type="response")[1] ),
           `End` = c(predict(ibtracs_ni_results$pois_fit, type="response")[length(predict(ibtracs_ni_results$pois_fit, type="response"))], predict(hursat_ni_results$pois_fit, type="response")[length(predict(hursat_ni_results$pois_fit, type="response"))], predict(trimmed_ni_results$pois_fit, type="response")[length(predict(trimmed_ni_results$pois_fit, type="response"))] ),
           `Change per Decade` = c( (predict(ibtracs_ni_results$pois_fit, type="response")[length(predict(ibtracs_ni_results$pois_fit, type="response"))]-predict(ibtracs_ni_results$pois_fit, type="response")[1])/length(predict(ibtracs_ni_results$pois_fit, type="response"))*10,
                                    (predict(hursat_ni_results$pois_fit, type="response")[length(predict(hursat_ni_results$pois_fit, type="response"))]-predict(hursat_ni_results$pois_fit, type="response")[1])/length(predict(hursat_ni_results$pois_fit, type="response"))*10,
                                    (predict(trimmed_ni_results$pois_fit, type="response")[length(predict(trimmed_ni_results$pois_fit, type="response"))]-predict(trimmed_ni_results$pois_fit, type="response")[1])/length(predict(trimmed_ni_results$pois_fit, type="response"))*10 ) ) |>
  knitr::kable(digits=5)
```

**SCUSUM Change Point Test**

```{r, echo=FALSE}
data.frame(Data=c("IBTracs 1980-2024", "HURSAT 1981-2017", "IBTracs 1981-2017"),
           `Chpt Location` = c(ibtracs_ni_results$pois_chpt_test[2]+1980,hursat_ni_results$pois_chpt_test[2]+1981,trimmed_ni_results$pois_chpt_test[2]+1981),
           `p-value` =c(ibtracs_ni_results$pois_chpt_test[3],hursat_ni_results$pois_chpt_test[3],trimmed_ni_results$pois_chpt_test[3]),
           `Regime 1` = c(mean(ibtracs_ni_results$data$Total_Storms[1:ibtracs_ni_results$pois_chpt_test[2]]),
                          mean(hursat_ni_results$data$Total_Storms[1:hursat_ni_results$pois_chpt_test[2]]),
                          mean(trimmed_ni_results$data$Total_Storms[1:trimmed_ni_results$pois_chpt_test[2]]) ),
           `Regime 2` = c(mean(ibtracs_ni_results$data$Total_Storms[(ibtracs_ni_results$pois_chpt_test[2]+1):length(ibtracs_ni_results$data$Total_Storms)]),
                          mean(hursat_ni_results$data$Total_Storms[(hursat_ni_results$pois_chpt_test[2]+1):length(hursat_ni_results$data$Total_Storms)]),
                            mean(trimmed_ni_results$data$Total_Storms[(trimmed_ni_results$pois_chpt_test[2]+1):length(trimmed_ni_results$data$Total_Storms)]) ),
           `Change` = c(mean(ibtracs_ni_results$data$Total_Storms[(ibtracs_ni_results$pois_chpt_test[2]+1):length(ibtracs_ni_results$data$Total_Storms)]) - mean(ibtracs_ni_results$data$Total_Storms[1:ibtracs_ni_results$pois_chpt_test[2]]),
                        mean(hursat_ni_results$data$Total_Storms[(hursat_ni_results$pois_chpt_test[2]+1):length(hursat_ni_results$data$Total_Storms)]) - mean(hursat_ni_results$data$Total_Storms[1:hursat_ni_results$pois_chpt_test[2]]),
                        mean(trimmed_ni_results$data$Total_Storms[(trimmed_ni_results$pois_chpt_test[2]+1):length(trimmed_ni_results$data$Total_Storms)]) - mean(trimmed_ni_results$data$Total_Storms[1:trimmed_ni_results$pois_chpt_test[2]]) ) ) |>
  knitr::kable(digits=5)
```

### Major Storm Counts

**Poisson Regression**

```{r, echo=FALSE}
data.frame(Data=c("IBTracs 1980-2024", "HURSAT 1981-2017", "IBTracs 1981-2017"),
           Coefficent = c(coef(ibtracs_ni_results$major_fit)[2],
                          coef(hursat_ni_results$major_fit)[2],
                          coef(trimmed_ni_results$major_fit)[2]),
           `p-value` = c(anova(ibtracs_ni_results$major_fit)$`Pr(>Chi)`[2],
                         anova(hursat_ni_results$major_fit)$`Pr(>Chi)`[2],
                         anova(trimmed_ni_results$major_fit)$`Pr(>Chi)`[2]),
           `Start` = c(predict(ibtracs_ni_results$major_fit, type="response")[1], 
                       predict(hursat_ni_results$major_fit, type="response")[1], 
                       predict(trimmed_ni_results$major_fit, type="response")[1] ),
           `End` = c(predict(ibtracs_ni_results$major_fit,type="response")[length(predict(ibtracs_ni_results$major_fit, type="response"))],
                     predict(hursat_ni_results$major_fit, type="response")[length(predict(hursat_ni_results$major_fit, type="response"))], 
                     predict(trimmed_ni_results$major_fit, type="response")[length(predict(trimmed_ni_results$major_fit, type="response"))] ),
           `Change per Decade` = c( (predict(ibtracs_ni_results$major_fit, type="response")[length(predict(ibtracs_ni_results$major_fit, type="response"))]-predict(ibtracs_ni_results$major_fit, type="response")[1])/length(predict(ibtracs_ni_results$major_fit, type="response"))*10,
                                    (predict(hursat_ni_results$major_fit, type="response")[length(predict(hursat_ni_results$major_fit, type="response"))]-predict(hursat_ni_results$major_fit, type="response")[1])/length(predict(hursat_ni_results$major_fit, type="response"))*10,
                                    (predict(trimmed_ni_results$major_fit, type="response")[length(predict(trimmed_ni_results$major_fit, type="response"))]-predict(trimmed_ni_results$major_fit, type="response")[1])/length(predict(trimmed_ni_results$major_fit, type="response"))*10 ) ) |>
  knitr::kable(digits=5)
```

**SCUSUM Change Point Test**

```{r, echo=FALSE}
data.frame(Data=c("IBTracs 1980-2024", "HURSAT 1981-2017", "IBTracs 1981-2017"),
           `Chpt Location` = c(ibtracs_ni_results$major_pois_chpt_test[2]+1980,hursat_ni_results$major_pois_chpt_test[2]+1981,trimmed_ni_results$major_pois_chpt_test[2]+1981),
           `p-value` =c(ibtracs_ni_results$major_pois_chpt_test[3],hursat_ni_results$major_pois_chpt_test[3],trimmed_ni_results$major_pois_chpt_test[3]),
           `Regime 1` = c(mean(ibtracs_ni_results$data$Major_Storms[1:ibtracs_ni_results$major_pois_chpt_test[2]]),
                          mean(hursat_ni_results$data$Major_Storms[1:hursat_ni_results$major_pois_chpt_test[2]]),
                          mean(trimmed_ni_results$data$Major_Storms[1:trimmed_ni_results$major_pois_chpt_test[2]]) ),
           `Regime 2` = c(mean(ibtracs_ni_results$data$Major_Storms[(ibtracs_ni_results$major_pois_chpt_test[2]+1):length(ibtracs_ni_results$data$Major_Storms)]),
                          mean(hursat_ni_results$data$Major_Storms[(hursat_ni_results$major_pois_chpt_test[2]+1):length(hursat_ni_results$data$Major_Storms)]),
                            mean(trimmed_ni_results$data$Major_Storms[(trimmed_ni_results$major_pois_chpt_test[2]+1):length(trimmed_ni_results$data$Major_Storms)]) ),
           `Change` = c(mean(ibtracs_ni_results$data$Major_Storms[(ibtracs_ni_results$major_pois_chpt_test[2]+1):length(ibtracs_ni_results$data$Major_Storms)]) - mean(ibtracs_ni_results$data$Major_Storms[1:ibtracs_ni_results$major_pois_chpt_test[2]]),
                        mean(hursat_ni_results$data$Major_Storms[(hursat_ni_results$major_pois_chpt_test[2]+1):length(hursat_ni_results$data$Major_Storms)]) - mean(hursat_ni_results$data$Major_Storms[1:hursat_ni_results$major_pois_chpt_test[2]]),
                        mean(trimmed_ni_results$data$Major_Storms[(trimmed_ni_results$major_pois_chpt_test[2]+1):length(trimmed_ni_results$data$Major_Storms)]) - mean(trimmed_ni_results$data$Major_Storms[1:trimmed_ni_results$major_pois_chpt_test[2]]) ) ) |>
  knitr::kable(digits=5)
```

### Proportion of Major Storms

**Logistic Regression**

```{r, echo=FALSE}
data.frame(Data=c("IBTracs 1980-2024", "HURSAT 1981-2017", "IBTracs 1981-2017"),
           Coefficent = c(coef(ibtracs_ni_results$major_logit_fit)[2],coef(hursat_ni_results$major_logit_fit)[2],coef(trimmed_ni_results$major_logit_fit)[2]),
           `p-value` = c(anova(ibtracs_ni_results$major_logit_fit)$`Pr(>Chi)`[2], anova(hursat_ni_results$major_logit_fit)$`Pr(>Chi)`[2], anova(trimmed_ni_results$major_logit_fit)$`Pr(>Chi)`[2]),
                      `Start` = c(predict(ibtracs_ni_results$major_logit_fit, type="response")[1]*100, 
                       predict(hursat_ni_results$major_logit_fit, type="response")[1]*100, 
                       predict(trimmed_ni_results$major_logit_fit, type="response")[1]*100 ),
           `End` = c(predict(ibtracs_ni_results$major_logit_fit,type="response")[length(predict(ibtracs_ni_results$major_fit, type="response"))]*100,
                     predict(hursat_ni_results$major_logit_fit, type="response")[length(predict(hursat_ni_results$major_logit_fit, type="response"))]*100, 
                     predict(trimmed_ni_results$major_logit_fit, type="response")[length(predict(trimmed_ni_results$major_logit_fit, type="response"))]*100 ),
           `Change per Decade` = c( (predict(ibtracs_ni_results$major_logit_fit, type="response")[length(predict(ibtracs_ni_results$major_logit_fit, type="response"))]-predict(ibtracs_ni_results$major_logit_fit, type="response")[1])/length(predict(ibtracs_ni_results$major_logit_fit, type="response"))*10*100,
                                    (predict(hursat_ni_results$major_logit_fit, type="response")[length(predict(hursat_ni_results$major_logit_fit, type="response"))]-predict(hursat_ni_results$major_logit_fit, type="response")[1])/length(predict(hursat_ni_results$major_logit_fit, type="response"))*10*100,
                                    (predict(trimmed_ni_results$major_logit_fit, type="response")[length(predict(trimmed_ni_results$major_logit_fit, type="response"))]-predict(trimmed_ni_results$major_logit_fit, type="response")[1])/length(predict(trimmed_ni_results$major_logit_fit, type="response"))*10*100 ) ) |>
  knitr::kable()
```

**SCUSUM Change Point Test**

```{r, echo=FALSE}
data.frame(Data=c("IBTracs 1980-2024", "HURSAT 1981-2017", "IBTracs 1981-2017"),
           `Chpt Location` = c(ibtracs_ni_results$prop_major_chpt_test[2]+1980,hursat_ni_results$prop_major_chpt_test[2]+1981,trimmed_ni_results$prop_major_chpt_test[2]+1981),
           `p-value` =c(ibtracs_ni_results$prop_major_chpt_test[3],hursat_ni_results$prop_major_chpt_test[3],trimmed_ni_results$prop_major_chpt_test[3]),
           `Regime 1 %` = c(sum(ibtracs_ni_results$data$Major_Storms[1:ibtracs_ni_results$prop_major_chpt_test[2]])/sum(ibtracs_ni_results$data$Total_Storms[1:ibtracs_ni_results$prop_major_chpt_test[2]])*100,
                            sum(hursat_ni_results$data$Major_Storms[1:hursat_ni_results$prop_major_chpt_test[2]])/sum(hursat_ni_results$data$Total_Storms[1:hursat_ni_results$prop_major_chpt_test[2]])*100,
                            sum(ibtracs_ni_results$data$Major_Storms[1:ibtracs_ni_results$prop_major_chpt_test[2]])/sum(ibtracs_ni_results$data$Total_Storms[1:ibtracs_ni_results$prop_major_chpt_test[2]])*100 ),
           `Regime 2 %` = c(sum(ibtracs_ni_results$data$Major_Storms[(ibtracs_ni_results$prop_major_chpt_test[2]+1):length(ibtracs_ni_results$data$Major_Storms)])/sum(ibtracs_ni_results$data$Total_Storms[(ibtracs_ni_results$prop_major_chpt_test[2]+1):length(ibtracs_ni_results$data$Major_Storms)])*100,
                            sum(hursat_ni_results$data$Major_Storms[(hursat_ni_results$prop_major_chpt_test[2]+1):length(hursat_ni_results$data$Major_Storms)])/sum(hursat_ni_results$data$Total_Storms[(hursat_ni_results$prop_major_chpt_test[2]+1):length(hursat_ni_results$data$Major_Storms)])*100,
                            sum(trimmed_ni_results$data$Major_Storms[(trimmed_ni_results$prop_major_chpt_test[2]+1):length(trimmed_ni_results$data$Major_Storms)])/sum(trimmed_ni_results$data$Total_Storms[(trimmed_ni_results$prop_major_chpt_test[2]+1):length(trimmed_ni_results$data$Major_Storms)])*100 ),
           `Change %` = c((sum(ibtracs_ni_results$data$Major_Storms[(ibtracs_ni_results$prop_major_chpt_test[2]+1):length(ibtracs_ni_results$data$Major_Storms)])/sum(ibtracs_ni_results$data$Total_Storms[(ibtracs_ni_results$prop_major_chpt_test[2]+1):length(ibtracs_ni_results$data$Major_Storms)]) - sum(ibtracs_ni_results$data$Major_Storms[1:ibtracs_ni_results$prop_major_chpt_test[2]])/sum(ibtracs_ni_results$data$Total_Storms[1:ibtracs_ni_results$prop_major_chpt_test[2]]) )*100,
                          (sum(hursat_ni_results$data$Major_Storms[(hursat_ni_results$prop_major_chpt_test[2]+1):length(hursat_ni_results$data$Major_Storms)])/sum(hursat_ni_results$data$Total_Storms[(hursat_ni_results$prop_major_chpt_test[2]+1):length(hursat_ni_results$data$Major_Storms)]) - sum(hursat_ni_results$data$Major_Storms[1:hursat_ni_results$prop_major_chpt_test[2]])/sum(hursat_ni_results$data$Total_Storms[1:hursat_ni_results$prop_major_chpt_test[2]]) )*100,
                          (sum(trimmed_ni_results$data$Major_Storms[(trimmed_ni_results$prop_major_chpt_test[2]+1):length(trimmed_ni_results$data$Major_Storms)])/sum(trimmed_ni_results$data$Total_Storms[(trimmed_ni_results$prop_major_chpt_test[2]+1):length(trimmed_ni_results$data$Major_Storms)]) - sum(trimmed_ni_results$data$Major_Storms[1:trimmed_ni_results$prop_major_chpt_test[2]])/sum(trimmed_ni_results$data$Total_Storms[1:trimmed_ni_results$prop_major_chpt_test[2]]) )*100 ) ) |>
  knitr::kable()
```



### Intense Storm Counts

**Poisson Regression**

```{r, echo=FALSE}
data.frame(Data=c("IBTracs 1980-2024", "HURSAT 1981-2017", "IBTracs 1981-2017"),
           Coefficent = c(coef(ibtracs_ni_results$intense_fit)[2],
                          coef(hursat_ni_results$intense_fit)[2],
                          coef(trimmed_ni_results$intense_fit)[2]),
           `p-value` = c(anova(ibtracs_ni_results$intense_fit)$`Pr(>Chi)`[2],
                         anova(hursat_ni_results$intense_fit)$`Pr(>Chi)`[2],
                         anova(trimmed_ni_results$intense_fit)$`Pr(>Chi)`[2]),
           `Start` = c(predict(ibtracs_ni_results$intense_fit, type="response")[1], 
                       predict(hursat_ni_results$intense_fit, type="response")[1], 
                       predict(trimmed_ni_results$intense_fit, type="response")[1] ),
           `End` = c(predict(ibtracs_ni_results$intense_fit,type="response")[length(predict(ibtracs_ni_results$intense_fit, type="response"))],
                     predict(hursat_ni_results$intense_fit, type="response")[length(predict(hursat_ni_results$intense_fit, type="response"))], 
                     predict(trimmed_ni_results$intense_fit, type="response")[length(predict(trimmed_ni_results$intense_fit, type="response"))] ),
           `Change per Decade` = c( (predict(ibtracs_ni_results$intense_fit, type="response")[length(predict(ibtracs_ni_results$intense_fit, type="response"))]-predict(ibtracs_ni_results$intense_fit, type="response")[1])/length(predict(ibtracs_ni_results$intense_fit, type="response"))*10,
                                    (predict(hursat_ni_results$intense_fit, type="response")[length(predict(hursat_ni_results$intense_fit, type="response"))]-predict(hursat_ni_results$intense_fit, type="response")[1])/length(predict(hursat_ni_results$intense_fit, type="response"))*10,
                                    (predict(trimmed_ni_results$intense_fit, type="response")[length(predict(trimmed_ni_results$intense_fit, type="response"))]-predict(trimmed_ni_results$intense_fit, type="response")[1])/length(predict(trimmed_ni_results$intense_fit, type="response"))*10 ) ) |>
  knitr::kable(digits=5)
```

**SCUSUM Change Point Test**

```{r, echo=FALSE}
data.frame(Data=c("IBTracs 1980-2024", "HURSAT 1981-2017", "IBTracs 1981-2017"),
           `Chpt Location` = c(ibtracs_ni_results$intense_pois_chpt_test[2]+1980,hursat_ni_results$intense_pois_chpt_test[2]+1981,trimmed_ni_results$intense_pois_chpt_test[2]+1981),
           `p-value` =c(ibtracs_ni_results$intense_pois_chpt_test[3],hursat_ni_results$intense_pois_chpt_test[3],trimmed_ni_results$intense_pois_chpt_test[3]),
           `Regime 1` = c(mean(ibtracs_ni_results$data$Intense_Storms[1:ibtracs_ni_results$intense_pois_chpt_test[2]]),
                          mean(hursat_ni_results$data$Intense_Storms[1:hursat_ni_results$intense_pois_chpt_test[2]]),
                          mean(trimmed_ni_results$data$Intense_Storms[1:trimmed_ni_results$intense_pois_chpt_test[2]]) ),
           `Regime 2` = c(mean(ibtracs_ni_results$data$Intense_Storms[(ibtracs_ni_results$intense_pois_chpt_test[2]+1):length(ibtracs_ni_results$data$Intense_Storms)]),
                          mean(hursat_ni_results$data$Intense_Storms[(hursat_ni_results$intense_pois_chpt_test[2]+1):length(hursat_ni_results$data$Intense_Storms)]),
                            mean(trimmed_ni_results$data$Intense_Storms[(trimmed_ni_results$intense_pois_chpt_test[2]+1):length(trimmed_ni_results$data$Intense_Storms)]) ),
           `Change` = c(mean(ibtracs_ni_results$data$Intense_Storms[(ibtracs_ni_results$intense_pois_chpt_test[2]+1):length(ibtracs_ni_results$data$Intense_Storms)]) - mean(ibtracs_ni_results$data$Intense_Storms[1:ibtracs_ni_results$intense_pois_chpt_test[2]]),
                        mean(hursat_ni_results$data$Intense_Storms[(hursat_ni_results$intense_pois_chpt_test[2]+1):length(hursat_ni_results$data$Intense_Storms)]) - mean(hursat_ni_results$data$Intense_Storms[1:hursat_ni_results$intense_pois_chpt_test[2]]),
                        mean(trimmed_ni_results$data$Intense_Storms[(trimmed_ni_results$intense_pois_chpt_test[2]+1):length(trimmed_ni_results$data$Intense_Storms)]) - mean(trimmed_ni_results$data$Intense_Storms[1:trimmed_ni_results$intense_pois_chpt_test[2]]) ) ) |>
  knitr::kable(digits=5)
```

### Proportion of Intense Storms

**Logistic Regression**

```{r, echo=FALSE}
data.frame(Data=c("IBTracs 1980-2024", "HURSAT 1981-2017", "IBTracs 1981-2017"),
           Coefficent = c(coef(ibtracs_ni_results$intense_logit_fit)[2],coef(hursat_ni_results$intense_logit_fit)[2],coef(trimmed_ni_results$intense_logit_fit)[2]),
           `p-value` = c(anova(ibtracs_ni_results$intense_logit_fit)$`Pr(>Chi)`[2], anova(hursat_ni_results$intense_logit_fit)$`Pr(>Chi)`[2], anova(trimmed_ni_results$intense_logit_fit)$`Pr(>Chi)`[2]),
                      `Start` = c(predict(ibtracs_ni_results$intense_logit_fit, type="response")[1]*100, 
                       predict(hursat_ni_results$intense_logit_fit, type="response")[1]*100, 
                       predict(trimmed_ni_results$intense_logit_fit, type="response")[1]*100 ),
           `End` = c(predict(ibtracs_ni_results$intense_logit_fit,type="response")[length(predict(ibtracs_ni_results$intense_fit, type="response"))]*100,
                     predict(hursat_ni_results$intense_logit_fit, type="response")[length(predict(hursat_ni_results$intense_logit_fit, type="response"))]*100, 
                     predict(trimmed_ni_results$intense_logit_fit, type="response")[length(predict(trimmed_ni_results$intense_logit_fit, type="response"))]*100 ),
           `Change per Decade` = c( (predict(ibtracs_ni_results$intense_logit_fit, type="response")[length(predict(ibtracs_ni_results$intense_logit_fit, type="response"))]-predict(ibtracs_ni_results$intense_logit_fit, type="response")[1])/length(predict(ibtracs_ni_results$intense_logit_fit, type="response"))*10*100,
                                    (predict(hursat_ni_results$intense_logit_fit, type="response")[length(predict(hursat_ni_results$intense_logit_fit, type="response"))]-predict(hursat_ni_results$intense_logit_fit, type="response")[1])/length(predict(hursat_ni_results$intense_logit_fit, type="response"))*10*100,
                                    (predict(trimmed_ni_results$intense_logit_fit, type="response")[length(predict(trimmed_ni_results$intense_logit_fit, type="response"))]-predict(trimmed_ni_results$intense_logit_fit, type="response")[1])/length(predict(trimmed_ni_results$intense_logit_fit, type="response"))*10*100 ) ) |>
  knitr::kable()
```

**SCUSUM Change Point Test**

```{r, echo=FALSE}
data.frame(Data=c("IBTracs 1980-2024", "HURSAT 1981-2017", "IBTracs 1981-2017"),
           `Chpt Location` = c(ibtracs_ni_results$prop_intense_chpt_test[2]+1980,hursat_ni_results$prop_intense_chpt_test[2]+1981,trimmed_ni_results$prop_intense_chpt_test[2]+1981),
           `p-value` =c(ibtracs_ni_results$prop_intense_chpt_test[3],hursat_ni_results$prop_intense_chpt_test[3],trimmed_ni_results$prop_intense_chpt_test[3]),
           `Regime 1 %` = c(sum(ibtracs_ni_results$data$Intense_Storms[1:ibtracs_ni_results$prop_intense_chpt_test[2]])/sum(ibtracs_ni_results$data$Total_Storms[1:ibtracs_ni_results$prop_intense_chpt_test[2]])*100,
                          sum(hursat_ni_results$data$Intense_Storms[1:hursat_ni_results$prop_intense_chpt_test[2]])/sum(hursat_ni_results$data$Total_Storms[1:hursat_ni_results$prop_intense_chpt_test[2]])*100,
                          sum(ibtracs_ni_results$data$Intense_Storms[1:ibtracs_ni_results$prop_intense_chpt_test[2]])/sum(ibtracs_ni_results$data$Total_Storms[1:ibtracs_ni_results$prop_intense_chpt_test[2]])*100 ),
           `Regime 2 %` = c(sum(ibtracs_ni_results$data$Intense_Storms[(ibtracs_ni_results$prop_intense_chpt_test[2]+1):length(ibtracs_ni_results$data$Intense_Storms)])/sum(ibtracs_ni_results$data$Total_Storms[(ibtracs_ni_results$prop_intense_chpt_test[2]+1):length(ibtracs_ni_results$data$Intense_Storms)])*100,
                          sum(hursat_ni_results$data$Intense_Storms[(hursat_ni_results$prop_intense_chpt_test[2]+1):length(hursat_ni_results$data$Intense_Storms)])/sum(hursat_ni_results$data$Total_Storms[(hursat_ni_results$prop_intense_chpt_test[2]+1):length(hursat_ni_results$data$Intense_Storms)])*100,
                          sum(trimmed_ni_results$data$Intense_Storms[(trimmed_ni_results$prop_intense_chpt_test[2]+1):length(trimmed_ni_results$data$Intense_Storms)])/sum(trimmed_ni_results$data$Total_Storms[(trimmed_ni_results$prop_intense_chpt_test[2]+1):length(trimmed_ni_results$data$Intense_Storms)])*100 ),
           `Change %` = c((sum(ibtracs_ni_results$data$Intense_Storms[(ibtracs_ni_results$prop_intense_chpt_test[2]+1):length(ibtracs_ni_results$data$Intense_Storms)])/sum(ibtracs_ni_results$data$Total_Storms[(ibtracs_ni_results$prop_intense_chpt_test[2]+1):length(ibtracs_ni_results$data$Intense_Storms)]) - sum(ibtracs_ni_results$data$Intense_Storms[1:ibtracs_ni_results$prop_intense_chpt_test[2]])/sum(ibtracs_ni_results$data$Total_Storms[1:ibtracs_ni_results$prop_intense_chpt_test[2]]) )*100,
                        (sum(hursat_ni_results$data$Intense_Storms[(hursat_ni_results$prop_intense_chpt_test[2]+1):length(hursat_ni_results$data$Intense_Storms)])/sum(hursat_ni_results$data$Total_Storms[(hursat_ni_results$prop_intense_chpt_test[2]+1):length(hursat_ni_results$data$Intense_Storms)]) - sum(hursat_ni_results$data$Intense_Storms[1:hursat_ni_results$prop_intense_chpt_test[2]])/sum(hursat_ni_results$data$Total_Storms[1:hursat_ni_results$prop_intense_chpt_test[2]]) )*100,
                        (sum(trimmed_ni_results$data$Intense_Storms[(trimmed_ni_results$prop_intense_chpt_test[2]+1):length(trimmed_ni_results$data$Intense_Storms)])/sum(trimmed_ni_results$data$Total_Storms[(trimmed_ni_results$prop_intense_chpt_test[2]+1):length(trimmed_ni_results$data$Intense_Storms)]) - sum(trimmed_ni_results$data$Intense_Storms[1:trimmed_ni_results$prop_intense_chpt_test[2]])/sum(trimmed_ni_results$data$Total_Storms[1:trimmed_ni_results$prop_intense_chpt_test[2]]) )*100 ) ) |>
  knitr::kable()
```

