
# Global Cyclone Record

We begin with a study of the record across the entire globe -- the Global "basin". Our analysis looks at all non-shorties (storms that lasted at least 48 hours) and achieved tropical storm strength. We also look at those storms that achieved major hurricane strength.

## Full IBTracs Record

```{r, echo=FALSE}
ibtracs_global_results <- get_report_basin(basin="Global", all_basins=ibtracs_all_basins)
```

```{r, echo=FALSE}
ibtracs_global_results$plot_data
```


### Summary Statistics (overall)

```{r, echo=FALSE}
ibtracs_global_results$summary_stat |>
  knitr::kable()
```

### Regression Methods

#### Poisson Regression on non-Shorties

```{r, echo=FALSE}
summary(ibtracs_global_results$pois_fit)
```



Note, this says a decrease of `r abs(coef(ibtracs_global_results$pois_fit)[2])` log-storms per year. This corresponds to a `r (1-exp(coef(ibtracs_global_results$pois_fit)[2]))*100` percent decrease per season. So over `r num_years_ibtracs` years, that is a decrease of `r (1-exp(coef(ibtracs_global_results$pois_fit)[2])^num_years_ibtracs)*100` percent.

Using this model, we would predict `r predict(ibtracs_global_results$pois_fit, type="response")[1]` storms in 1980 and `r predict(ibtracs_global_results$pois_fit, type="response")[num_years_ibtracs]` storms in 2025.

That is a change of `r (predict(ibtracs_global_results$pois_fit, type="response")[num_years_ibtracs]-predict(ibtracs_global_results$pois_fit, type="response")[1])/num_years_ibtracs*10` storms per decade.

#### Poisson Regression on Major Storms

```{r, echo=FALSE}
summary(ibtracs_global_results$major_fit)
```

Using this model the number of a major storm in 1980 was `r predict(ibtracs_global_results$major_fit, type="response")[1]` and `r predict(ibtracs_global_results$major_fit, type="response")[num_years_ibtracs]` in 2025.

That is a change of `r (predict(ibtracs_global_results$major_fit, type="response")[num_years_ibtracs]-predict(ibtracs_global_results$major_fit, type="response")[1])/num_years_ibtracs*10` storms per decade.

#### Poisson Regression on Intense Storms

```{r, echo=FALSE}
summary(ibtracs_global_results$intense_fit)
```


Using this model the number of a major storm in 1980 was `r predict(ibtracs_global_results$intense_fit, type="response")[1]` and `r predict(ibtracs_global_results$intense_fit, type="response")[num_years_ibtracs]` in 2025.

That is a change of `r (predict(ibtracs_global_results$intense_fit, type="response")[num_years_ibtracs]-predict(ibtracs_global_results$intense_fit, type="response")[1])/num_years_ibtracs*10` storms per decade.

#### Logistic Regression on Proportions of Major

```{r, echo=FALSE}
summary(ibtracs_global_results$major_logit_fit)
```

We see a significant result with a positive coefficient, so the log-odds of major storms has increased. This tells us the proportion of major storms has increased.

Specifically, using this model the probability of a major storm in 1980 was `r predict(ibtracs_global_results$major_logit_fit, type="response")[1]` and `r predict(ibtracs_global_results$major_logit_fit, type="response")[num_years_ibtracs]` in 2025.

That is a change of `r (predict(ibtracs_global_results$major_logit_fit, type="response")[num_years_ibtracs]-predict(ibtracs_global_results$major_logit_fit, type="response")[1])/num_years_ibtracs*10*100` percent per decade.

#### Logistic Regression on Proportions of Intense

```{r, echo=FALSE}
summary(ibtracs_global_results$intense_logit_fit)
```


Specifically, using this model the probability of a major storm in 1980 was `r predict(ibtracs_global_results$intense_logit_fit, type="response")[1]` and `r predict(ibtracs_global_results$intense_logit_fit, type="response")[num_years_ibtracs]` in 2025.

That is a change of `r (predict(ibtracs_global_results$intense_logit_fit, type="response")[num_years_ibtracs]-predict(ibtracs_global_results$intense_logit_fit, type="response")[1])/num_years_ibtracs*10*100` percent per decade.

### Change Point Methods

#### Change in Non-Shorty Counts

```{r, echo=FALSE}
ibtracs_global_results$pois_chpt_test
```

The dates suggest the changepoint would occur from `r 1979+ibtracs_global_results$pois_chpt_test[2]` to `r 1979+ibtracs_global_results$pois_chpt_test[2]+1` with $p$-value `r ibtracs_global_results$pois_chpt_test[3]`.

In regime one, the mean number of total storms is `r mean(ibtracs_global_results$data$Total_Storms[1:ibtracs_global_results$pois_chpt_test[2]])` and in regime two, the mean number of storms is `r mean(ibtracs_global_results$data$Total_Storms[(ibtracs_global_results$pois_chpt_test[2]+1):num_years_ibtracs])`; a change of `r mean(ibtracs_global_results$data$Total_Storms[(ibtracs_global_results$pois_chpt_test[2]+1):num_years_ibtracs]) - mean(ibtracs_global_results$data$Total_Storms[1:ibtracs_global_results$pois_chpt_test[2]])`.

#### Change in counts of major storms

```{r, echo=FALSE}
ibtracs_global_results$major_pois_chpt_test
```

The dates suggest the changepoint would occur from `r 1979+ibtracs_global_results$major_pois_chpt_test[2]` to `r 1979+ibtracs_global_results$major_pois_chpt_test[2]+1` with $p$-value `r ibtracs_global_results$major_pois_chpt_test[3]`.

In regime one, the mean number of major storms is `r mean(ibtracs_global_results$data$Major_Storms[1:ibtracs_global_results$major_pois_chpt_test[2]])` and in regime two, the mean number of storms is `r mean(ibtracs_global_results$data$Major_Storms[(ibtracs_global_results$major_pois_chpt_test[2]+1):num_years_ibtracs])`; a change of `r mean(ibtracs_global_results$data$Major_Storms[(ibtracs_global_results$major_pois_chpt_test[2]+1):num_years_ibtracs]) - mean(ibtracs_global_results$data$Major_Storms[1:ibtracs_global_results$major_pois_chpt_test[2]])`.

#### Change in counts of intense storms

```{r, echo=FALSE}
ibtracs_global_results$intense_pois_chpt_test
```


The dates suggest the changepoint would occur from `r 1979+ibtracs_global_results$intense_pois_chpt_test[2]` to `r 1979+ibtracs_global_results$intense_pois_chpt_test[2]+1` with $p$-value `r ibtracs_global_results$intense_pois_chpt_test[3]`.

In regime one, the mean number of intense storms is `r mean(ibtracs_global_results$data$Intense_Storms[1:ibtracs_global_results$intense_pois_chpt_test[2]])` and in regime two, the mean number of storms is `r mean(ibtracs_global_results$data$Intense_Storms[(ibtracs_global_results$intense_pois_chpt_test[2]+1):num_years_ibtracs])`; a change of `r mean(ibtracs_global_results$data$Intense_Storms[(ibtracs_global_results$intense_pois_chpt_test[2]+1):num_years_ibtracs]) - mean(ibtracs_global_results$data$Intense_Storms[1:ibtracs_global_results$intense_pois_chpt_test[2]])`.

#### Proportion of major storms

A change in the proportion of major storms?
  
```{r, echo=FALSE}
ibtracs_global_results$prop_major_chpt_test
```

The change in proportions occurs `r 1979+ibtracs_global_results$prop_major_chpt_test[2]` to `r 1979+ibtracs_global_results$prop_major_chpt_test[2]+1` with $p$-value `r ibtracs_global_results$prop_major_chpt_test[3]`. 

In regime one, the proportion of major storms is `r sum(ibtracs_global_results$data$Major_Storms[1:ibtracs_global_results$prop_major_chpt_test[2]])/sum(ibtracs_global_results$data$Total_Storms[1:ibtracs_global_results$prop_major_chpt_test[2]])` and in regime two, the proportion of major storms is `r sum(ibtracs_global_results$data$Major_Storms[(ibtracs_global_results$prop_major_chpt_test[2]+1):num_years_ibtracs])/sum(ibtracs_global_results$data$Total_Storms[(ibtracs_global_results$prop_major_chpt_test[2]+1):num_years_ibtracs])`; a change of `r sum(ibtracs_global_results$data$Major_Storms[(ibtracs_global_results$prop_major_chpt_test[2]+1):num_years_ibtracs])/sum(ibtracs_global_results$data$Total_Storms[(ibtracs_global_results$prop_major_chpt_test[2]+1):num_years_ibtracs]) - sum(ibtracs_global_results$data$Major_Storms[1:ibtracs_global_results$prop_major_chpt_test[2]])/sum(ibtracs_global_results$data$Total_Storms[1:ibtracs_global_results$prop_major_chpt_test[2]])`.

#### Proportion of intense storms

A change in the proportion of major storms?
  
```{r, echo=FALSE}
ibtracs_global_results$prop_intense_chpt_test
```

The change in proportions occurs `r 1979+ibtracs_global_results$prop_intense_chpt_test[2]` to `r 1979+ibtracs_global_results$prop_intense_chpt_test[2]+1` with $p$-value `r ibtracs_global_results$prop_intense_chpt_test[3]`.


In regime one, the proportion of major storms is `r sum(ibtracs_global_results$data$Intense_Storms[1:ibtracs_global_results$prop_intense_chpt_test[2]])/sum(ibtracs_global_results$data$Total_Storms[1:ibtracs_global_results$prop_intense_chpt_test[2]])` and in regime two, the proportion of major storms is `r sum(ibtracs_global_results$data$Intense_Storms[(ibtracs_global_results$prop_intense_chpt_test[2]+1):num_years_ibtracs])/sum(ibtracs_global_results$data$Total_Storms[(ibtracs_global_results$prop_intense_chpt_test[2]+1):num_years_ibtracs])`; a change of `r sum(ibtracs_global_results$data$Intense_Storms[(ibtracs_global_results$prop_intense_chpt_test[2]+1):num_years_ibtracs])/sum(ibtracs_global_results$data$Total_Storms[(ibtracs_global_results$prop_intense_chpt_test[2]+1):num_years_ibtracs]) - sum(ibtracs_global_results$data$Intense_Storms[1:ibtracs_global_results$prop_intense_chpt_test[2]])/sum(ibtracs_global_results$data$Total_Storms[1:ibtracs_global_results$prop_intense_chpt_test[2]])`.


### Overall Change Point Findings

Basically, the total number of storms is decreasing a bit while the number of major storms is more-or-less constant, so therefore the proportion of major storms has increased. The number of intense storms appears to have increased. 

A very suggestive result of an overall change to the number of non-shorties. So for exploration look at how counts changed in the potential change point time.

```{r, echo=FALSE}
bind_rows(
  ibtracs_all_basins |>
    dplyr::filter(BASIN=="Global") |>
    summarize(`Total Storms (avg)` = mean(Total_Storms),
              `Total Storms (stddev)` = sd(Total_Storms) ) |>
    mutate(Regime = "Entire Record"),
  ibtracs_all_basins |>
    dplyr::filter(BASIN=="Global") |>
    mutate(Regime = ifelse(SEASON < 1998, "1980-1997", "1998-2025") ) |>
    group_by(Regime) |>
    summarize(`Total Storms (avg)` = mean(Total_Storms),
              `Total Storms (stddev)` = sd(Total_Storms) ) 
)|>
  dplyr::select(Regime, `Total Storms (avg)`, `Total Storms (stddev)`) |>
  knitr::kable()
```

We do see a decrease of nearly 7 storms per year in the 1990s.

We do see a statistically meaningful change in proportion of major storms and note the changepoint location was at a different point.

```{r, echo=FALSE}
bind_rows(
  ibtracs_all_basins |>
    dplyr::filter(BASIN=="Global") |>
    summarize(`Total Storms (avg)` = mean(Total_Storms),
              `Major Storms (avg)` = mean(Major_Storms),
              `Proportion Major` = sum(Major_Storms)/sum(Total_Storms)) |>
    mutate(Regime = "Entire Record"),
  ibtracs_all_basins |>
    dplyr::filter(BASIN=="Global") |>
    mutate(Regime = ifelse(SEASON < 1991, "1980-1990", "1991-2025") ) |>
    group_by(Regime) |>
    summarize(`Total Storms (avg)` = mean(Total_Storms),
              `Major Storms (avg)` = mean(Major_Storms),
              `Proportion Major` = sum(Major_Storms)/sum(Total_Storms)) ) |>
  dplyr::select(Regime, `Total Storms (avg)`, `Major Storms (avg)`, `Proportion Major`) |>
  knitr::kable()
```

Proportion of major storms increased by almost 7%. The numbers are fairly convincing when looking at these two regimes. The total number of storms dropped by about 5 (albeit, we found that not really significant) but the number of major storms actually increased across these two regimes.
  
  
```{r, echo=FALSE}
ibtracs_global_results$plot_seg_fits
```


For the intense storms, we found a change point a year later


```{r, echo=FALSE}
bind_rows(
  ibtracs_all_basins |>
    dplyr::filter(BASIN=="Global") |>
    summarize(`Total Storms (avg)` = mean(Total_Storms),
              `Intense Storms (avg)` = mean(Intense_Storms),
              `Proportion Intense` = sum(Intense_Storms)/sum(Total_Storms)) |>
    mutate(Regime = "Entire Record"),
  ibtracs_all_basins |>
    dplyr::filter(BASIN=="Global") |>
    mutate(Regime = ifelse(SEASON < 1992, "1980-1991", "1992-2025") ) |>
    group_by(Regime) |>
    summarize(`Total Storms (avg)` = mean(Total_Storms),
              `Intense Storms (avg)` = mean(Intense_Storms),
              `Proportion Intense` = sum(Intense_Storms)/sum(Total_Storms)) ) |>
  dplyr::select(Regime, `Total Storms (avg)`, `Intense Storms (avg)`, `Proportion Intense`) |>
  knitr::kable()
```

The proportion of intense storms has increased quite a bit since early in the record.



## Full HURSAT Record
  
We now repeat the analysis with the full HURSAT record. Note, this data runs from 1981 through 2017.

```{r, echo=FALSE}
hursat_global_results <- get_report_basin(basin="Global", all_basins=hursat_all_basins)
```


```{r, echo=FALSE}
hursat_global_results$plot_data
```


### Summary Statistics (overall)

```{r, echo=FALSE}
hursat_global_results$summary_stat |>
  knitr::kable()
```

### Regression Methods

#### Poisson Regression on non-Shorties

```{r, echo=FALSE}
summary(hursat_global_results$pois_fit)
```

Note, this says a decrease of `r abs(coef(hursat_global_results$pois_fit)[2])` log-storms per year. This corresponds to a `r (1-exp(coef(hursat_global_results$pois_fit)[2]))*100` percent decrease per season. So over `r num_years_hursat` years, that is a decrease of `r (1-exp(coef(hursat_global_results$pois_fit)[2])^num_years_hursat)*100` percent.

Using this model, we would predict `r predict(hursat_global_results$pois_fit, type="response")[1]` storms in 1981 and `r predict(hursat_global_results$pois_fit, type="response")[num_years_hursat]` storms in 2017.

That is a change of `r (predict(hursat_global_results$pois_fit, type="response")[num_years_hursat]-predict(hursat_global_results$pois_fit, type="response")[1])/num_years_hursat*10` storms per decade.


#### Poisson Regression on Major Storms

```{r, echo=FALSE}
summary(hursat_global_results$major_fit)
```


Using this model the number of a major storm in 1981 was `r predict(hursat_global_results$major_fit, type="response")[1]` and `r predict(hursat_global_results$major_fit, type="response")[num_years_hursat]` in 2017.

That is a change of `r (predict(hursat_global_results$major_fit, type="response")[num_years_hursat]-predict(hursat_global_results$major_fit, type="response")[1])/num_years_hursat*10` storms per decade.

#### Poisson Regression on Intense Storms

```{r, echo=FALSE}
summary(hursat_global_results$intense_fit)
```

We see a non-significant increasing trend in the number of intense storms.

Using this model the number of a intense storm in 1981 was `r predict(hursat_global_results$intense_fit, type="response")[1]` and `r predict(hursat_global_results$intense_fit, type="response")[num_years_hursat]` in 2017.

That is a change of `r (predict(hursat_global_results$intense_fit, type="response")[num_years_hursat]-predict(hursat_global_results$intense_fit, type="response")[1])/num_years_hursat*10` storms per decade.


#### Logistic Regression on Proportions

```{r, echo=FALSE}
summary(hursat_global_results$major_logit_fit)
```


Specifically, using this model the probability of a major storm in 1981 was `r predict(hursat_global_results$major_logit_fit, type="response")[1]` and `r predict(hursat_global_results$major_logit_fit, type="response")[num_years_hursat]` in 2017.

That is a change of `r (predict(hursat_global_results$major_logit_fit, type="response")[num_years_hursat]-predict(hursat_global_results$major_logit_fit, type="response")[1])/num_years_hursat*10*100` percent per decade.


#### Logistic Regression on Proportions of Intense

```{r, echo=FALSE}
summary(hursat_global_results$intense_logit_fit)
```


Specifically, using this model the probability of a major storm in 1981 was `r predict(hursat_global_results$intense_logit_fit, type="response")[1]` and `r predict(hursat_global_results$intense_logit_fit, type="response")[num_years_hursat]` in 2017.

That is a change of `r (predict(hursat_global_results$intense_logit_fit, type="response")[num_years_hursat]-predict(hursat_global_results$intense_logit_fit, type="response")[1])/num_years_hursat*10*100` percent per decade.


### Change Point Methods

#### Change in Non-Shorty Counts

```{r, echo=FALSE}
hursat_global_results$pois_chpt_test
```

The dates suggest the changepoint would occur from `r 1980+hursat_global_results$pois_chpt_test[2]` to `r 1980+hursat_global_results$pois_chpt_test[2]+1` with $p$-value `r hursat_global_results$pois_chpt_test[3]`.

In regime one, the mean number of total storms is `r mean(hursat_global_results$data$Total_Storms[1:hursat_global_results$pois_chpt_test[2]])` and in regime two, the mean number of storms is `r mean(hursat_global_results$data$Total_Storms[(hursat_global_results$pois_chpt_test[2]+1):num_years_hursat])`; a change of `r mean(hursat_global_results$data$Total_Storms[(hursat_global_results$pois_chpt_test[2]+1):num_years_hursat]) - mean(hursat_global_results$data$Total_Storms[1:hursat_global_results$pois_chpt_test[2]])`.

#### Change in counts of major storms

```{r, echo=FALSE}
hursat_global_results$major_pois_chpt_test
```

The dates suggest the changepoint would occur from `r 1980+hursat_global_results$major_pois_chpt_test[2]` to `r 1980+hursat_global_results$major_pois_chpt_test[2]+1` with $p$-value `r hursat_global_results$major_pois_chpt_test[3]`.

In regime one, the mean number of total storms is `r mean(hursat_global_results$data$Major_Storms[1:hursat_global_results$major_pois_chpt_test[2]])` and in regime two, the mean number of storms is `r mean(hursat_global_results$data$Major_Storms[(hursat_global_results$major_pois_chpt_test[2]+1):num_years_hursat])`; a change of `r mean(hursat_global_results$data$Major_Storms[(hursat_global_results$major_pois_chpt_test[2]+1):num_years_hursat]) - mean(hursat_global_results$data$Major_Storms[1:hursat_global_results$major_pois_chpt_test[2]])`.

#### Change in counts of intense storms

```{r, echo=FALSE}
hursat_global_results$intense_pois_chpt_test
```

The dates suggest the changepoint would occur from `r 1980+hursat_global_results$intense_pois_chpt_test[2]` to `r 1980+hursat_global_results$intense_pois_chpt_test[2]+1` with $p$-value `r hursat_global_results$intense_pois_chpt_test[3]`.

In regime one, the mean number of total storms is `r mean(hursat_global_results$data$Intense_Storms[1:hursat_global_results$intense_pois_chpt_test[2]])` and in regime two, the mean number of storms is `r mean(hursat_global_results$data$Intense_Storms[(hursat_global_results$intense_pois_chpt_test[2]+1):num_years_hursat])`; a change of `r mean(hursat_global_results$data$Intense_Storms[(hursat_global_results$intense_pois_chpt_test[2]+1):num_years_hursat]) - mean(hursat_global_results$data$Intense_Storms[1:hursat_global_results$intense_pois_chpt_test[2]])`.


#### Proportion of major storms

A change in the proportion of major storms?
  
```{r, echo=FALSE}
hursat_global_results$prop_major_chpt_test
```

The change in proportions occurs `r 1979+hursat_global_results$prop_major_chpt_test[2]` to `r 1979+hursat_global_results$prop_major_chpt_test[2]+1` with $p$-value `r hursat_global_results$prop_major_chpt_test[3]`. 

In regime one, the proportion of major storms is `r sum(hursat_global_results$data$Major_Storms[1:hursat_global_results$prop_major_chpt_test[2]])/sum(hursat_global_results$data$Total_Storms[1:hursat_global_results$prop_major_chpt_test[2]])` and in regime two, the proportion of major storms is `r sum(hursat_global_results$data$Major_Storms[(hursat_global_results$prop_major_chpt_test[2]+1):num_years_hursat])/sum(hursat_global_results$data$Total_Storms[(hursat_global_results$prop_major_chpt_test[2]+1):num_years_hursat])`; a change of `r sum(hursat_global_results$data$Major_Storms[(hursat_global_results$prop_major_chpt_test[2]+1):num_years_hursat])/sum(hursat_global_results$data$Total_Storms[(hursat_global_results$prop_major_chpt_test[2]+1):num_years_hursat]) - sum(hursat_global_results$data$Major_Storms[1:hursat_global_results$prop_major_chpt_test[2]])/sum(hursat_global_results$data$Total_Storms[1:hursat_global_results$prop_major_chpt_test[2]])`.



#### Proportion of intense storms

A change in the proportion of major storms?
  
```{r, echo=FALSE}
hursat_global_results$prop_intense_chpt_test
```

The change in proportions occurs `r 1979+hursat_global_results$prop_intense_chpt_test[2]` to `r 1979+hursat_global_results$prop_intense_chpt_test[2]+1` with $p$-value `r hursat_global_results$prop_intense_chpt_test[3]`.


In regime one, the proportion of major storms is `r sum(hursat_global_results$data$Intense_Storms[1:hursat_global_results$prop_intense_chpt_test[2]])/sum(hursat_global_results$data$Total_Storms[1:hursat_global_results$prop_intense_chpt_test[2]])` and in regime two, the proportion of major storms is `r sum(hursat_global_results$data$Intense_Storms[(hursat_global_results$prop_intense_chpt_test[2]+1):num_years_hursat])/sum(hursat_global_results$data$Total_Storms[(hursat_global_results$prop_intense_chpt_test[2]+1):num_years_hursat])`; a change of `r sum(hursat_global_results$data$Intense_Storms[(hursat_global_results$prop_intense_chpt_test[2]+1):num_years_hursat])/sum(hursat_global_results$data$Total_Storms[(hursat_global_results$prop_intense_chpt_test[2]+1):num_years_hursat]) - sum(hursat_global_results$data$Intense_Storms[1:hursat_global_results$prop_intense_chpt_test[2]])/sum(hursat_global_results$data$Total_Storms[1:hursat_global_results$prop_intense_chpt_test[2]])`.


### Overall Change Point Findings

Basically, the total number of storms is decreasing a bit while the number of major storms is more-or-less constant, but not seeing enough change to suggest a change in proportions.

A very suggestive result of an overall change to the number of non-shorties. So for exploration look at how counts changed in the potential change point time.

```{r, echo=FALSE}
bind_rows(
  hursat_all_basins |>
    dplyr::filter(BASIN=="Global") |>
    summarize(`Total Storms (avg)` = mean(Total_Storms),
              `Total Storms (stddev)` = sd(Total_Storms) ) |>
    mutate(Regime = "Entire Record"),
  hursat_all_basins |>
    dplyr::filter(BASIN=="Global") |>
    mutate(Regime = ifelse(SEASON < 2001, "1981-2000", "2001-2017") ) |>
    group_by(Regime) |>
    summarize(`Total Storms (avg)` = mean(Total_Storms),
              `Total Storms (stddev)` = sd(Total_Storms) ) 
)|>
  dplyr::select(Regime, `Total Storms (avg)`, `Total Storms (stddev)`) |>
  knitr::kable()
```

We do see a decrease of nearly 9 storms per year.

A table to compare the proportion of major storms and note the change point location was a year later albeit not-significant. We also include the intense storms here.

```{r, echo=FALSE}
bind_rows(
  hursat_all_basins |>
    dplyr::filter(BASIN=="Global") |>
    summarize(`Total Storms (avg)` = mean(Total_Storms),
              `Major Storms (avg)` = mean(Major_Storms),
              `Proportion Major` = sum(Major_Storms)/sum(Total_Storms),
              `Intense Storms (avg)` = mean(Intense_Storms),
              `Proportion Intense` = sum(Intense_Storms)/sum(Total_Storms) ) |>
    mutate(Regime = "Entire Record"),
  hursat_all_basins |>
    dplyr::filter(BASIN=="Global") |>
    mutate(Regime = ifelse(SEASON < 2002, "1981-2001", "2002-2017") ) |>
    group_by(Regime) |>
    summarize(`Total Storms (avg)` = mean(Total_Storms),
              `Major Storms (avg)` = mean(Major_Storms),
              `Proportion Major` = sum(Major_Storms)/sum(Total_Storms),
              `Intense Storms (avg)` = mean(Intense_Storms),
              `Proportion Intense` = sum(Intense_Storms)/sum(Total_Storms) ) ) |>
  dplyr::select(Regime, `Total Storms (avg)`, `Major Storms (avg)`, `Proportion Major`, `Intense Storms (avg)`, `Proportion Intense`) |>
  knitr::kable()
```

About 9 less total storms per year, with the number of major storms staying constant, resulting in about a 3% increase in major storms.

```{r, echo=FALSE}
hursat_global_results$plot_seg_fits
```


## Trimmed IBTRacs

We now repeat the analysis with the a version of the IBTRacs that has been trimmed to match HURSAT record. So this data runs from 1981 through 2017.

```{r, echo=FALSE}
trimmed_global_results <- get_report_basin(basin="Global", all_basins=ibtracs_trimmed)
```


```{r, echo=FALSE}
trimmed_global_results$plot_data
```


### Summary Statistics (overall)

```{r, echo=FALSE}
trimmed_global_results$summary_stat |>
  knitr::kable()
```

### Regression Methods

#### Poisson Regression

```{r, echo=FALSE}
summary(trimmed_global_results$pois_fit)
```

Note, this says a decrease of `r abs(coef(trimmed_global_results$pois_fit)[2])` log-storms per year. This corresponds to a `r (1-exp(coef(trimmed_global_results$pois_fit)[2]))*100` percent decrease per season. So over `r num_years_hursat` years, that is a decrease of `r (1-exp(coef(trimmed_global_results$pois_fit)[2])^num_years_hursat)*100` percent.

Using this model, we would predict `r predict(trimmed_global_results$pois_fit, type="response")[1]` storms in 1981 and `r predict(trimmed_global_results$pois_fit, type="response")[num_years_hursat]` storms in 2017

That is a change of `r (predict(trimmed_global_results$pois_fit, type="response")[num_years_hursat]-predict(trimmed_global_results$pois_fit, type="response")[1])/num_years_hursat*10` storms per decade.

#### Poisson Regression on Major Storms

```{r, echo=FALSE}
summary(trimmed_global_results$major_fit)
```

We see a non-significant trend in the number of major storms.

Using this model the number of a major storm in 1981 was `r predict(trimmed_global_results$major_fit, type="response")[1]` and `r predict(trimmed_global_results$major_fit, type="response")[num_years_hursat]` in 2017.

That is a change of `r (predict(trimmed_global_results$major_fit, type="response")[num_years_hursat]-predict(trimmed_global_results$major_fit, type="response")[1])/num_years_hursat*10` storms per decade.


#### Poisson Regression on Intense Storms

```{r, echo=FALSE}
summary(trimmed_global_results$intense_fit)
```

Using this model the number of a major storm in 1981 was `r predict(trimmed_global_results$intense_fit, type="response")[1]` and `r predict(trimmed_global_results$intense_fit, type="response")[num_years_hursat]` in 2017

That is a change of `r (predict(trimmed_global_results$intense_fit, type="response")[num_years_hursat]-predict(trimmed_global_results$intense_fit, type="response")[1])/num_years_hursat*10` storms per decade.


#### Logistic Regression on Proportions

```{r, echo=FALSE}
summary(trimmed_global_results$major_logit_fit)
```


Specifically, using this model the probability of a major storm in 1981 was `r predict(trimmed_global_results$major_logit_fit, type="response")[1]` and `r predict(trimmed_global_results$major_logit_fit, type="response")[num_years_hursat]` in 2017.


That is a change of `r (predict(trimmed_global_results$major_logit_fit, type="response")[num_years_hursat]-predict(trimmed_global_results$major_logit_fit, type="response")[1])/num_years_hursat*10*100` percent per decade.



#### Logistic Regression on Proportions of Intense

```{r, echo=FALSE}
summary(trimmed_global_results$intense_logit_fit)
```


Specifically, using this model the probability of a major storm in 1981 was `r predict(trimmed_global_results$intense_logit_fit, type="response")[1]` and `r predict(trimmed_global_results$intense_logit_fit, type="response")[num_years_hursat]` in 2017.

That is a change of `r (predict(trimmed_global_results$intense_logit_fit, type="response")[num_years_hursat]-predict(trimmed_global_results$intense_logit_fit, type="response")[1])/num_years_hursat*10*100` percent per decade.


### Change Point Methods

#### Change in Non-Shorty Counts

```{r, echo=FALSE}
trimmed_global_results$pois_chpt_test
```

We have evidence the number of non-shorties has changed. The dates suggest the changepoint would occur from `r 1980+trimmed_global_results$pois_chpt_test[2]` to `r 1980+trimmed_global_results$pois_chpt_test[2]+1` with $p$-value `r hursat_global_results$pois_chpt_test[3]`.

In regime one, the mean number of total storms is `r mean(trimmed_global_results$data$Total_Storms[1:trimmed_global_results$pois_chpt_test[2]])` and in regime two, the mean number of storms is `r mean(trimmed_global_results$data$Total_Storms[(trimmed_global_results$pois_chpt_test[2]+1):num_years_hursat])`; a change of `r mean(trimmed_global_results$data$Total_Storms[(trimmed_global_results$pois_chpt_test[2]+1):num_years_hursat]) - mean(trimmed_global_results$data$Total_Storms[1:trimmed_global_results$pois_chpt_test[2]])`.

#### Change in counts of major storms

```{r, echo=FALSE}
trimmed_global_results$major_pois_chpt_test
```


The dates suggest the changepoint would occur from `r 1980+trimmed_global_results$major_pois_chpt_test[2]` to `r 1980+trimmed_global_results$major_pois_chpt_test[2]+1` with $p$-value `r trimmed_global_results$major_pois_chpt_test[3]`.

In regime one, the mean number of total storms is `r mean(trimmed_global_results$data$Major_Storms[1:trimmed_global_results$major_pois_chpt_test[2]])` and in regime two, the mean number of storms is `r mean(trimmed_global_results$data$Major_Storms[(trimmed_global_results$major_pois_chpt_test[2]+1):num_years_hursat])`; a change of `r mean(trimmed_global_results$data$Major_Storms[(trimmed_global_results$major_pois_chpt_test[2]+1):num_years_hursat]) - mean(trimmed_global_results$data$Major_Storms[1:trimmed_global_results$major_pois_chpt_test[2]])`.


#### Change in counts of intense storms

```{r, echo=FALSE}
trimmed_global_results$intense_pois_chpt_test
```


The dates suggest the changepoint would occur from `r 1980+trimmed_global_results$intense_pois_chpt_test[2]` to `r 1980+trimmed_global_results$intense_pois_chpt_test[2]+1` with $p$-value `r trimmed_global_results$intense_pois_chpt_test[3]`.

In regime one, the mean number of total storms is `r mean(trimmed_global_results$data$Intense_Storms[1:trimmed_global_results$intense_pois_chpt_test[2]])` and in regime two, the mean number of storms is `r mean(trimmed_global_results$data$Intense_Storms[(trimmed_global_results$intense_pois_chpt_test[2]+1):num_years_hursat])`; a change of `r mean(trimmed_global_results$data$Intense_Storms[(trimmed_global_results$intense_pois_chpt_test[2]+1):num_years_hursat]) - mean(trimmed_global_results$data$Intense_Storms[1:trimmed_global_results$intense_pois_chpt_test[2]])`.




#### Proportion of major storms

A change in the proportion of major storms?
  
```{r, echo=FALSE}
trimmed_global_results$prop_major_chpt_test
```


The change in proportions occurs `r 1979+trimmed_global_results$prop_major_chpt_test[2]` to `r 1979+trimmed_global_results$prop_major_chpt_test[2]+1` with $p$-value `r trimmed_global_results$prop_major_chpt_test[3]`. 

In regime one, the proportion of major storms is `r sum(trimmed_global_results$data$Major_Storms[1:trimmed_global_results$prop_major_chpt_test[2]])/sum(trimmed_global_results$data$Total_Storms[1:trimmed_global_results$prop_major_chpt_test[2]])` and in regime two, the proportion of major storms is `r sum(trimmed_global_results$data$Major_Storms[(trimmed_global_results$prop_major_chpt_test[2]+1):num_years_hursat])/sum(trimmed_global_results$data$Total_Storms[(trimmed_global_results$prop_major_chpt_test[2]+1):num_years_hursat])`; a change of `r sum(trimmed_global_results$data$Major_Storms[(trimmed_global_results$prop_major_chpt_test[2]+1):num_years_hursat])/sum(trimmed_global_results$data$Total_Storms[(trimmed_global_results$prop_major_chpt_test[2]+1):num_years_hursat]) - sum(trimmed_global_results$data$Major_Storms[1:trimmed_global_results$prop_major_chpt_test[2]])/sum(trimmed_global_results$data$Total_Storms[1:trimmed_global_results$prop_major_chpt_test[2]])`.


#### Proportion of intense storms

A change in the proportion of major storms?
  
```{r, echo=FALSE}
trimmed_global_results$prop_intense_chpt_test
```

The change in proportions occurs `r 1979+trimmed_global_results$prop_intense_chpt_test[2]` to `r 1979+trimmed_global_results$prop_intense_chpt_test[2]+1` with $p$-value `r trimmed_global_results$prop_intense_chpt_test[3]`.


In regime one, the proportion of major storms is `r sum(trimmed_global_results$data$Intense_Storms[1:trimmed_global_results$prop_intense_chpt_test[2]])/sum(trimmed_global_results$data$Total_Storms[1:trimmed_global_results$prop_intense_chpt_test[2]])` and in regime two, the proportion of major storms is `r sum(trimmed_global_results$data$Intense_Storms[(trimmed_global_results$prop_intense_chpt_test[2]+1):num_years_hursat])/sum(trimmed_global_results$data$Total_Storms[(trimmed_global_results$prop_intense_chpt_test[2]+1):num_years_hursat])`; a change of `r sum(trimmed_global_results$data$Intense_Storms[(trimmed_global_results$prop_intense_chpt_test[2]+1):num_years_hursat])/sum(trimmed_global_results$data$Total_Storms[(trimmed_global_results$prop_intense_chpt_test[2]+1):num_years_hursat]) - sum(trimmed_global_results$data$Intense_Storms[1:trimmed_global_results$prop_intense_chpt_test[2]])/sum(trimmed_global_results$data$Total_Storms[1:trimmed_global_results$prop_intense_chpt_test[2]])`.



### Overall Change Point Findings

Similar results to the HURSAT and full IBTracs, the total number of storms is decreasing a bit while the number of major storms is more-or-less constant, so therefore the proportion of major storms has increased. The number of intense storms has changed, as well as the proportion of intense storms.

A very suggestive result of an overall change to the number of non-shorties. So for exploration look at how counts changed in the potential change point time.

```{r, echo=FALSE}
bind_rows(
  ibtracs_trimmed |>
    dplyr::filter(BASIN=="Global") |>
    summarize(`Total Storms (avg)` = mean(Total_Storms),
              `Total Storms (stddev)` = sd(Total_Storms) ) |>
    mutate(Regime = "Entire Record"),
  ibtracs_trimmed |>
    dplyr::filter(BASIN=="Global") |>
    mutate(Regime = ifelse(SEASON < 1998, "1981-1997", "1998-2017") ) |>
    group_by(Regime) |>
    summarize(`Total Storms (avg)` = mean(Total_Storms),
              `Total Storms (stddev)` = sd(Total_Storms) ) 
)|>
  dplyr::select(Regime, `Total Storms (avg)`, `Total Storms (stddev)`) |>
  knitr::kable()
```

We do see a decrease of about 9 storms per year.

A table to compare the proportion of major storms. Note the change point method suggested this shift occurred earlier.

```{r, echo=FALSE}
bind_rows(
  ibtracs_trimmed |>
    dplyr::filter(BASIN=="Global") |>
    summarize(`Total Storms (avg)` = mean(Total_Storms),
              `Major Storms (avg)` = mean(Major_Storms),
              `Proportion Major` = sum(Major_Storms)/sum(Total_Storms)) |>
    mutate(Regime = "Entire Record"),
  ibtracs_trimmed |>
    dplyr::filter(BASIN=="Global") |>
    mutate(Regime = ifelse(SEASON < 1991, "1981-1990", "1991-2017") ) |>
    group_by(Regime) |>
    summarize(`Total Storms (avg)` = mean(Total_Storms),
              `Major Storms (avg)` = mean(Major_Storms),
              `Proportion Major` = sum(Major_Storms)/sum(Total_Storms)) ) |>
  dplyr::select(Regime, `Total Storms (avg)`, `Major Storms (avg)`, `Proportion Major`) |>
  knitr::kable()
```

Proportion of major storms increased by almost 7%. The numbers are fairly convincing when looking at these two regimes. The total number of storms dropped by about 6 but the number of major storms actually increased; so proportionally we see a clear jump.

The results for intense storms is just a year later

```{r, echo=FALSE}
bind_rows(
  ibtracs_trimmed |>
    dplyr::filter(BASIN=="Global") |>
    summarize(`Total Storms (avg)` = mean(Total_Storms),
              `Intense Storms (avg)` = mean(Intense_Storms),
              `Proportion Intense` = sum(Intense_Storms)/sum(Total_Storms)) |>
    mutate(Regime = "Entire Record"),
  ibtracs_trimmed |>
    dplyr::filter(BASIN=="Global") |>
    mutate(Regime = ifelse(SEASON < 1992, "1981-1991", "1992-2017") ) |>
    group_by(Regime) |>
    summarize(`Total Storms (avg)` = mean(Total_Storms),
              `Intense Storms (avg)` = mean(Intense_Storms),
              `Proportion Intense` = sum(Intense_Storms)/sum(Total_Storms)) ) |>
  dplyr::select(Regime, `Total Storms (avg)`, `Intense Storms (avg)`, `Proportion Intense`) |>
  knitr::kable()
```


```{r, echo=FALSE}
trimmed_global_results$plot_seg_fits
```


## Overall Hypothesis Test Results

### Total Number of storms (non-shorties)
  
**Poisson Regression**

```{r, echo=FALSE}
data.frame(Data=c("IBTracs 1980-2025", "HURSAT 1981-2017", "IBTracs 1981-2017"),
           Coefficent = c(coef(ibtracs_global_results$pois_fit)[2],coef(hursat_global_results$pois_fit)[2],coef(trimmed_global_results$pois_fit)[2]),
           `p-value` = c(anova(ibtracs_global_results$pois_fit)$`Pr(>Chi)`[2], anova(hursat_global_results$pois_fit)$`Pr(>Chi)`[2], anova(trimmed_global_results$pois_fit)$`Pr(>Chi)`[2]),
           `Start` = c(predict(ibtracs_global_results$pois_fit, type="response")[1], predict(hursat_global_results$pois_fit, type="response")[1], predict(trimmed_global_results$pois_fit, type="response")[1] ),
           `End` = c(predict(ibtracs_global_results$pois_fit, type="response")[length(predict(ibtracs_global_results$pois_fit, type="response"))], predict(hursat_global_results$pois_fit, type="response")[length(predict(hursat_global_results$pois_fit, type="response"))], predict(trimmed_global_results$pois_fit, type="response")[length(predict(trimmed_global_results$pois_fit, type="response"))] ),
           `Change per Decade` = c( (predict(ibtracs_global_results$pois_fit, type="response")[length(predict(ibtracs_global_results$pois_fit, type="response"))]-predict(ibtracs_global_results$pois_fit, type="response")[1])/length(predict(ibtracs_global_results$pois_fit, type="response"))*10,
                                    (predict(hursat_global_results$pois_fit, type="response")[length(predict(hursat_global_results$pois_fit, type="response"))]-predict(hursat_global_results$pois_fit, type="response")[1])/length(predict(hursat_global_results$pois_fit, type="response"))*10,
                                    (predict(trimmed_global_results$pois_fit, type="response")[length(predict(trimmed_global_results$pois_fit, type="response"))]-predict(trimmed_global_results$pois_fit, type="response")[1])/length(predict(trimmed_global_results$pois_fit, type="response"))*10 ) ) |>
  knitr::kable(digits=5)
```

**SCUSUM Change Point Test**

```{r, echo=FALSE}
data.frame(Data=c("IBTracs 1980-2025", "HURSAT 1981-2017", "IBTracs 1981-2017"),
           `Chpt Location` = c(ibtracs_global_results$pois_chpt_test[2]+1980,hursat_global_results$pois_chpt_test[2]+1981,trimmed_global_results$pois_chpt_test[2]+1981),
           `p-value` =c(ibtracs_global_results$pois_chpt_test[3],hursat_global_results$pois_chpt_test[3],trimmed_global_results$pois_chpt_test[3]),
           `Regime 1` = c(mean(ibtracs_global_results$data$Total_Storms[1:ibtracs_global_results$pois_chpt_test[2]]),
                          mean(hursat_global_results$data$Total_Storms[1:hursat_global_results$pois_chpt_test[2]]),
                          mean(trimmed_global_results$data$Total_Storms[1:trimmed_global_results$pois_chpt_test[2]]) ),
           `Regime 2` = c(mean(ibtracs_global_results$data$Total_Storms[(ibtracs_global_results$pois_chpt_test[2]+1):length(ibtracs_global_results$data$Total_Storms)]),
                          mean(hursat_global_results$data$Total_Storms[(hursat_global_results$pois_chpt_test[2]+1):length(hursat_global_results$data$Total_Storms)]),
                            mean(trimmed_global_results$data$Total_Storms[(trimmed_global_results$pois_chpt_test[2]+1):length(trimmed_global_results$data$Total_Storms)]) ),
           `Change` = c(mean(ibtracs_global_results$data$Total_Storms[(ibtracs_global_results$pois_chpt_test[2]+1):length(ibtracs_global_results$data$Total_Storms)]) - mean(ibtracs_global_results$data$Total_Storms[1:ibtracs_global_results$pois_chpt_test[2]]),
                        mean(hursat_global_results$data$Total_Storms[(hursat_global_results$pois_chpt_test[2]+1):length(hursat_global_results$data$Total_Storms)]) - mean(hursat_global_results$data$Total_Storms[1:hursat_global_results$pois_chpt_test[2]]),
                        mean(trimmed_global_results$data$Total_Storms[(trimmed_global_results$pois_chpt_test[2]+1):length(trimmed_global_results$data$Total_Storms)]) - mean(trimmed_global_results$data$Total_Storms[1:trimmed_global_results$pois_chpt_test[2]]) ) ) |>
  mutate(Perc_Change = Change/`Regime.1`*100) |>
  knitr::kable(digits=5)
```

### Major Storm Counts

**Poisson Regression**

```{r, echo=FALSE}
data.frame(Data=c("IBTracs 1980-2025", "HURSAT 1981-2017", "IBTracs 1981-2017"),
           Coefficent = c(coef(ibtracs_global_results$major_fit)[2],
                          coef(hursat_global_results$major_fit)[2],
                          coef(trimmed_global_results$major_fit)[2]),
           `p-value` = c(anova(ibtracs_global_results$major_fit)$`Pr(>Chi)`[2],
                         anova(hursat_global_results$major_fit)$`Pr(>Chi)`[2],
                         anova(trimmed_global_results$major_fit)$`Pr(>Chi)`[2]),
           `Start` = c(predict(ibtracs_global_results$major_fit, type="response")[1], 
                       predict(hursat_global_results$major_fit, type="response")[1], 
                       predict(trimmed_global_results$major_fit, type="response")[1] ),
           `End` = c(predict(ibtracs_global_results$major_fit,type="response")[length(predict(ibtracs_global_results$major_fit, type="response"))],
                     predict(hursat_global_results$major_fit, type="response")[length(predict(hursat_global_results$major_fit, type="response"))], 
                     predict(trimmed_global_results$major_fit, type="response")[length(predict(trimmed_global_results$major_fit, type="response"))] ),
           `Change per Decade` = c( (predict(ibtracs_global_results$major_fit, type="response")[length(predict(ibtracs_global_results$major_fit, type="response"))]-predict(ibtracs_global_results$major_fit, type="response")[1])/length(predict(ibtracs_global_results$major_fit, type="response"))*10,
                                    (predict(hursat_global_results$major_fit, type="response")[length(predict(hursat_global_results$major_fit, type="response"))]-predict(hursat_global_results$major_fit, type="response")[1])/length(predict(hursat_global_results$major_fit, type="response"))*10,
                                    (predict(trimmed_global_results$major_fit, type="response")[length(predict(trimmed_global_results$major_fit, type="response"))]-predict(trimmed_global_results$major_fit, type="response")[1])/length(predict(trimmed_global_results$major_fit, type="response"))*10 ) ) |>
  knitr::kable(digits=5)
```

**SCUSUM Change Point Test**

```{r, echo=FALSE}
data.frame(Data=c("IBTracs 1980-2025", "HURSAT 1981-2017", "IBTracs 1981-2017"),
           `Chpt Location` = c(ibtracs_global_results$major_pois_chpt_test[2]+1980,hursat_global_results$major_pois_chpt_test[2]+1981,trimmed_global_results$major_pois_chpt_test[2]+1981),
           `p-value` =c(ibtracs_global_results$major_pois_chpt_test[3],hursat_global_results$major_pois_chpt_test[3],trimmed_global_results$major_pois_chpt_test[3]),
           `Regime 1` = c(mean(ibtracs_global_results$data$Major_Storms[1:ibtracs_global_results$major_pois_chpt_test[2]]),
                          mean(hursat_global_results$data$Major_Storms[1:hursat_global_results$major_pois_chpt_test[2]]),
                          mean(trimmed_global_results$data$Major_Storms[1:trimmed_global_results$major_pois_chpt_test[2]]) ),
           `Regime 2` = c(mean(ibtracs_global_results$data$Major_Storms[(ibtracs_global_results$major_pois_chpt_test[2]+1):length(ibtracs_global_results$data$Major_Storms)]),
                          mean(hursat_global_results$data$Major_Storms[(hursat_global_results$major_pois_chpt_test[2]+1):length(hursat_global_results$data$Major_Storms)]),
                            mean(trimmed_global_results$data$Major_Storms[(trimmed_global_results$major_pois_chpt_test[2]+1):length(trimmed_global_results$data$Major_Storms)]) ),
           `Change` = c(mean(ibtracs_global_results$data$Major_Storms[(ibtracs_global_results$major_pois_chpt_test[2]+1):length(ibtracs_global_results$data$Major_Storms)]) - mean(ibtracs_global_results$data$Major_Storms[1:ibtracs_global_results$major_pois_chpt_test[2]]),
                        mean(hursat_global_results$data$Major_Storms[(hursat_global_results$major_pois_chpt_test[2]+1):length(hursat_global_results$data$Major_Storms)]) - mean(hursat_global_results$data$Major_Storms[1:hursat_global_results$major_pois_chpt_test[2]]),
                        mean(trimmed_global_results$data$Major_Storms[(trimmed_global_results$major_pois_chpt_test[2]+1):length(trimmed_global_results$data$Major_Storms)]) - mean(trimmed_global_results$data$Major_Storms[1:trimmed_global_results$major_pois_chpt_test[2]]) ) ) |>
  mutate(Perc_Change = Change/`Regime.1`*100) |>
  knitr::kable(digits=5)
```

### Proportion of Major Storms

**Logistic Regression**

```{r, echo=FALSE}
data.frame(Data=c("IBTracs 1980-2025", "HURSAT 1981-2017", "IBTracs 1981-2017"),
           Coefficent = c(coef(ibtracs_global_results$major_logit_fit)[2],coef(hursat_global_results$major_logit_fit)[2],coef(trimmed_global_results$major_logit_fit)[2]),
           `p-value` = c(anova(ibtracs_global_results$major_logit_fit)$`Pr(>Chi)`[2], anova(hursat_global_results$major_logit_fit)$`Pr(>Chi)`[2], anova(trimmed_global_results$major_logit_fit)$`Pr(>Chi)`[2]),
                      `Start` = c(predict(ibtracs_global_results$major_logit_fit, type="response")[1]*100, 
                       predict(hursat_global_results$major_logit_fit, type="response")[1]*100, 
                       predict(trimmed_global_results$major_logit_fit, type="response")[1]*100 ),
           `End` = c(predict(ibtracs_global_results$major_logit_fit,type="response")[length(predict(ibtracs_global_results$major_fit, type="response"))]*100,
                     predict(hursat_global_results$major_logit_fit, type="response")[length(predict(hursat_global_results$major_logit_fit, type="response"))]*100, 
                     predict(trimmed_global_results$major_logit_fit, type="response")[length(predict(trimmed_global_results$major_logit_fit, type="response"))]*100 ),
           `Change per Decade` = c( (predict(ibtracs_global_results$major_logit_fit, type="response")[length(predict(ibtracs_global_results$major_logit_fit, type="response"))]-predict(ibtracs_global_results$major_logit_fit, type="response")[1])/length(predict(ibtracs_global_results$major_logit_fit, type="response"))*10*100,
                                    (predict(hursat_global_results$major_logit_fit, type="response")[length(predict(hursat_global_results$major_logit_fit, type="response"))]-predict(hursat_global_results$major_logit_fit, type="response")[1])/length(predict(hursat_global_results$major_logit_fit, type="response"))*10*100,
                                    (predict(trimmed_global_results$major_logit_fit, type="response")[length(predict(trimmed_global_results$major_logit_fit, type="response"))]-predict(trimmed_global_results$major_logit_fit, type="response")[1])/length(predict(trimmed_global_results$major_logit_fit, type="response"))*10*100 ) ) |>
  knitr::kable()
```

**SCUSUM Change Point Test**

```{r, echo=FALSE}
data.frame(Data=c("IBTracs 1980-2025", "HURSAT 1981-2017", "IBTracs 1981-2017"),
           `Chpt Location` = c(ibtracs_global_results$prop_major_chpt_test[2]+1980,hursat_global_results$prop_major_chpt_test[2]+1981,trimmed_global_results$prop_major_chpt_test[2]+1981),
           `p-value` =c(ibtracs_global_results$prop_major_chpt_test[3],hursat_global_results$prop_major_chpt_test[3],trimmed_global_results$prop_major_chpt_test[3]),
           `Regime 1 %` = c(sum(ibtracs_global_results$data$Major_Storms[1:ibtracs_global_results$prop_major_chpt_test[2]])/sum(ibtracs_global_results$data$Total_Storms[1:ibtracs_global_results$prop_major_chpt_test[2]])*100,
                            sum(hursat_global_results$data$Major_Storms[1:hursat_global_results$prop_major_chpt_test[2]])/sum(hursat_global_results$data$Total_Storms[1:hursat_global_results$prop_major_chpt_test[2]])*100,
                            sum(ibtracs_global_results$data$Major_Storms[1:ibtracs_global_results$prop_major_chpt_test[2]])/sum(ibtracs_global_results$data$Total_Storms[1:ibtracs_global_results$prop_major_chpt_test[2]])*100 ),
           `Regime 2 %` = c(sum(ibtracs_global_results$data$Major_Storms[(ibtracs_global_results$prop_major_chpt_test[2]+1):length(ibtracs_global_results$data$Major_Storms)])/sum(ibtracs_global_results$data$Total_Storms[(ibtracs_global_results$prop_major_chpt_test[2]+1):length(ibtracs_global_results$data$Major_Storms)])*100,
                            sum(hursat_global_results$data$Major_Storms[(hursat_global_results$prop_major_chpt_test[2]+1):length(hursat_global_results$data$Major_Storms)])/sum(hursat_global_results$data$Total_Storms[(hursat_global_results$prop_major_chpt_test[2]+1):length(hursat_global_results$data$Major_Storms)])*100,
                            sum(trimmed_global_results$data$Major_Storms[(trimmed_global_results$prop_major_chpt_test[2]+1):length(trimmed_global_results$data$Major_Storms)])/sum(trimmed_global_results$data$Total_Storms[(trimmed_global_results$prop_major_chpt_test[2]+1):length(trimmed_global_results$data$Major_Storms)])*100 ),
           `Change %` = c((sum(ibtracs_global_results$data$Major_Storms[(ibtracs_global_results$prop_major_chpt_test[2]+1):length(ibtracs_global_results$data$Major_Storms)])/sum(ibtracs_global_results$data$Total_Storms[(ibtracs_global_results$prop_major_chpt_test[2]+1):length(ibtracs_global_results$data$Major_Storms)]) - sum(ibtracs_global_results$data$Major_Storms[1:ibtracs_global_results$prop_major_chpt_test[2]])/sum(ibtracs_global_results$data$Total_Storms[1:ibtracs_global_results$prop_major_chpt_test[2]]) )*100,
                          (sum(hursat_global_results$data$Major_Storms[(hursat_global_results$prop_major_chpt_test[2]+1):length(hursat_global_results$data$Major_Storms)])/sum(hursat_global_results$data$Total_Storms[(hursat_global_results$prop_major_chpt_test[2]+1):length(hursat_global_results$data$Major_Storms)]) - sum(hursat_global_results$data$Major_Storms[1:hursat_global_results$prop_major_chpt_test[2]])/sum(hursat_global_results$data$Total_Storms[1:hursat_global_results$prop_major_chpt_test[2]]) )*100,
                          (sum(trimmed_global_results$data$Major_Storms[(trimmed_global_results$prop_major_chpt_test[2]+1):length(trimmed_global_results$data$Major_Storms)])/sum(trimmed_global_results$data$Total_Storms[(trimmed_global_results$prop_major_chpt_test[2]+1):length(trimmed_global_results$data$Major_Storms)]) - sum(trimmed_global_results$data$Major_Storms[1:trimmed_global_results$prop_major_chpt_test[2]])/sum(trimmed_global_results$data$Total_Storms[1:trimmed_global_results$prop_major_chpt_test[2]]) )*100 ) ) |>
  knitr::kable()
```



### Intense Storm Counts

**Poisson Regression**

```{r, echo=FALSE}
data.frame(Data=c("IBTracs 1980-2025", "HURSAT 1981-2017", "IBTracs 1981-2017"),
           Coefficent = c(coef(ibtracs_global_results$intense_fit)[2],
                          coef(hursat_global_results$intense_fit)[2],
                          coef(trimmed_global_results$intense_fit)[2]),
           `p-value` = c(anova(ibtracs_global_results$intense_fit)$`Pr(>Chi)`[2],
                         anova(hursat_global_results$intense_fit)$`Pr(>Chi)`[2],
                         anova(trimmed_global_results$intense_fit)$`Pr(>Chi)`[2]),
           `Start` = c(predict(ibtracs_global_results$intense_fit, type="response")[1], 
                       predict(hursat_global_results$intense_fit, type="response")[1], 
                       predict(trimmed_global_results$intense_fit, type="response")[1] ),
           `End` = c(predict(ibtracs_global_results$intense_fit,type="response")[length(predict(ibtracs_global_results$intense_fit, type="response"))],
                     predict(hursat_global_results$intense_fit, type="response")[length(predict(hursat_global_results$intense_fit, type="response"))], 
                     predict(trimmed_global_results$intense_fit, type="response")[length(predict(trimmed_global_results$intense_fit, type="response"))] ),
           `Change per Decade` = c( (predict(ibtracs_global_results$intense_fit, type="response")[length(predict(ibtracs_global_results$intense_fit, type="response"))]-predict(ibtracs_global_results$intense_fit, type="response")[1])/length(predict(ibtracs_global_results$intense_fit, type="response"))*10,
                                    (predict(hursat_global_results$intense_fit, type="response")[length(predict(hursat_global_results$intense_fit, type="response"))]-predict(hursat_global_results$intense_fit, type="response")[1])/length(predict(hursat_global_results$intense_fit, type="response"))*10,
                                    (predict(trimmed_global_results$intense_fit, type="response")[length(predict(trimmed_global_results$intense_fit, type="response"))]-predict(trimmed_global_results$intense_fit, type="response")[1])/length(predict(trimmed_global_results$intense_fit, type="response"))*10 ) ) |>
  knitr::kable(digits=5)
```

**SCUSUM Change Point Test**

```{r, echo=FALSE}
data.frame(Data=c("IBTracs 1980-2025", "HURSAT 1981-2017", "IBTracs 1981-2017"),
           `Chpt Location` = c(ibtracs_global_results$intense_pois_chpt_test[2]+1980,hursat_global_results$intense_pois_chpt_test[2]+1981,trimmed_global_results$intense_pois_chpt_test[2]+1981),
           `p-value` =c(ibtracs_global_results$intense_pois_chpt_test[3],hursat_global_results$intense_pois_chpt_test[3],trimmed_global_results$intense_pois_chpt_test[3]),
           `Regime 1` = c(mean(ibtracs_global_results$data$Intense_Storms[1:ibtracs_global_results$intense_pois_chpt_test[2]]),
                          mean(hursat_global_results$data$Intense_Storms[1:hursat_global_results$intense_pois_chpt_test[2]]),
                          mean(trimmed_global_results$data$Intense_Storms[1:trimmed_global_results$intense_pois_chpt_test[2]]) ),
           `Regime 2` = c(mean(ibtracs_global_results$data$Intense_Storms[(ibtracs_global_results$intense_pois_chpt_test[2]+1):length(ibtracs_global_results$data$Intense_Storms)]),
                          mean(hursat_global_results$data$Intense_Storms[(hursat_global_results$intense_pois_chpt_test[2]+1):length(hursat_global_results$data$Intense_Storms)]),
                            mean(trimmed_global_results$data$Intense_Storms[(trimmed_global_results$intense_pois_chpt_test[2]+1):length(trimmed_global_results$data$Intense_Storms)]) ),
           `Change` = c(mean(ibtracs_global_results$data$Intense_Storms[(ibtracs_global_results$intense_pois_chpt_test[2]+1):length(ibtracs_global_results$data$Intense_Storms)]) - mean(ibtracs_global_results$data$Intense_Storms[1:ibtracs_global_results$intense_pois_chpt_test[2]]),
                        mean(hursat_global_results$data$Intense_Storms[(hursat_global_results$intense_pois_chpt_test[2]+1):length(hursat_global_results$data$Intense_Storms)]) - mean(hursat_global_results$data$Intense_Storms[1:hursat_global_results$intense_pois_chpt_test[2]]),
                        mean(trimmed_global_results$data$Intense_Storms[(trimmed_global_results$intense_pois_chpt_test[2]+1):length(trimmed_global_results$data$Intense_Storms)]) - mean(trimmed_global_results$data$Intense_Storms[1:trimmed_global_results$intense_pois_chpt_test[2]]) ) ) |>
  mutate(Perc_Change = Change/`Regime.1`*100) |>
  knitr::kable(digits=5)
```

### Proportion of Intense Storms

**Logistic Regression**

```{r, echo=FALSE}
data.frame(Data=c("IBTracs 1980-2025", "HURSAT 1981-2017", "IBTracs 1981-2017"),
           Coefficent = c(coef(ibtracs_global_results$intense_logit_fit)[2],coef(hursat_global_results$intense_logit_fit)[2],coef(trimmed_global_results$intense_logit_fit)[2]),
           `p-value` = c(anova(ibtracs_global_results$intense_logit_fit)$`Pr(>Chi)`[2], anova(hursat_global_results$intense_logit_fit)$`Pr(>Chi)`[2], anova(trimmed_global_results$intense_logit_fit)$`Pr(>Chi)`[2]),
                      `Start` = c(predict(ibtracs_global_results$intense_logit_fit, type="response")[1]*100, 
                       predict(hursat_global_results$intense_logit_fit, type="response")[1]*100, 
                       predict(trimmed_global_results$intense_logit_fit, type="response")[1]*100 ),
           `End` = c(predict(ibtracs_global_results$intense_logit_fit,type="response")[length(predict(ibtracs_global_results$intense_fit, type="response"))]*100,
                     predict(hursat_global_results$intense_logit_fit, type="response")[length(predict(hursat_global_results$intense_logit_fit, type="response"))]*100, 
                     predict(trimmed_global_results$intense_logit_fit, type="response")[length(predict(trimmed_global_results$intense_logit_fit, type="response"))]*100 ),
           `Change per Decade` = c( (predict(ibtracs_global_results$intense_logit_fit, type="response")[length(predict(ibtracs_global_results$intense_logit_fit, type="response"))]-predict(ibtracs_global_results$intense_logit_fit, type="response")[1])/length(predict(ibtracs_global_results$intense_logit_fit, type="response"))*10*100,
                                    (predict(hursat_global_results$intense_logit_fit, type="response")[length(predict(hursat_global_results$intense_logit_fit, type="response"))]-predict(hursat_global_results$intense_logit_fit, type="response")[1])/length(predict(hursat_global_results$intense_logit_fit, type="response"))*10*100,
                                    (predict(trimmed_global_results$intense_logit_fit, type="response")[length(predict(trimmed_global_results$intense_logit_fit, type="response"))]-predict(trimmed_global_results$intense_logit_fit, type="response")[1])/length(predict(trimmed_global_results$intense_logit_fit, type="response"))*10*100 ) ) |>
  knitr::kable()
```

**SCUSUM Change Point Test**

```{r, echo=FALSE}
data.frame(Data=c("IBTracs 1980-2025", "HURSAT 1981-2017", "IBTracs 1981-2017"),
           `Chpt Location` = c(ibtracs_global_results$prop_intense_chpt_test[2]+1980,hursat_global_results$prop_intense_chpt_test[2]+1981,trimmed_global_results$prop_intense_chpt_test[2]+1981),
           `p-value` =c(ibtracs_global_results$prop_intense_chpt_test[3],hursat_global_results$prop_intense_chpt_test[3],trimmed_global_results$prop_intense_chpt_test[3]),
           `Regime 1 %` = c(sum(ibtracs_global_results$data$Intense_Storms[1:ibtracs_global_results$prop_intense_chpt_test[2]])/sum(ibtracs_global_results$data$Total_Storms[1:ibtracs_global_results$prop_intense_chpt_test[2]])*100,
                          sum(hursat_global_results$data$Intense_Storms[1:hursat_global_results$prop_intense_chpt_test[2]])/sum(hursat_global_results$data$Total_Storms[1:hursat_global_results$prop_intense_chpt_test[2]])*100,
                          sum(ibtracs_global_results$data$Intense_Storms[1:ibtracs_global_results$prop_intense_chpt_test[2]])/sum(ibtracs_global_results$data$Total_Storms[1:ibtracs_global_results$prop_intense_chpt_test[2]])*100 ),
           `Regime 2 %` = c(sum(ibtracs_global_results$data$Intense_Storms[(ibtracs_global_results$prop_intense_chpt_test[2]+1):length(ibtracs_global_results$data$Intense_Storms)])/sum(ibtracs_global_results$data$Total_Storms[(ibtracs_global_results$prop_intense_chpt_test[2]+1):length(ibtracs_global_results$data$Intense_Storms)])*100,
                          sum(hursat_global_results$data$Intense_Storms[(hursat_global_results$prop_intense_chpt_test[2]+1):length(hursat_global_results$data$Intense_Storms)])/sum(hursat_global_results$data$Total_Storms[(hursat_global_results$prop_intense_chpt_test[2]+1):length(hursat_global_results$data$Intense_Storms)])*100,
                          sum(trimmed_global_results$data$Intense_Storms[(trimmed_global_results$prop_intense_chpt_test[2]+1):length(trimmed_global_results$data$Intense_Storms)])/sum(trimmed_global_results$data$Total_Storms[(trimmed_global_results$prop_intense_chpt_test[2]+1):length(trimmed_global_results$data$Intense_Storms)])*100 ),
           `Change %` = c((sum(ibtracs_global_results$data$Intense_Storms[(ibtracs_global_results$prop_intense_chpt_test[2]+1):length(ibtracs_global_results$data$Intense_Storms)])/sum(ibtracs_global_results$data$Total_Storms[(ibtracs_global_results$prop_intense_chpt_test[2]+1):length(ibtracs_global_results$data$Intense_Storms)]) - sum(ibtracs_global_results$data$Intense_Storms[1:ibtracs_global_results$prop_intense_chpt_test[2]])/sum(ibtracs_global_results$data$Total_Storms[1:ibtracs_global_results$prop_intense_chpt_test[2]]) )*100,
                        (sum(hursat_global_results$data$Intense_Storms[(hursat_global_results$prop_intense_chpt_test[2]+1):length(hursat_global_results$data$Intense_Storms)])/sum(hursat_global_results$data$Total_Storms[(hursat_global_results$prop_intense_chpt_test[2]+1):length(hursat_global_results$data$Intense_Storms)]) - sum(hursat_global_results$data$Intense_Storms[1:hursat_global_results$prop_intense_chpt_test[2]])/sum(hursat_global_results$data$Total_Storms[1:hursat_global_results$prop_intense_chpt_test[2]]) )*100,
                        (sum(trimmed_global_results$data$Intense_Storms[(trimmed_global_results$prop_intense_chpt_test[2]+1):length(trimmed_global_results$data$Intense_Storms)])/sum(trimmed_global_results$data$Total_Storms[(trimmed_global_results$prop_intense_chpt_test[2]+1):length(trimmed_global_results$data$Intense_Storms)]) - sum(trimmed_global_results$data$Intense_Storms[1:trimmed_global_results$prop_intense_chpt_test[2]])/sum(trimmed_global_results$data$Total_Storms[1:trimmed_global_results$prop_intense_chpt_test[2]]) )*100 ) ) |>
  knitr::kable()
```

