

## HURSAT Quick Results


### Summary Statistics and Data

```{r, echo=FALSE}
hursat_ni_results$summary_stat |>
  knitr::kable()
```

```{r, echo=FALSE, warning=FALSE}
hursat_ni_results$plot_data
```

### Testing Results

**Changepoint Methods**

Here, we report the estimate mean for each segment or regime regardless of whether the test found a significant shift.  

```{r, echo=FALSE}
df <- data.frame(
  Response = c("Cyclone Freq", "Hurricanes", "Major Freq", "Prop Major", "Intense Freq", "Prop Freq"),
  `Chpt Location` = c(hursat_ni_results$pois_chpt_test[2]+1981,
                      hursat_ni_results$hur_pois_chpt_test[2]+1981,
                      hursat_ni_results$major_pois_chpt_test[2]+1981,
                      hursat_ni_results$prop_major_chpt_test[2]+1981,
                      hursat_ni_results$intense_pois_chpt_test[2]+1981,
                      hursat_ni_results$prop_intense_chpt_test[2]+1981),
  `p-value` =c(hursat_ni_results$pois_chpt_test[3],
               hursat_ni_results$hur_pois_chpt_test[3],
               hursat_ni_results$major_pois_chpt_test[3],
               hursat_ni_results$prop_major_chpt_test[3],
               hursat_ni_results$intense_pois_chpt_test[3],
               hursat_ni_results$prop_intense_chpt_test[3] ),
  `Regime1` = c(mean(hursat_ni_results$data$Total_Storms[1:hursat_ni_results$pois_chpt_test[2]]),
                mean(hursat_ni_results$data$Hurricanes[1:hursat_ni_results$hur_pois_chpt_test[2]]),
                mean(hursat_ni_results$data$Major_Storms[1:hursat_ni_results$major_pois_chpt_test[2]]),
                sum(hursat_ni_results$data$Major_Storms[1:hursat_ni_results$prop_major_chpt_test[2]])/sum(hursat_ni_results$data$Total_Storms[1:hursat_ni_results$prop_major_chpt_test[2]])*100,
                mean(hursat_ni_results$data$Intense_Storms[1:hursat_ni_results$intense_pois_chpt_test[2]]),
                sum(hursat_ni_results$data$Intense_Storms[1:hursat_ni_results$prop_intense_chpt_test[2]])/sum(hursat_ni_results$data$Total_Storms[1:hursat_ni_results$prop_intense_chpt_test[2]])*100 ),
  `Regime2` = c(mean(hursat_ni_results$data$Total_Storms[(hursat_ni_results$pois_chpt_test[2]+1):num_years_hursat]),
                mean(hursat_ni_results$data$Hurricanes[(hursat_ni_results$hur_pois_chpt_test[2]+1):num_years_hursat]),
                mean(hursat_ni_results$data$Major_Storms[(hursat_ni_results$major_pois_chpt_test[2]+1):num_years_hursat]),
                sum(hursat_ni_results$data$Major_Storms[(hursat_ni_results$prop_major_chpt_test[2]+1):num_years_hursat])/sum(hursat_ni_results$data$Total_Storms[(hursat_ni_results$prop_major_chpt_test[2]+1):num_years_hursat])*100,
                mean(hursat_ni_results$data$Intense_Storms[(hursat_ni_results$intense_pois_chpt_test[2]+1):num_years_hursat]),
                sum(hursat_ni_results$data$Intense_Storms[(hursat_ni_results$prop_intense_chpt_test[2]+1):num_years_hursat])/sum(hursat_ni_results$data$Total_Storms[(hursat_ni_results$prop_intense_chpt_test[2]+1):num_years_hursat])*100)
  )

df |>
  mutate(Change = Regime2-Regime1,
         Perc_Change = Change/Regime1*100) |>
  knitr::kable(digits=5)
  
```


**Generalized Linear Model Approaches**

Note:

* we report the Likelihood Ratio Test for the generalized linear models, this is better to use than the default $Z$-test provided in the `summary()` output.
* the Change per decade is found by comparing the predicted value at the start time (`r min(hursat_ni_results$data$SEASON)`) and end time (`r max(hursat_ni_results$data$SEASON)`) and scaling by the number of years $\times 10$.

```{r, echo=FALSE}
df <- data.frame(
  Response = c("Cyclone Freq", "Hurricanes", "Major Freq", "Prop Major", "Intense Freq", "Prop Freq"),
  Coefficent = c(coef(hursat_ni_results$pois_fit)[2], 
                 coef(hursat_ni_results$hur_fit)[2],
                 coef(hursat_ni_results$major_fit)[2], 
                 coef(hursat_ni_results$major_logit_fit)[2], 
                 coef(hursat_ni_results$intense_fit)[2], 
                 coef(hursat_ni_results$intense_logit_fit)[2] ),
  `p-value` = c(anova(hursat_ni_results$pois_fit)$`Pr(>Chi)`[2], 
                anova(hursat_ni_results$hur_fit)$`Pr(>Chi)`[2], 
                anova(hursat_ni_results$major_fit)$`Pr(>Chi)`[2], 
                anova(hursat_ni_results$major_logit_fit)$`Pr(>Chi)`[2], 
                anova(hursat_ni_results$intense_fit)$`Pr(>Chi)`[2], 
                anova(hursat_ni_results$intense_logit_fit)$`Pr(>Chi)`[2]),
  `Start` = c(predict(hursat_ni_results$pois_fit, type="response")[1], 
              predict(hursat_ni_results$hur_fit, type="response")[1],
              predict(hursat_ni_results$major_fit, type="response")[1], 
              predict(hursat_ni_results$major_logit_fit, type="response")[1]*100, 
              predict(hursat_ni_results$intense_fit, type="response")[1],
              predict(hursat_ni_results$intense_logit_fit, type="response")[1]*100 ),
  `End` = c(predict(hursat_ni_results$pois_fit, type="response")[num_years_hursat],
            predict(hursat_ni_results$hur_fit, type="response")[num_years_hursat],
            predict(hursat_ni_results$major_fit, type="response")[num_years_hursat],
            predict(hursat_ni_results$major_logit_fit, type="response")[num_years_hursat]*100,
            predict(hursat_ni_results$intense_fit, type="response")[num_years_hursat],
            predict(hursat_ni_results$intense_logit_fit, type="response")[num_years_hursat]*100)
)

df |>
  mutate(`Change per Decade` = (End- Start)/num_years_hursat*10) |>
  knitr::kable(digits=5)
```



## HURSAT Full Analysis


### Changepoint Methods

#### Change in Non-Shorty Counts

```{r, echo=FALSE}
hursat_ni_results$pois_chpt_test
```

The dates suggest the changepoint would occur from `r 1980+hursat_ni_results$pois_chpt_test[2]` to `r 1980+hursat_ni_results$pois_chpt_test[2]+1` with $p$-value `r hursat_ni_results$pois_chpt_test[3]`.


In regime one, the mean number of storms is `r predict(hursat_ni_results$pois_chpt_model, type="response")[1]` (CI: `r exp(predict(hursat_ni_results$pois_chpt_model, type="link")[1]-1.96*predict(hursat_ni_results$pois_chpt_model, type="link", se=TRUE)$se.fit[1])`, `r exp(predict(hursat_ni_results$pois_chpt_model, type="link")[1]+1.96*predict(hursat_ni_results$pois_chpt_model, type="link", se=TRUE)$se.fit[1])`) and in regime two, the mean number of storms is `r predict(hursat_ni_results$pois_chpt_model, type="response")[num_years_hursat]` (CI: `r exp(predict(hursat_ni_results$pois_chpt_model, type="link")[num_years_hursat]-1.96*predict(hursat_ni_results$pois_chpt_model, type="link", se=TRUE)$se.fit[num_years_hursat])`, `r exp(predict(hursat_ni_results$pois_chpt_model, type="link")[num_years_hursat]+1.96*predict(hursat_ni_results$pois_chpt_model, type="link", se=TRUE)$se.fit[num_years_hursat])`); a change of `r predict(hursat_ni_results$pois_chpt_model, type="response")[num_years_hursat] - predict(hursat_ni_results$pois_chpt_model, type="response")[1]`.


**Model Assessment**

The ratio of the residual deviance is `r hursat_ni_results$pois_chpt_model$deviance/hursat_ni_results$pois_chpt_model$df.residual`.

The $p$-value for the Ljung-Box test on the deviance residuals `r Box.test(residuals(hursat_ni_results$pois_chpt_model, type="deviance"), lag=ceiling(log(num_years_hursat)))$p.value`.

```{r, message=FALSE, warning=FALSE, echo=FALSE, fig.height=8}
residual_plots(hursat_ni_results$pois_chpt_model)
```


#### Change in counts of hurricanes

```{r, echo=FALSE}
hursat_ni_results$hur_pois_chpt_test
```

The dates suggest the changepoint would occur from `r 1980+hursat_ni_results$hur_pois_chpt_test[2]` to `r 1980+hursat_ni_results$hur_pois_chpt_test[2]+1` with $p$-value `r hursat_ni_results$hur_pois_chpt_test[3]`.


In regime one, the mean number of hurricanes is `r predict(hursat_ni_results$hur_chpt_model, type="response")[1]` (CI: `r exp(predict(hursat_ni_results$hur_chpt_model, type="link")[1]-1.96*predict(hursat_ni_results$hur_chpt_model, type="link", se=TRUE)$se.fit[1])`, `r exp(predict(hursat_ni_results$hur_chpt_model, type="link")[1]+1.96*predict(hursat_ni_results$hur_chpt_model, type="link", se=TRUE)$se.fit[1])`) and in regime two, the mean number of hurricanes is `r predict(hursat_ni_results$hur_chpt_model, type="response")[num_years_hursat]` (CI: `r exp(predict(hursat_ni_results$hur_chpt_model, type="link")[num_years_hursat]-1.96*predict(hursat_ni_results$hur_chpt_model, type="link", se=TRUE)$se.fit[num_years_hursat])`, `r exp(predict(hursat_ni_results$hur_chpt_model, type="link")[num_years_hursat]+1.96*predict(hursat_ni_results$hur_chpt_model, type="link", se=TRUE)$se.fit[num_years_hursat])`); a change of `r predict(hursat_ni_results$hur_chpt_model, type="response")[num_years_hursat] - predict(hursat_ni_results$hur_chpt_model, type="response")[1]`.



**Model Assessment**

The ratio of the residual deviance is `r hursat_ni_results$hur_chpt_model$deviance/hursat_ni_results$hur_chpt_model$df.residual`.

The $p$-value for the Ljung-Box test on the deviance residuals `r Box.test(residuals(hursat_ni_results$hur_chpt_model, type="deviance"), lag=ceiling(log(num_years_hursat)))$p.value`.

```{r, message=FALSE, warning=FALSE, echo=FALSE, fig.height=8}
residual_plots(hursat_ni_results$hur_chpt_model)
```

#### Change in counts of major storms

```{r, echo=FALSE}
hursat_ni_results$major_pois_chpt_test
```

The dates suggest the changepoint would occur from `r 1980+hursat_ni_results$major_pois_chpt_test[2]` to `r 1980+hursat_ni_results$major_pois_chpt_test[2]+1` with $p$-value `r hursat_ni_results$major_pois_chpt_test[3]`.


In regime one, the mean number of storms is `r predict(hursat_ni_results$major_pois_chpt_model, type="response")[1]` (CI: `r exp(predict(hursat_ni_results$major_pois_chpt_model, type="link")[1]-1.96*predict(hursat_ni_results$major_pois_chpt_model, type="link", se=TRUE)$se.fit[1])`, `r exp(predict(hursat_ni_results$major_pois_chpt_model, type="link")[1]+1.96*predict(hursat_ni_results$major_pois_chpt_model, type="link", se=TRUE)$se.fit[1])`) and in regime two, the mean number of storms is `r predict(hursat_ni_results$major_pois_chpt_model, type="response")[num_years_hursat]` (CI: `r exp(predict(hursat_ni_results$major_pois_chpt_model, type="link")[num_years_hursat]-1.96*predict(hursat_ni_results$major_pois_chpt_model, type="link", se=TRUE)$se.fit[num_years_hursat])`, `r exp(predict(hursat_ni_results$major_pois_chpt_model, type="link")[num_years_hursat]+1.96*predict(hursat_ni_results$major_pois_chpt_model, type="link", se=TRUE)$se.fit[num_years_hursat])`); a change of `r predict(hursat_ni_results$major_pois_chpt_model, type="response")[num_years_hursat] - predict(hursat_ni_results$major_pois_chpt_model, type="response")[1]`.



**Model Assessment**

The ratio of the residual deviance is `r hursat_ni_results$major_pois_chpt_model$deviance/hursat_ni_results$major_pois_chpt_model$df.residual`.

The $p$-value for the Ljung-Box test on the deviance residuals `r Box.test(residuals(hursat_ni_results$major_pois_chpt_model, type="deviance"), lag=ceiling(log(num_years_hursat)))$p.value`.

```{r, message=FALSE, warning=FALSE, echo=FALSE, fig.height=8}
residual_plots(hursat_ni_results$major_pois_chpt_model)
```




#### Change in counts of intense storms

```{r, echo=FALSE}
hursat_ni_results$intense_pois_chpt_test
```


The dates suggest the changepoint would occur from `r 1980+hursat_ni_results$intense_pois_chpt_test[2]` to `r 1980+hursat_ni_results$intense_pois_chpt_test[2]+1` with $p$-value `r hursat_ni_results$intense_pois_chpt_test[3]`.


In regime one, the mean number of storms is `r predict(hursat_ni_results$intense_pois_chpt_model, type="response")[1]` (CI: `r exp(predict(hursat_ni_results$intense_pois_chpt_model, type="link")[1]-1.96*predict(hursat_ni_results$intense_pois_chpt_model, type="link", se=TRUE)$se.fit[1])`, `r exp(predict(hursat_ni_results$intense_pois_chpt_model, type="link")[1]+1.96*predict(hursat_ni_results$intense_pois_chpt_model, type="link", se=TRUE)$se.fit[1])`) and in regime two, the mean number of storms is `r predict(hursat_ni_results$intense_pois_chpt_model, type="response")[num_years_hursat]` (CI: `r exp(predict(hursat_ni_results$intense_pois_chpt_model, type="link")[num_years_hursat]-1.96*predict(hursat_ni_results$intense_pois_chpt_model, type="link", se=TRUE)$se.fit[num_years_hursat])`, `r exp(predict(hursat_ni_results$intense_pois_chpt_model, type="link")[num_years_hursat]+1.96*predict(hursat_ni_results$intense_pois_chpt_model, type="link", se=TRUE)$se.fit[num_years_hursat])`); a change of `r predict(hursat_ni_results$intense_pois_chpt_model, type="response")[num_years_hursat] - predict(hursat_ni_results$intense_pois_chpt_model, type="response")[1]`.


**Model Assessment**

The ratio of the residual deviance is `r hursat_ni_results$intense_pois_chpt_model$deviance/hursat_ni_results$intense_pois_chpt_model$df.residual`.

The $p$-value for the Ljung-Box test on the deviance residuals `r Box.test(residuals(hursat_ni_results$intense_pois_chpt_model, type="deviance"), lag=ceiling(log(num_years_hursat)))$p.value`.

```{r, message=FALSE, warning=FALSE, echo=FALSE, fig.height=8}
residual_plots(hursat_ni_results$intense_pois_chpt_model)
```



#### Proportion of major storms

A change in the proportion of major storms?
  
```{r, echo=FALSE}
hursat_ni_results$prop_major_chpt_test
```

The change in proportions occurs `r 1980+hursat_ni_results$prop_major_chpt_test[2]` to `r 1980+hursat_ni_results$prop_major_chpt_test[2]+1` with $p$-value `r hursat_ni_results$prop_major_chpt_test[3]`. 

In regime one, the proportion of major storms is `r plogis(predict(hursat_ni_results$prop_major_chpt_model)[1])` (CI: `r plogis(predict(hursat_ni_results$prop_major_chpt_model)[1] - 1.96*predict(hursat_ni_results$prop_major_chpt_model, se=TRUE)$se.fit[1])`, `r plogis(predict(hursat_ni_results$prop_major_chpt_model)[1] + 1.96*predict(hursat_ni_results$prop_major_chpt_model, se=TRUE)$se.fit[1])`) and in regime two, the proportion of major storms is `r plogis(predict(hursat_ni_results$prop_major_chpt_model)[num_years_hursat])` (CI: `r plogis(predict(hursat_ni_results$prop_major_chpt_model)[num_years_hursat] - 1.96*predict(hursat_ni_results$prop_major_chpt_model, se=TRUE)$se.fit[num_years_hursat])`, `r plogis(predict(hursat_ni_results$prop_major_chpt_model)[num_years_hursat] + 1.96*predict(hursat_ni_results$prop_major_chpt_model, se=TRUE)$se.fit[num_years_hursat])`); a change of `r plogis(predict(hursat_ni_results$prop_major_chpt_model)[num_years_hursat]) - plogis(predict(hursat_ni_results$prop_major_chpt_model)[1])`.



**Model Assessment**

The ratio of the residual deviance is `r hursat_ni_results$prop_major_chpt_model$deviance/hursat_ni_results$prop_major_chpt_model$df.residual`.

The $p$-value for the Ljung-Box test on the deviance residuals `r Box.test(residuals(hursat_ni_results$prop_major_chpt_model, type="deviance"), lag=ceiling(log(num_years_hursat)))$p.value`.

```{r, message=FALSE, warning=FALSE, echo=FALSE, fig.height=8}
residual_plots(hursat_ni_results$prop_major_chpt_model)
```



#### Proportion of intense storms

A change in the proportion of intense storms?
  
```{r, echo=FALSE}
hursat_ni_results$prop_intense_chpt_test
```

The change in proportions occurs `r 1980+hursat_ni_results$prop_intense_chpt_test[2]` to `r 1980+hursat_ni_results$prop_intense_chpt_test[2]+1` with $p$-value `r hursat_ni_results$prop_intense_chpt_test[3]`.


In regime one, the proportion of intense storms is `r plogis(predict(hursat_ni_results$prop_intense_chpt_model)[1])` (CI: `r plogis(predict(hursat_ni_results$prop_intense_chpt_model)[1] - 1.96*predict(hursat_ni_results$prop_intense_chpt_model, se=TRUE)$se.fit[1])`, `r plogis(predict(hursat_ni_results$prop_intense_chpt_model)[1] + 1.96*predict(hursat_ni_results$prop_intense_chpt_model, se=TRUE)$se.fit[1])`) and in regime two, the proportion of intense storms is `r plogis(predict(hursat_ni_results$prop_intense_chpt_model)[num_years_hursat])` (CI: `r plogis(predict(hursat_ni_results$prop_intense_chpt_model)[num_years_hursat] - 1.96*predict(hursat_ni_results$prop_intense_chpt_model, se=TRUE)$se.fit[num_years_hursat])`, `r plogis(predict(hursat_ni_results$prop_intense_chpt_model)[num_years_hursat] + 1.96*predict(hursat_ni_results$prop_intense_chpt_model, se=TRUE)$se.fit[num_years_hursat])`); a change of `r plogis(predict(hursat_ni_results$prop_intense_chpt_model)[num_years_hursat]) - plogis(predict(hursat_ni_results$prop_intense_chpt_model)[1])`.

**Model Assessment**

The ratio of the residual deviance is `r hursat_ni_results$prop_intense_chpt_model$deviance/hursat_ni_results$prop_intense_chpt_model$df.residual`.

The $p$-value for the Ljung-Box test on the deviance residuals `r Box.test(residuals(hursat_ni_results$prop_intense_chpt_model, type="deviance"), lag=ceiling(log(num_years_hursat)))$p.value`.

```{r, message=FALSE, warning=FALSE, echo=FALSE, fig.height=8}
residual_plots(hursat_ni_results$prop_intense_chpt_model)
```



### Regression Methods

#### Poisson Regression on non-Shorties

```{r, echo=FALSE}
(hursat_ni_results$pois_fit)
drop1(hursat_ni_results$pois_fit, test="LRT")
```


Using this model, in `r min(hursat_ni_results$data$SEASON)` we would predict `r predict(hursat_ni_results$pois_fit, type="response")[1]` storms (CI: `r exp(predict(hursat_ni_results$pois_fit, type="link")[1]-1.96*predict(hursat_ni_results$pois_fit, type="link", se.fit=TRUE)$se.fit[1])`, `r exp(predict(hursat_ni_results$pois_fit, type="link")[1]+1.96*predict(hursat_ni_results$pois_fit, type="link", se.fit=TRUE)$se.fit[1])`) and in `r max(hursat_ni_results$data$SEASON)` we predict `r predict(hursat_ni_results$pois_fit, type="response")[num_years_hursat]` storms (CI: `r exp(predict(hursat_ni_results$pois_fit, type="link")[num_years_hursat]-1.96*predict(hursat_ni_results$pois_fit, type="link", se.fit=TRUE)$se.fit[num_years_hursat])`, `r exp(predict(hursat_ni_results$pois_fit, type="link")[num_years_hursat]+1.96*predict(hursat_ni_results$pois_fit, type="link", se.fit=TRUE)$se.fit[num_years_hursat])`).

That is a change of `r (predict(hursat_ni_results$pois_fit, type="response")[num_years_hursat]-predict(hursat_ni_results$pois_fit, type="response")[1])/num_years_hursat*10` storms per decade.

**Model Assessment**

The ratio of the residual deviance is `r hursat_ni_results$pois_fit$deviance/hursat_ni_results$pois_fit$df.residual`.

The $p$-value for the Ljung-Box test on the deviance residuals `r Box.test(residuals(hursat_ni_results$pois_fit, type="deviance"), lag=ceiling(log(num_years_hursat)))$p.value`.

```{r, message=FALSE, warning=FALSE, echo=FALSE, fig.height=8}
residual_plots(hursat_ni_results$pois_fit)
```



#### Poisson Regression on Hurricanes

```{r, echo=FALSE}
(hursat_ni_results$hur_fit)
drop1(hursat_ni_results$hur_fit, test="LRT")
```

Using this model, in `r min(hursat_ni_results$data$SEASON)` we would predict `r predict(hursat_ni_results$hur_fit, type="response")[1]` hurricanes (CI: `r exp(predict(hursat_ni_results$hur_fit, type="link")[1]-1.96*predict(hursat_ni_results$hur_fit, type="link", se.fit=TRUE)$se.fit[1])`, `r exp(predict(hursat_ni_results$hur_fit, type="link")[1]+1.96*predict(hursat_ni_results$hur_fit, type="link", se.fit=TRUE)$se.fit[1])`) and in `r max(hursat_ni_results$data$SEASON)` we predict `r predict(hursat_ni_results$hur_fit, type="response")[num_years_hursat]` hurricanes (CI: `r exp(predict(hursat_ni_results$hur_fit, type="link")[num_years_hursat]-1.96*predict(hursat_ni_results$hur_fit, type="link", se.fit=TRUE)$se.fit[num_years_hursat])`, `r exp(predict(hursat_ni_results$hur_fit, type="link")[num_years_hursat]+1.96*predict(hursat_ni_results$hur_fit, type="link", se.fit=TRUE)$se.fit[num_years_hursat])`).

That is a change of `r (predict(hursat_ni_results$hur_fit, type="response")[num_years_hursat]-predict(hursat_ni_results$hur_fit, type="response")[1])/num_years_hursat*10` hurricanes per decade.

**Model Assessment**

The ratio of the residual deviance is `r hursat_ni_results$hur_fit$deviance/hursat_ni_results$hur_fit$df.residual`.

The $p$-value for the Ljung-Box test on the deviance residuals `r Box.test(residuals(hursat_ni_results$hur_fit, type="deviance"), lag=ceiling(log(num_years_hursat)))$p.value`.

```{r, message=FALSE, warning=FALSE, echo=FALSE, fig.height=8}
residual_plots(hursat_ni_results$hur_fit)
```



#### Poisson Regression on Major Storms

```{r, echo=FALSE}
(hursat_ni_results$major_fit)
drop1(hursat_ni_results$major_fit, test="LRT")
```

Using this model, in `r min(hursat_ni_results$data$SEASON)` we would predict `r predict(hursat_ni_results$major_fit, type="response")[1]` major storms (CI: `r exp(predict(hursat_ni_results$major_fit, type="link")[1]-1.96*predict(hursat_ni_results$major_fit, type="link", se.fit=TRUE)$se.fit[1])`, `r exp(predict(hursat_ni_results$major_fit, type="link")[1]+1.96*predict(hursat_ni_results$major_fit, type="link", se.fit=TRUE)$se.fit[1])`) and in `r max(hursat_ni_results$data$SEASON)` we predict `r predict(hursat_ni_results$major_fit, type="response")[num_years_hursat]` major storms (CI: `r exp(predict(hursat_ni_results$major_fit, type="link")[num_years_hursat]-1.96*predict(hursat_ni_results$major_fit, type="link", se.fit=TRUE)$se.fit[num_years_hursat])`, `r exp(predict(hursat_ni_results$major_fit, type="link")[num_years_hursat]+1.96*predict(hursat_ni_results$major_fit, type="link", se.fit=TRUE)$se.fit[num_years_hursat])`).

That is a change of `r (predict(hursat_ni_results$major_fit, type="response")[num_years_hursat]-predict(hursat_ni_results$major_fit, type="response")[1])/num_years_hursat*10` major storms per decade.

**Model Assessment**

The ratio of the residual deviance is `r hursat_ni_results$major_fit$deviance/hursat_ni_results$major_fit$df.residual`.

The $p$-value for the Ljung-Box test on the deviance residuals `r Box.test(residuals(hursat_ni_results$major_fit, type="deviance"), lag=ceiling(log(num_years_hursat)))$p.value`.

```{r, message=FALSE, warning=FALSE, echo=FALSE, fig.height=8}
residual_plots(hursat_ni_results$major_fit)
```




#### Poisson Regression on Intense Storms

```{r, echo=FALSE}
(hursat_ni_results$intense_fit)
drop1(hursat_ni_results$intense_fit, test="LRT")
```


Using this model, in `r min(hursat_ni_results$data$SEASON)` we would predict `r predict(hursat_ni_results$intense_fit, type="response")[1]` intense storms (CI: `r exp(predict(hursat_ni_results$intense_fit, type="link")[1]-1.96*predict(hursat_ni_results$intense_fit, type="link", se.fit=TRUE)$se.fit[1])`, `r exp(predict(hursat_ni_results$intense_fit, type="link")[1]+1.96*predict(hursat_ni_results$intense_fit, type="link", se.fit=TRUE)$se.fit[1])`) and in `r max(hursat_ni_results$data$SEASON)` we predict `r predict(hursat_ni_results$intense_fit, type="response")[num_years_hursat]` intense storms (CI: `r exp(predict(hursat_ni_results$intense_fit, type="link")[num_years_hursat]-1.96*predict(hursat_ni_results$intense_fit, type="link", se.fit=TRUE)$se.fit[num_years_hursat])`, `r exp(predict(hursat_ni_results$intense_fit, type="link")[num_years_hursat]+1.96*predict(hursat_ni_results$intense_fit, type="link", se.fit=TRUE)$se.fit[num_years_hursat])`).

That is a change of `r (predict(hursat_ni_results$intense_fit, type="response")[num_years_hursat]-predict(hursat_ni_results$intense_fit, type="response")[1])/num_years_hursat*10` intense storms per decade.

**Model Assessment**

The ratio of the residual deviance is `r hursat_ni_results$intense_fit$deviance/hursat_ni_results$intense_fit$df.residual`.

The $p$-value for the Ljung-Box test on the deviance residuals `r Box.test(residuals(hursat_ni_results$intense_fit, type="deviance"), lag=ceiling(log(num_years_hursat)))$p.value`.

```{r, message=FALSE, warning=FALSE, echo=FALSE, fig.height=8}
residual_plots(hursat_ni_results$intense_fit)
```




#### Logistic Regression on Proportions of Major

```{r, echo=FALSE}
(hursat_ni_results$major_logit_fit)
drop1(hursat_ni_results$major_logit_fit, test="LRT")
```



Using this model, in `r min(hursat_ni_results$data$SEASON)` the probability of a major storm in  was `r predict(hursat_ni_results$major_logit_fit, type="response")[1]` (CI: `r plogis(predict(hursat_ni_results$major_logit_fit, type="link")[1]-1.96*predict(hursat_ni_results$major_logit_fit, type="link", se=TRUE)$se.fit[1])`, `r plogis(predict(hursat_ni_results$major_logit_fit, type="link")[1]+1.96*predict(hursat_ni_results$major_logit_fit, type="link", se=TRUE)$se.fit[1])`)  and in `r max(hursat_ni_results$data$SEASON)` the probability of a major storm was `r predict(hursat_ni_results$major_logit_fit, type="response")[num_years_hursat]` (CI: `r plogis(predict(hursat_ni_results$major_logit_fit, type="link")[num_years_hursat]-1.96*predict(hursat_ni_results$major_logit_fit, type="link", se=TRUE)$se.fit[num_years_hursat])`, `r plogis(predict(hursat_ni_results$major_logit_fit, type="link")[num_years_hursat]+1.96*predict(hursat_ni_results$major_logit_fit, type="link", se=TRUE)$se.fit[num_years_hursat])`).

That is a change of `r (predict(hursat_ni_results$major_logit_fit, type="response")[num_years_hursat]-predict(hursat_ni_results$major_logit_fit, type="response")[1])/num_years_hursat*10*100` percent per decade.

**Model Assessment**

The ratio of the residual deviance is `r hursat_ni_results$major_logit_fit$deviance/hursat_ni_results$major_logit_fit$df.residual`.

The $p$-value for the Ljung-Box test on the deviance residuals `r Box.test(residuals(hursat_ni_results$major_logit_fit, type="deviance"), lag=ceiling(log(num_years_hursat)))$p.value`.

```{r, message=FALSE, warning=FALSE, echo=FALSE, fig.height=8}
residual_plots(hursat_ni_results$major_logit_fit)
```




#### Logistic Regression on Proportions of Intense

```{r, echo=FALSE}
(hursat_ni_results$intense_logit_fit)
drop1(hursat_ni_results$intense_logit_fit, test="LRT")
```



Using this model, in `r min(hursat_ni_results$data$SEASON)` the probability of a major storm in  was `r predict(hursat_ni_results$intense_logit_fit, type="response")[1]` (CI: `r plogis(predict(hursat_ni_results$intense_logit_fit, type="link")[1]-1.96*predict(hursat_ni_results$intense_logit_fit, type="link", se=TRUE)$se.fit[1])`, `r plogis(predict(hursat_ni_results$intense_logit_fit, type="link")[1]+1.96*predict(hursat_ni_results$intense_logit_fit, type="link", se=TRUE)$se.fit[1])`)  and in `r max(hursat_ni_results$data$SEASON)` the probability of a major storm was `r predict(hursat_ni_results$intense_logit_fit, type="response")[num_years_hursat]` (CI: `r plogis(predict(hursat_ni_results$intense_logit_fit, type="link")[num_years_hursat]-1.96*predict(hursat_ni_results$intense_logit_fit, type="link", se=TRUE)$se.fit[num_years_hursat])`, `r plogis(predict(hursat_ni_results$intense_logit_fit, type="link")[num_years_hursat]+1.96*predict(hursat_ni_results$intense_logit_fit, type="link", se=TRUE)$se.fit[num_years_hursat])`).



That is a change of `r (predict(hursat_ni_results$intense_logit_fit, type="response")[num_years_hursat]-predict(hursat_ni_results$intense_logit_fit, type="response")[1])/num_years_hursat*10*100` percent per decade.


**Model Assessment**

The ratio of the residual deviance is `r hursat_ni_results$intense_logit_fit$deviance/hursat_ni_results$intense_logit_fit$df.residual`.

The $p$-value for the Ljung-Box test on the deviance residuals `r Box.test(residuals(hursat_ni_results$intense_logit_fit, type="deviance"), lag=ceiling(log(num_years_hursat)))$p.value`.

```{r, message=FALSE, warning=FALSE, echo=FALSE, fig.height=8}
residual_plots(hursat_ni_results$intense_logit_fit)
```






