
# Western Pacific

Here, we look at the Western North Pacific basin (the most active!). 


## Full IBTracs Record

```{r, echo=FALSE}
ibtracs_wp_results <- get_report_basin(basin="Western North Pacific", all_basins=ibtracs_all_basins)
```

```{r, echo=FALSE}
ibtracs_wp_results$plot_data
```

### Summary Statistics (overall)

```{r, echo=FALSE}
ibtracs_wp_results$summary_stat |>
  knitr::kable()
```

### Regression Methods

#### Poisson Regression on non-Shorties

```{r, echo=FALSE}
summary(ibtracs_wp_results$pois_fit)
```



Note, this says a decrease of `r abs(coef(ibtracs_wp_results$pois_fit)[2])` log-storms per year. This corresponds to a `r (1-exp(coef(ibtracs_wp_results$pois_fit)[2]))*100` percent decrease per season. So over 45 years, that is a decrease of `r (1-exp(coef(ibtracs_wp_results$pois_fit)[2])^45)*100` percent.

Using this model, we would predict `r predict(ibtracs_wp_results$pois_fit, type="response")[1]` storms in 1980 and `r predict(ibtracs_wp_results$pois_fit, type="response")[45]` storms in 2024.

That is a change of `r (predict(ibtracs_wp_results$pois_fit, type="response")[45]-predict(ibtracs_wp_results$pois_fit, type="response")[1])/45*10` storms per decade.

#### Poisson Regression on Major Storms

```{r, echo=FALSE}
summary(ibtracs_wp_results$major_fit)
```

Using this model the number of a major storm in 1980 was `r predict(ibtracs_wp_results$major_fit, type="response")[1]` and `r predict(ibtracs_wp_results$major_fit, type="response")[45]` in 2024.

That is a change of `r (predict(ibtracs_wp_results$major_fit, type="response")[45]-predict(ibtracs_wp_results$major_fit, type="response")[1])/45*10` storms per decade.

#### Poisson Regression on Intense Storms

```{r, echo=FALSE}
summary(ibtracs_wp_results$intense_fit)
```


Using this model the number of a major storm in 1980 was `r predict(ibtracs_wp_results$intense_fit, type="response")[1]` and `r predict(ibtracs_wp_results$intense_fit, type="response")[45]` in 2024.

That is a change of `r (predict(ibtracs_wp_results$intense_fit, type="response")[45]-predict(ibtracs_wp_results$intense_fit, type="response")[1])/45*10` storms per decade.

#### Logistic Regression on Proportions of Major

```{r, echo=FALSE}
summary(ibtracs_wp_results$major_logit_fit)
```

We see a significant result with a positive coefficient, so the log-odds of major storms has increased. This tells us the proportion of major storms has increased.

Specifically, using this model the probability of a major storm in 1980 was `r predict(ibtracs_wp_results$major_logit_fit, type="response")[1]` and `r predict(ibtracs_wp_results$major_logit_fit, type="response")[45]` in 2024.

That is a change of `r (predict(ibtracs_wp_results$major_logit_fit, type="response")[45]-predict(ibtracs_wp_results$major_logit_fit, type="response")[1])/45*10*100` percent per decade.

#### Logistic Regression on Proportions of Intense

```{r, echo=FALSE}
summary(ibtracs_wp_results$intense_logit_fit)
```


Specifically, using this model the probability of a major storm in 1980 was `r predict(ibtracs_wp_results$intense_logit_fit, type="response")[1]` and `r predict(ibtracs_wp_results$intense_logit_fit, type="response")[45]` in 2024.

That is a change of `r (predict(ibtracs_wp_results$intense_logit_fit, type="response")[45]-predict(ibtracs_wp_results$intense_logit_fit, type="response")[1])/45*10*100` percent per decade.

### Change Point Methods

#### Change in Non-Shorty Counts

```{r, echo=FALSE}
ibtracs_wp_results$pois_chpt_test
```

The dates suggest the changepoint would occur from `r 1979+ibtracs_wp_results$pois_chpt_test[2]` to `r 1979+ibtracs_wp_results$pois_chpt_test[2]+1` with $p$-value `r ibtracs_wp_results$pois_chpt_test[3]`.

In regime one, the mean number of total storms is `r mean(ibtracs_wp_results$data$Total_Storms[1:ibtracs_wp_results$pois_chpt_test[2]])` and in regime two, the mean number of storms is `r mean(ibtracs_wp_results$data$Total_Storms[(ibtracs_wp_results$pois_chpt_test[2]+1):45])`; a change of `r mean(ibtracs_wp_results$data$Total_Storms[(ibtracs_wp_results$pois_chpt_test[2]+1):45]) - mean(ibtracs_wp_results$data$Total_Storms[1:ibtracs_wp_results$pois_chpt_test[2]])`.

#### Change in counts of major storms

```{r, echo=FALSE}
ibtracs_wp_results$major_pois_chpt_test
```

The dates suggest the changepoint would occur from `r 1979+ibtracs_wp_results$major_pois_chpt_test[2]` to `r 1979+ibtracs_wp_results$major_pois_chpt_test[2]+1` with $p$-value `r ibtracs_wp_results$major_pois_chpt_test[3]`.

In regime one, the mean number of major storms is `r mean(ibtracs_wp_results$data$Major_Storms[1:ibtracs_wp_results$major_pois_chpt_test[2]])` and in regime two, the mean number of storms is `r mean(ibtracs_wp_results$data$Major_Storms[(ibtracs_wp_results$major_pois_chpt_test[2]+1):45])`; a change of `r mean(ibtracs_wp_results$data$Major_Storms[(ibtracs_wp_results$major_pois_chpt_test[2]+1):45]) - mean(ibtracs_wp_results$data$Major_Storms[1:ibtracs_wp_results$major_pois_chpt_test[2]])`.

#### Change in counts of intense storms

```{r, echo=FALSE}
ibtracs_wp_results$intense_pois_chpt_test
```


The dates suggest the changepoint would occur from `r 1979+ibtracs_wp_results$intense_pois_chpt_test[2]` to `r 1979+ibtracs_wp_results$intense_pois_chpt_test[2]+1` with $p$-value `r ibtracs_wp_results$intense_pois_chpt_test[3]`.

In regime one, the mean number of intense storms is `r mean(ibtracs_wp_results$data$Intense_Storms[1:ibtracs_wp_results$intense_pois_chpt_test[2]])` and in regime two, the mean number of storms is `r mean(ibtracs_wp_results$data$Intense_Storms[(ibtracs_wp_results$intense_pois_chpt_test[2]+1):45])`; a change of `r mean(ibtracs_wp_results$data$Intense_Storms[(ibtracs_wp_results$intense_pois_chpt_test[2]+1):45]) - mean(ibtracs_wp_results$data$Intense_Storms[1:ibtracs_wp_results$intense_pois_chpt_test[2]])`.

#### Proportion of major storms

A change in the proportion of major storms?
  
```{r, echo=FALSE}
ibtracs_wp_results$prop_major_chpt_test
```

The change in proportions occurs `r 1979+ibtracs_wp_results$prop_major_chpt_test[2]` to `r 1979+ibtracs_wp_results$prop_major_chpt_test[2]+1` with $p$-value `r ibtracs_wp_results$prop_major_chpt_test[3]`. 

In regime one, the proportion of major storms is `r sum(ibtracs_wp_results$data$Major_Storms[1:ibtracs_wp_results$prop_major_chpt_test[2]])/sum(ibtracs_wp_results$data$Total_Storms[1:ibtracs_wp_results$prop_major_chpt_test[2]])` and in regime two, the proportion of major storms is `r sum(ibtracs_wp_results$data$Major_Storms[(ibtracs_wp_results$prop_major_chpt_test[2]+1):45])/sum(ibtracs_wp_results$data$Total_Storms[(ibtracs_wp_results$prop_major_chpt_test[2]+1):45])`; a change of `r sum(ibtracs_wp_results$data$Major_Storms[(ibtracs_wp_results$prop_major_chpt_test[2]+1):45])/sum(ibtracs_wp_results$data$Total_Storms[(ibtracs_wp_results$prop_major_chpt_test[2]+1):45]) - sum(ibtracs_wp_results$data$Major_Storms[1:ibtracs_wp_results$prop_major_chpt_test[2]])/sum(ibtracs_wp_results$data$Total_Storms[1:ibtracs_wp_results$prop_major_chpt_test[2]])`.

#### Proportion of intense storms

A change in the proportion of major storms?
  
```{r, echo=FALSE}
ibtracs_wp_results$prop_intense_chpt_test
```

The change in proportions occurs `r 1979+ibtracs_wp_results$prop_intense_chpt_test[2]` to `r 1979+ibtracs_wp_results$prop_intense_chpt_test[2]+1` with $p$-value `r ibtracs_wp_results$prop_intense_chpt_test[3]`.


In regime one, the proportion of major storms is `r sum(ibtracs_wp_results$data$Intense_Storms[1:ibtracs_wp_results$prop_intense_chpt_test[2]])/sum(ibtracs_wp_results$data$Total_Storms[1:ibtracs_wp_results$prop_intense_chpt_test[2]])` and in regime two, the proportion of major storms is `r sum(ibtracs_wp_results$data$Intense_Storms[(ibtracs_wp_results$prop_intense_chpt_test[2]+1):45])/sum(ibtracs_wp_results$data$Total_Storms[(ibtracs_wp_results$prop_intense_chpt_test[2]+1):45])`; a change of `r sum(ibtracs_wp_results$data$Intense_Storms[(ibtracs_wp_results$prop_intense_chpt_test[2]+1):45])/sum(ibtracs_wp_results$data$Total_Storms[(ibtracs_wp_results$prop_intense_chpt_test[2]+1):45]) - sum(ibtracs_wp_results$data$Intense_Storms[1:ibtracs_wp_results$prop_intense_chpt_test[2]])/sum(ibtracs_wp_results$data$Total_Storms[1:ibtracs_wp_results$prop_intense_chpt_test[2]])`.

### Overall Change Point Findings

Basically, the total number of storms is decreasing while the number of major storms aappears constant. So we see a change in the proportion of major storms.


```{r, echo=FALSE}
bind_rows(
  ibtracs_all_basins |>
    dplyr::filter(BASIN=="Western North Pacific") |>
    summarize(`Total Storms (avg)` = mean(Total_Storms),
              `Total Storms (stddev)` = sd(Total_Storms) ) |>
    mutate(Regime = "Entire Record"),
  ibtracs_all_basins |>
    dplyr::filter(BASIN=="Western North Pacific") |>
    mutate(Regime = ifelse(SEASON < 1998, "1980-1997", "1998-2024") ) |>
    group_by(Regime) |>
    summarize(`Total Storms (avg)` = mean(Total_Storms),
              `Total Storms (stddev)` = sd(Total_Storms) ) 
)|>
  dplyr::select(Regime, `Total Storms (avg)`, `Total Storms (stddev)`) |>
  knitr::kable()
```

We do see a decrease of about 5 or 6 storms per year circa 1998.

We do see a statistically meaningful change in proportion of major storms.

```{r, echo=FALSE}
bind_rows(
  ibtracs_all_basins |>
    dplyr::filter(BASIN=="Western North Pacific") |>
    summarize(`Total Storms (avg)` = mean(Total_Storms),
              `Major Storms (avg)` = mean(Major_Storms),
              `Proportion Major` = sum(Major_Storms)/sum(Total_Storms)) |>
    mutate(Regime = "Entire Record"),
  ibtracs_all_basins |>
    dplyr::filter(BASIN=="Western North Pacific") |>
    mutate(Regime = ifelse(SEASON < 1998, "1980-1997", "1998-2024") ) |>
    group_by(Regime) |>
    summarize(`Total Storms (avg)` = mean(Total_Storms),
              `Major Storms (avg)` = mean(Major_Storms),
              `Proportion Major` = sum(Major_Storms)/sum(Total_Storms)) ) |>
  dplyr::select(Regime, `Total Storms (avg)`, `Major Storms (avg)`, `Proportion Major`) |>
  knitr::kable()
```

The total number of storms decreased by about 20% while the number major storms has not really changed, thus the proportion of major storms has increased. 

```{r, echo=FALSE}
ibtracs_wp_results$plot_seg_fits
```



## Full HURSAT Record
  
We now repeat the analysis with the full HURSAT record. Note, this data runs from 1981 through 2017.

```{r, echo=FALSE}
hursat_wp_results <- get_report_basin(basin="Western North Pacific", all_basins=hursat_all_basins)
```

```{r, echo=FALSE}
hursat_wp_results$plot_data
```

### Summary Statistics (overall)

```{r, echo=FALSE}
hursat_wp_results$summary_stat |>
  knitr::kable()
```

### Regression Methods

#### Poisson Regression on non-Shorties

```{r, echo=FALSE}
summary(hursat_wp_results$pois_fit)
```

Note, this says a decrease of `r abs(coef(hursat_wp_results$pois_fit)[2])` log-storms per year. This corresponds to a `r (1-exp(coef(hursat_wp_results$pois_fit)[2]))*100` percent decrease per season. So over 37 years, that is a decrease of `r (1-exp(coef(hursat_wp_results$pois_fit)[2])^37)*100` percent.

Using this model, we would predict `r predict(hursat_wp_results$pois_fit, type="response")[1]` storms in 1981 and `r predict(hursat_wp_results$pois_fit, type="response")[37]` storms in 2017.

That is a change of `r (predict(hursat_wp_results$pois_fit, type="response")[37]-predict(hursat_wp_results$pois_fit, type="response")[1])/37*10` storms per decade.


#### Poisson Regression on Major Storms

```{r, echo=FALSE}
summary(hursat_wp_results$major_fit)
```


Using this model the number of a major storm in 1981 was `r predict(hursat_wp_results$major_fit, type="response")[1]` and `r predict(hursat_wp_results$major_fit, type="response")[37]` in 2017.

That is a change of `r (predict(hursat_wp_results$major_fit, type="response")[37]-predict(hursat_wp_results$major_fit, type="response")[1])/37*10` storms per decade.

#### Poisson Regression on Intense Storms

```{r, echo=FALSE}
summary(hursat_wp_results$intense_fit)
```

We see a significant increasing trend in the number of intense storms.

Using this model the number of a major storm in 1981 was `r predict(hursat_wp_results$intense_fit, type="response")[1]` and `r predict(hursat_wp_results$intense_fit, type="response")[37]` in 2017.

That is a change of `r (predict(hursat_wp_results$intense_fit, type="response")[37]-predict(hursat_wp_results$intense_fit, type="response")[1])/37*10` storms per decade.


#### Logistic Regression on Proportions

```{r, echo=FALSE}
summary(hursat_wp_results$major_logit_fit)
```


Specifically, using this model the probability of a major storm in 1981 was `r predict(hursat_wp_results$major_logit_fit, type="response")[1]` and `r predict(hursat_wp_results$major_logit_fit, type="response")[37]` in 2017.

That is a change of `r (predict(hursat_wp_results$major_logit_fit, type="response")[37]-predict(hursat_wp_results$major_logit_fit, type="response")[1])/37*10*100` percent per decade.


#### Logistic Regression on Proportions of Intense

```{r, echo=FALSE}
summary(hursat_wp_results$intense_logit_fit)
```


Specifically, using this model the probability of a major storm in 1981 was `r predict(hursat_wp_results$intense_logit_fit, type="response")[1]` and `r predict(hursat_wp_results$intense_logit_fit, type="response")[37]` in 2017.

That is a change of `r (predict(hursat_wp_results$intense_logit_fit, type="response")[37]-predict(hursat_wp_results$intense_logit_fit, type="response")[1])/37*10*100` percent per decade.


### Change Point Methods

#### Change in Non-Shorty Counts

```{r, echo=FALSE}
hursat_wp_results$pois_chpt_test
```

The dates suggest the changepoint would occur from `r 1980+hursat_wp_results$pois_chpt_test[2]` to `r 1980+hursat_wp_results$pois_chpt_test[2]+1` with $p$-value `r hursat_wp_results$pois_chpt_test[3]`.

In regime one, the mean number of total storms is `r mean(hursat_wp_results$data$Total_Storms[1:hursat_wp_results$pois_chpt_test[2]])` and in regime two, the mean number of storms is `r mean(hursat_wp_results$data$Total_Storms[(hursat_wp_results$pois_chpt_test[2]+1):37])`; a change of `r mean(hursat_wp_results$data$Total_Storms[(hursat_wp_results$pois_chpt_test[2]+1):37]) - mean(hursat_wp_results$data$Total_Storms[1:hursat_wp_results$pois_chpt_test[2]])`.

#### Change in counts of major storms

```{r, echo=FALSE}
hursat_wp_results$major_pois_chpt_test
```

The dates suggest the changepoint would occur from `r 1980+hursat_wp_results$major_pois_chpt_test[2]` to `r 1980+hursat_wp_results$major_pois_chpt_test[2]+1` with $p$-value `r hursat_wp_results$major_pois_chpt_test[3]`.

In regime one, the mean number of total storms is `r mean(hursat_wp_results$data$Major_Storms[1:hursat_wp_results$major_pois_chpt_test[2]])` and in regime two, the mean number of storms is `r mean(hursat_wp_results$data$Major_Storms[(hursat_wp_results$major_pois_chpt_test[2]+1):37])`; a change of `r mean(hursat_wp_results$data$Major_Storms[(hursat_wp_results$major_pois_chpt_test[2]+1):37]) - mean(hursat_wp_results$data$Major_Storms[1:hursat_wp_results$major_pois_chpt_test[2]])`.

#### Change in counts of intense storms

```{r, echo=FALSE}
hursat_wp_results$intense_pois_chpt_test
```

The dates suggest the changepoint would occur from `r 1980+hursat_wp_results$intense_pois_chpt_test[2]` to `r 1980+hursat_wp_results$intense_pois_chpt_test[2]+1` with $p$-value `r hursat_wp_results$intense_pois_chpt_test[3]`.

In regime one, the mean number of total storms is `r mean(hursat_wp_results$data$Intense_Storms[1:hursat_wp_results$intense_pois_chpt_test[2]])` and in regime two, the mean number of storms is `r mean(hursat_wp_results$data$Intense_Storms[(hursat_wp_results$intense_pois_chpt_test[2]+1):37])`; a change of `r mean(hursat_wp_results$data$Intense_Storms[(hursat_wp_results$intense_pois_chpt_test[2]+1):37]) - mean(hursat_wp_results$data$Intense_Storms[1:hursat_wp_results$intense_pois_chpt_test[2]])`.


#### Proportion of major storms

A change in the proportion of major storms?
  
```{r, echo=FALSE}
hursat_wp_results$prop_major_chpt_test
```

The change in proportions occurs `r 1979+hursat_wp_results$prop_major_chpt_test[2]` to `r 1979+hursat_wp_results$prop_major_chpt_test[2]+1` with $p$-value `r hursat_wp_results$prop_major_chpt_test[3]`. 

In regime one, the proportion of major storms is `r sum(hursat_wp_results$data$Major_Storms[1:hursat_wp_results$prop_major_chpt_test[2]])/sum(hursat_wp_results$data$Total_Storms[1:hursat_wp_results$prop_major_chpt_test[2]])` and in regime two, the proportion of major storms is `r sum(hursat_wp_results$data$Major_Storms[(hursat_wp_results$prop_major_chpt_test[2]+1):37])/sum(hursat_wp_results$data$Total_Storms[(hursat_wp_results$prop_major_chpt_test[2]+1):37])`; a change of `r sum(hursat_wp_results$data$Major_Storms[(hursat_wp_results$prop_major_chpt_test[2]+1):37])/sum(hursat_wp_results$data$Total_Storms[(hursat_wp_results$prop_major_chpt_test[2]+1):37]) - sum(hursat_wp_results$data$Major_Storms[1:hursat_wp_results$prop_major_chpt_test[2]])/sum(hursat_wp_results$data$Total_Storms[1:hursat_wp_results$prop_major_chpt_test[2]])`.



#### Proportion of intense storms

A change in the proportion of major storms?
  
```{r, echo=FALSE}
hursat_wp_results$prop_intense_chpt_test
```

The change in proportions occurs `r 1979+hursat_wp_results$prop_intense_chpt_test[2]` to `r 1979+hursat_wp_results$prop_intense_chpt_test[2]+1` with $p$-value `r hursat_wp_results$prop_intense_chpt_test[3]`.


In regime one, the proportion of major storms is `r sum(hursat_wp_results$data$Intense_Storms[1:hursat_wp_results$prop_intense_chpt_test[2]])/sum(hursat_wp_results$data$Total_Storms[1:hursat_wp_results$prop_intense_chpt_test[2]])` and in regime two, the proportion of major storms is `r sum(hursat_wp_results$data$Intense_Storms[(hursat_wp_results$prop_intense_chpt_test[2]+1):37])/sum(hursat_wp_results$data$Total_Storms[(hursat_wp_results$prop_intense_chpt_test[2]+1):37])`; a change of `r sum(hursat_wp_results$data$Intense_Storms[(hursat_wp_results$prop_intense_chpt_test[2]+1):37])/sum(hursat_wp_results$data$Total_Storms[(hursat_wp_results$prop_intense_chpt_test[2]+1):37]) - sum(hursat_wp_results$data$Intense_Storms[1:hursat_wp_results$prop_intense_chpt_test[2]])/sum(hursat_wp_results$data$Total_Storms[1:hursat_wp_results$prop_intense_chpt_test[2]])`.

### Overall Change Point Findings

The results suggest the total number of storms has decreased, with the number of major storms being fairly constant, so the proportion of major storms has increased as well.


```{r, echo=FALSE}
bind_rows(
  hursat_all_basins |>
    dplyr::filter(BASIN=="Western North Pacific") |>
    summarize(`Total Storms (avg)` = mean(Total_Storms),
              `Total Storms (stddev)` = sd(Total_Storms) ) |>
    mutate(Regime = "Entire Record"),
  hursat_all_basins |>
    dplyr::filter(BASIN=="Western North Pacific") |>
    mutate(Regime = ifelse(SEASON < 2003, "1981-2002", "2003-2017") ) |>
    group_by(Regime) |>
    summarize(`Total Storms (avg)` = mean(Total_Storms),
              `Total Storms (stddev)` = sd(Total_Storms) ) 
)|>
  dplyr::select(Regime, `Total Storms (avg)`, `Total Storms (stddev)`) |>
  knitr::kable()
```

A decrease of about 5 storms per year.


A table to compare the proportion of major storms. Note the change point time was a year different.

```{r, echo=FALSE}
bind_rows(
  hursat_all_basins |>
    dplyr::filter(BASIN=="Western North Pacific") |>
    summarize(`Total Storms (avg)` = mean(Total_Storms),
              `Major Storms (avg)` = mean(Major_Storms),
              `Proportion Major` = sum(Major_Storms)/sum(Total_Storms)) |>
    mutate(Regime = "Entire Record"),
  hursat_all_basins |>
    dplyr::filter(BASIN=="Western North Pacific") |>
    mutate(Regime = ifelse(SEASON < 2002, "1981-2001", "2002-2017") ) |>
    group_by(Regime) |>
    summarize(`Total Storms (avg)` = mean(Total_Storms),
              `Major Storms (avg)` = mean(Major_Storms),
              `Proportion Major` = sum(Major_Storms)/sum(Total_Storms)) ) |>
  dplyr::select(Regime, `Total Storms (avg)`, `Major Storms (avg)`, `Proportion Major`) |>
  knitr::kable()
```

Still a decrease of about 5 storms per year with a slight decrease-to-no change in major storms, thus the proportion went up!

```{r, echo=FALSE}
hursat_wp_results$plot_seg_fits
```

## Trimmed IBTRacs

We now repeat the analysis with the a version of the IBTRacs that has been trimmed to match HURSAT record. So this data runs from 1981 through 2017.

```{r, echo=FALSE}
trimmed_wp_results <- get_report_basin(basin="Western North Pacific", all_basins=ibtracs_trimmed)
```

```{r, echo=FALSE}
trimmed_wp_results$plot_data
```

### Summary Statistics (overall)

```{r, echo=FALSE}
trimmed_wp_results$summary_stat |>
  knitr::kable()
```

### Regression Methods

#### Poisson Regression

```{r, echo=FALSE}
summary(trimmed_wp_results$pois_fit)
```

Note, this says a decrease of `r abs(coef(trimmed_wp_results$pois_fit)[2])` log-storms per year. This corresponds to a `r (1-exp(coef(trimmed_wp_results$pois_fit)[2]))*100` percent decrease per season. So over 37 years, that is a decrease of `r (1-exp(coef(trimmed_wp_results$pois_fit)[2])^37)*100` percent.

Using this model, we would predict `r predict(trimmed_wp_results$pois_fit, type="response")[1]` storms in 1981 and `r predict(trimmed_wp_results$pois_fit, type="response")[37]` storms in 2017

That is a change of `r (predict(trimmed_wp_results$pois_fit, type="response")[37]-predict(trimmed_wp_results$pois_fit, type="response")[1])/37*10` storms per decade.

#### Poisson Regression on Major Storms

```{r, echo=FALSE}
summary(trimmed_wp_results$major_fit)
```

We see a non-significant trend in the number of major storms.

Using this model the number of a major storm in 1981 was `r predict(trimmed_wp_results$major_fit, type="response")[1]` and `r predict(trimmed_wp_results$major_fit, type="response")[37]` in 2017.

That is a change of `r (predict(trimmed_wp_results$major_fit, type="response")[37]-predict(trimmed_wp_results$major_fit, type="response")[1])/37*10` storms per decade.


#### Poisson Regression on Intense Storms

```{r, echo=FALSE}
summary(trimmed_wp_results$intense_fit)
```

Using this model the number of a major storm in 1981 was `r predict(trimmed_wp_results$intense_fit, type="response")[1]` and `r predict(trimmed_wp_results$intense_fit, type="response")[37]` in 2017

That is a change of `r (predict(trimmed_wp_results$intense_fit, type="response")[37]-predict(trimmed_wp_results$intense_fit, type="response")[1])/37*10` storms per decade.


#### Logistic Regression on Proportions

```{r, echo=FALSE}
summary(trimmed_wp_results$major_logit_fit)
```


Specifically, using this model the probability of a major storm in 1981 was `r predict(trimmed_wp_results$major_logit_fit, type="response")[1]` and `r predict(trimmed_wp_results$major_logit_fit, type="response")[37]` in 2017.


That is a change of `r (predict(trimmed_wp_results$major_logit_fit, type="response")[37]-predict(trimmed_wp_results$major_logit_fit, type="response")[1])/37*10*100` percent per decade.



#### Logistic Regression on Proportions of Intense

```{r, echo=FALSE}
summary(trimmed_wp_results$intense_logit_fit)
```


Specifically, using this model the probability of a major storm in 1981 was `r predict(trimmed_wp_results$intense_logit_fit, type="response")[1]` and `r predict(trimmed_wp_results$intense_logit_fit, type="response")[37]` in 2017.

That is a change of `r (predict(trimmed_wp_results$intense_logit_fit, type="response")[37]-predict(trimmed_wp_results$intense_logit_fit, type="response")[1])/37*10*100` percent per decade.


### Change Point Methods

#### Change in Non-Shorty Counts

```{r, echo=FALSE}
trimmed_wp_results$pois_chpt_test
```

We have evidence the number of non-shorties has changed. The dates suggest the changepoint would occur from `r 1980+trimmed_wp_results$pois_chpt_test[2]` to `r 1980+trimmed_wp_results$pois_chpt_test[2]+1` with $p$-value `r hursat_wp_results$pois_chpt_test[3]`.

In regime one, the mean number of total storms is `r mean(trimmed_wp_results$data$Total_Storms[1:trimmed_wp_results$pois_chpt_test[2]])` and in regime two, the mean number of storms is `r mean(trimmed_wp_results$data$Total_Storms[(trimmed_wp_results$pois_chpt_test[2]+1):37])`; a change of `r mean(trimmed_wp_results$data$Total_Storms[(trimmed_wp_results$pois_chpt_test[2]+1):37]) - mean(trimmed_wp_results$data$Total_Storms[1:trimmed_wp_results$pois_chpt_test[2]])`.

#### Change in counts of major storms

```{r, echo=FALSE}
trimmed_wp_results$major_pois_chpt_test
```


The dates suggest the changepoint would occur from `r 1980+trimmed_wp_results$major_pois_chpt_test[2]` to `r 1980+trimmed_wp_results$major_pois_chpt_test[2]+1` with $p$-value `r trimmed_wp_results$major_pois_chpt_test[3]`.

In regime one, the mean number of total storms is `r mean(trimmed_wp_results$data$Major_Storms[1:trimmed_wp_results$major_pois_chpt_test[2]])` and in regime two, the mean number of storms is `r mean(trimmed_wp_results$data$Major_Storms[(trimmed_wp_results$major_pois_chpt_test[2]+1):37])`; a change of `r mean(trimmed_wp_results$data$Major_Storms[(trimmed_wp_results$major_pois_chpt_test[2]+1):37]) - mean(trimmed_wp_results$data$Major_Storms[1:trimmed_wp_results$major_pois_chpt_test[2]])`.


#### Change in counts of intense storms

```{r, echo=FALSE}
trimmed_wp_results$intense_pois_chpt_test
```


The dates suggest the changepoint would occur from `r 1980+trimmed_wp_results$intense_pois_chpt_test[2]` to `r 1980+trimmed_wp_results$intense_pois_chpt_test[2]+1` with $p$-value `r trimmed_wp_results$intense_pois_chpt_test[3]`.

In regime one, the mean number of total storms is `r mean(trimmed_wp_results$data$Intense_Storms[1:trimmed_wp_results$intense_pois_chpt_test[2]])` and in regime two, the mean number of storms is `r mean(trimmed_wp_results$data$Intense_Storms[(trimmed_wp_results$intense_pois_chpt_test[2]+1):37])`; a change of `r mean(trimmed_wp_results$data$Intense_Storms[(trimmed_wp_results$intense_pois_chpt_test[2]+1):37]) - mean(trimmed_wp_results$data$Intense_Storms[1:trimmed_wp_results$intense_pois_chpt_test[2]])`.




#### Proportion of major storms

A change in the proportion of major storms?
  
```{r, echo=FALSE}
trimmed_wp_results$prop_major_chpt_test
```


The change in proportions occurs `r 1979+trimmed_wp_results$prop_major_chpt_test[2]` to `r 1979+trimmed_wp_results$prop_major_chpt_test[2]+1` with $p$-value `r trimmed_wp_results$prop_major_chpt_test[3]`. 

In regime one, the proportion of major storms is `r sum(trimmed_wp_results$data$Major_Storms[1:trimmed_wp_results$prop_major_chpt_test[2]])/sum(trimmed_wp_results$data$Total_Storms[1:trimmed_wp_results$prop_major_chpt_test[2]])` and in regime two, the proportion of major storms is `r sum(trimmed_wp_results$data$Major_Storms[(trimmed_wp_results$prop_major_chpt_test[2]+1):37])/sum(trimmed_wp_results$data$Total_Storms[(trimmed_wp_results$prop_major_chpt_test[2]+1):37])`; a change of `r sum(trimmed_wp_results$data$Major_Storms[(trimmed_wp_results$prop_major_chpt_test[2]+1):37])/sum(trimmed_wp_results$data$Total_Storms[(trimmed_wp_results$prop_major_chpt_test[2]+1):37]) - sum(trimmed_wp_results$data$Major_Storms[1:trimmed_wp_results$prop_major_chpt_test[2]])/sum(trimmed_wp_results$data$Total_Storms[1:trimmed_wp_results$prop_major_chpt_test[2]])`.


#### Proportion of intense storms

A change in the proportion of major storms?
  
```{r, echo=FALSE}
trimmed_wp_results$prop_intense_chpt_test
```

The change in proportions occurs `r 1979+trimmed_wp_results$prop_intense_chpt_test[2]` to `r 1979+trimmed_wp_results$prop_intense_chpt_test[2]+1` with $p$-value `r trimmed_wp_results$prop_intense_chpt_test[3]`.


In regime one, the proportion of major storms is `r sum(trimmed_wp_results$data$Intense_Storms[1:trimmed_wp_results$prop_intense_chpt_test[2]])/sum(trimmed_wp_results$data$Total_Storms[1:trimmed_wp_results$prop_intense_chpt_test[2]])` and in regime two, the proportion of major storms is `r sum(trimmed_wp_results$data$Intense_Storms[(trimmed_wp_results$prop_intense_chpt_test[2]+1):37])/sum(trimmed_wp_results$data$Total_Storms[(trimmed_wp_results$prop_intense_chpt_test[2]+1):37])`; a change of `r sum(trimmed_wp_results$data$Intense_Storms[(trimmed_wp_results$prop_intense_chpt_test[2]+1):37])/sum(trimmed_wp_results$data$Total_Storms[(trimmed_wp_results$prop_intense_chpt_test[2]+1):37]) - sum(trimmed_wp_results$data$Intense_Storms[1:trimmed_wp_results$prop_intense_chpt_test[2]])/sum(trimmed_wp_results$data$Total_Storms[1:trimmed_wp_results$prop_intense_chpt_test[2]])`.




### Overall Change Point Findings


Consistent results as the other two analyses -- the results suggest the total number of storms has decreased, while the number of major storms is constant, so the proportion of major storms has increased as well.


```{r, echo=FALSE}
bind_rows(
  ibtracs_trimmed |>
    dplyr::filter(BASIN=="Western North Pacific") |>
    summarize(`Total Storms (avg)` = mean(Total_Storms),
              `Total Storms (stddev)` = sd(Total_Storms) ) |>
    mutate(Regime = "Entire Record"),
  ibtracs_trimmed |>
    dplyr::filter(BASIN=="Western North Pacific") |>
    mutate(Regime = ifelse(SEASON < 1998, "1981-1997", "1998-2017") ) |>
    group_by(Regime) |>
    summarize(`Total Storms (avg)` = mean(Total_Storms),
              `Total Storms (stddev)` = sd(Total_Storms) ) 
)|>
  dplyr::select(Regime, `Total Storms (avg)`, `Total Storms (stddev)`) |>
  knitr::kable()
```

A decrease of about 5 storms per year (about 20% decrease)

A table to compare the proportion of major storms.

```{r, echo=FALSE}
bind_rows(
  ibtracs_trimmed |>
    dplyr::filter(BASIN=="Western North Pacific") |>
    summarize(`Total Storms (avg)` = mean(Total_Storms),
              `Major Storms (avg)` = mean(Major_Storms),
              `Proportion Major` = sum(Major_Storms)/sum(Total_Storms)) |>
    mutate(Regime = "Entire Record"),
  ibtracs_trimmed |>
    dplyr::filter(BASIN=="Western North Pacific") |>
    mutate(Regime = ifelse(SEASON < 1998, "1981-1997", "1998-2017") ) |>
    group_by(Regime) |>
    summarize(`Total Storms (avg)` = mean(Total_Storms),
              `Major Storms (avg)` = mean(Major_Storms),
              `Proportion Major` = sum(Major_Storms)/sum(Total_Storms)) ) |>
  dplyr::select(Regime, `Total Storms (avg)`, `Major Storms (avg)`, `Proportion Major`) |>
  knitr::kable()
```

Total number of storms decreased while major storms stayed the same, hence an increase in the proportion of major storms.

```{r, echo=FALSE}
trimmed_wp_results$plot_seg_fits
```
  



## Overall Hypothesis Test Results

### Total Number of storms (non-shorties)
  
**Poisson Regression**

```{r, echo=FALSE}
data.frame(Data=c("IBTracs 1980-2024", "HURSAT 1981-2017", "IBTracs 1981-2017"),
           Coefficent = c(coef(ibtracs_wp_results$pois_fit)[2],coef(hursat_wp_results$pois_fit)[2],coef(trimmed_wp_results$pois_fit)[2]),
           `p-value` = c(anova(ibtracs_wp_results$pois_fit)$`Pr(>Chi)`[2], anova(hursat_wp_results$pois_fit)$`Pr(>Chi)`[2], anova(trimmed_wp_results$pois_fit)$`Pr(>Chi)`[2]),
           `Start` = c(predict(ibtracs_wp_results$pois_fit, type="response")[1], predict(hursat_wp_results$pois_fit, type="response")[1], predict(trimmed_wp_results$pois_fit, type="response")[1] ),
           `End` = c(predict(ibtracs_wp_results$pois_fit, type="response")[length(predict(ibtracs_wp_results$pois_fit, type="response"))], predict(hursat_wp_results$pois_fit, type="response")[length(predict(hursat_wp_results$pois_fit, type="response"))], predict(trimmed_wp_results$pois_fit, type="response")[length(predict(trimmed_wp_results$pois_fit, type="response"))] ),
           `Change per Decade` = c( (predict(ibtracs_wp_results$pois_fit, type="response")[length(predict(ibtracs_wp_results$pois_fit, type="response"))]-predict(ibtracs_wp_results$pois_fit, type="response")[1])/length(predict(ibtracs_wp_results$pois_fit, type="response"))*10,
                                    (predict(hursat_wp_results$pois_fit, type="response")[length(predict(hursat_wp_results$pois_fit, type="response"))]-predict(hursat_wp_results$pois_fit, type="response")[1])/length(predict(hursat_wp_results$pois_fit, type="response"))*10,
                                    (predict(trimmed_wp_results$pois_fit, type="response")[length(predict(trimmed_wp_results$pois_fit, type="response"))]-predict(trimmed_wp_results$pois_fit, type="response")[1])/length(predict(trimmed_wp_results$pois_fit, type="response"))*10 ) ) |>
  knitr::kable(digits=5)
```

**SCUSUM Change Point Test**

```{r, echo=FALSE}
data.frame(Data=c("IBTracs 1980-2024", "HURSAT 1981-2017", "IBTracs 1981-2017"),
           `Chpt Location` = c(ibtracs_wp_results$pois_chpt_test[2]+1980,hursat_wp_results$pois_chpt_test[2]+1981,trimmed_wp_results$pois_chpt_test[2]+1981),
           `p-value` =c(ibtracs_wp_results$pois_chpt_test[3],hursat_wp_results$pois_chpt_test[3],trimmed_wp_results$pois_chpt_test[3]),
           `Regime 1` = c(mean(ibtracs_wp_results$data$Total_Storms[1:ibtracs_wp_results$pois_chpt_test[2]]),
                          mean(hursat_wp_results$data$Total_Storms[1:hursat_wp_results$pois_chpt_test[2]]),
                          mean(trimmed_wp_results$data$Total_Storms[1:trimmed_wp_results$pois_chpt_test[2]]) ),
           `Regime 2` = c(mean(ibtracs_wp_results$data$Total_Storms[(ibtracs_wp_results$pois_chpt_test[2]+1):length(ibtracs_wp_results$data$Total_Storms)]),
                          mean(hursat_wp_results$data$Total_Storms[(hursat_wp_results$pois_chpt_test[2]+1):length(hursat_wp_results$data$Total_Storms)]),
                            mean(trimmed_wp_results$data$Total_Storms[(trimmed_wp_results$pois_chpt_test[2]+1):length(trimmed_wp_results$data$Total_Storms)]) ),
           `Change` = c(mean(ibtracs_wp_results$data$Total_Storms[(ibtracs_wp_results$pois_chpt_test[2]+1):length(ibtracs_wp_results$data$Total_Storms)]) - mean(ibtracs_wp_results$data$Total_Storms[1:ibtracs_wp_results$pois_chpt_test[2]]),
                        mean(hursat_wp_results$data$Total_Storms[(hursat_wp_results$pois_chpt_test[2]+1):length(hursat_wp_results$data$Total_Storms)]) - mean(hursat_wp_results$data$Total_Storms[1:hursat_wp_results$pois_chpt_test[2]]),
                        mean(trimmed_wp_results$data$Total_Storms[(trimmed_wp_results$pois_chpt_test[2]+1):length(trimmed_wp_results$data$Total_Storms)]) - mean(trimmed_wp_results$data$Total_Storms[1:trimmed_wp_results$pois_chpt_test[2]]) ) ) |>
  knitr::kable(digits=5)
```

### Major Storm Counts

**Poisson Regression**

```{r, echo=FALSE}
data.frame(Data=c("IBTracs 1980-2024", "HURSAT 1981-2017", "IBTracs 1981-2017"),
           Coefficent = c(coef(ibtracs_wp_results$major_fit)[2],
                          coef(hursat_wp_results$major_fit)[2],
                          coef(trimmed_wp_results$major_fit)[2]),
           `p-value` = c(anova(ibtracs_wp_results$major_fit)$`Pr(>Chi)`[2],
                         anova(hursat_wp_results$major_fit)$`Pr(>Chi)`[2],
                         anova(trimmed_wp_results$major_fit)$`Pr(>Chi)`[2]),
           `Start` = c(predict(ibtracs_wp_results$major_fit, type="response")[1], 
                       predict(hursat_wp_results$major_fit, type="response")[1], 
                       predict(trimmed_wp_results$major_fit, type="response")[1] ),
           `End` = c(predict(ibtracs_wp_results$major_fit,type="response")[length(predict(ibtracs_wp_results$major_fit, type="response"))],
                     predict(hursat_wp_results$major_fit, type="response")[length(predict(hursat_wp_results$major_fit, type="response"))], 
                     predict(trimmed_wp_results$major_fit, type="response")[length(predict(trimmed_wp_results$major_fit, type="response"))] ),
           `Change per Decade` = c( (predict(ibtracs_wp_results$major_fit, type="response")[length(predict(ibtracs_wp_results$major_fit, type="response"))]-predict(ibtracs_wp_results$major_fit, type="response")[1])/length(predict(ibtracs_wp_results$major_fit, type="response"))*10,
                                    (predict(hursat_wp_results$major_fit, type="response")[length(predict(hursat_wp_results$major_fit, type="response"))]-predict(hursat_wp_results$major_fit, type="response")[1])/length(predict(hursat_wp_results$major_fit, type="response"))*10,
                                    (predict(trimmed_wp_results$major_fit, type="response")[length(predict(trimmed_wp_results$major_fit, type="response"))]-predict(trimmed_wp_results$major_fit, type="response")[1])/length(predict(trimmed_wp_results$major_fit, type="response"))*10 ) ) |>
  knitr::kable(digits=5)
```

**SCUSUM Change Point Test**

```{r, echo=FALSE}
data.frame(Data=c("IBTracs 1980-2024", "HURSAT 1981-2017", "IBTracs 1981-2017"),
           `Chpt Location` = c(ibtracs_wp_results$major_pois_chpt_test[2]+1980,hursat_wp_results$major_pois_chpt_test[2]+1981,trimmed_wp_results$major_pois_chpt_test[2]+1981),
           `p-value` =c(ibtracs_wp_results$major_pois_chpt_test[3],hursat_wp_results$major_pois_chpt_test[3],trimmed_wp_results$major_pois_chpt_test[3]),
           `Regime 1` = c(mean(ibtracs_wp_results$data$Major_Storms[1:ibtracs_wp_results$major_pois_chpt_test[2]]),
                          mean(hursat_wp_results$data$Major_Storms[1:hursat_wp_results$major_pois_chpt_test[2]]),
                          mean(trimmed_wp_results$data$Major_Storms[1:trimmed_wp_results$major_pois_chpt_test[2]]) ),
           `Regime 2` = c(mean(ibtracs_wp_results$data$Major_Storms[(ibtracs_wp_results$major_pois_chpt_test[2]+1):length(ibtracs_wp_results$data$Major_Storms)]),
                          mean(hursat_wp_results$data$Major_Storms[(hursat_wp_results$major_pois_chpt_test[2]+1):length(hursat_wp_results$data$Major_Storms)]),
                            mean(trimmed_wp_results$data$Major_Storms[(trimmed_wp_results$major_pois_chpt_test[2]+1):length(trimmed_wp_results$data$Major_Storms)]) ),
           `Change` = c(mean(ibtracs_wp_results$data$Major_Storms[(ibtracs_wp_results$major_pois_chpt_test[2]+1):length(ibtracs_wp_results$data$Major_Storms)]) - mean(ibtracs_wp_results$data$Major_Storms[1:ibtracs_wp_results$major_pois_chpt_test[2]]),
                        mean(hursat_wp_results$data$Major_Storms[(hursat_wp_results$major_pois_chpt_test[2]+1):length(hursat_wp_results$data$Major_Storms)]) - mean(hursat_wp_results$data$Major_Storms[1:hursat_wp_results$major_pois_chpt_test[2]]),
                        mean(trimmed_wp_results$data$Major_Storms[(trimmed_wp_results$major_pois_chpt_test[2]+1):length(trimmed_wp_results$data$Major_Storms)]) - mean(trimmed_wp_results$data$Major_Storms[1:trimmed_wp_results$major_pois_chpt_test[2]]) ) ) |>
  knitr::kable(digits=5)
```

### Proportion of Major Storms

**Logistic Regression**

```{r, echo=FALSE}
data.frame(Data=c("IBTracs 1980-2024", "HURSAT 1981-2017", "IBTracs 1981-2017"),
           Coefficent = c(coef(ibtracs_wp_results$major_logit_fit)[2],coef(hursat_wp_results$major_logit_fit)[2],coef(trimmed_wp_results$major_logit_fit)[2]),
           `p-value` = c(anova(ibtracs_wp_results$major_logit_fit)$`Pr(>Chi)`[2], anova(hursat_wp_results$major_logit_fit)$`Pr(>Chi)`[2], anova(trimmed_wp_results$major_logit_fit)$`Pr(>Chi)`[2]),
                      `Start` = c(predict(ibtracs_wp_results$major_logit_fit, type="response")[1]*100, 
                       predict(hursat_wp_results$major_logit_fit, type="response")[1]*100, 
                       predict(trimmed_wp_results$major_logit_fit, type="response")[1]*100 ),
           `End` = c(predict(ibtracs_wp_results$major_logit_fit,type="response")[length(predict(ibtracs_wp_results$major_fit, type="response"))]*100,
                     predict(hursat_wp_results$major_logit_fit, type="response")[length(predict(hursat_wp_results$major_logit_fit, type="response"))]*100, 
                     predict(trimmed_wp_results$major_logit_fit, type="response")[length(predict(trimmed_wp_results$major_logit_fit, type="response"))]*100 ),
           `Change per Decade` = c( (predict(ibtracs_wp_results$major_logit_fit, type="response")[length(predict(ibtracs_wp_results$major_logit_fit, type="response"))]-predict(ibtracs_wp_results$major_logit_fit, type="response")[1])/length(predict(ibtracs_wp_results$major_logit_fit, type="response"))*10*100,
                                    (predict(hursat_wp_results$major_logit_fit, type="response")[length(predict(hursat_wp_results$major_logit_fit, type="response"))]-predict(hursat_wp_results$major_logit_fit, type="response")[1])/length(predict(hursat_wp_results$major_logit_fit, type="response"))*10*100,
                                    (predict(trimmed_wp_results$major_logit_fit, type="response")[length(predict(trimmed_wp_results$major_logit_fit, type="response"))]-predict(trimmed_wp_results$major_logit_fit, type="response")[1])/length(predict(trimmed_wp_results$major_logit_fit, type="response"))*10*100 ) ) |>
  knitr::kable()
```

**SCUSUM Change Point Test**

```{r, echo=FALSE}
data.frame(Data=c("IBTracs 1980-2024", "HURSAT 1981-2017", "IBTracs 1981-2017"),
           `Chpt Location` = c(ibtracs_wp_results$prop_major_chpt_test[2]+1980,hursat_wp_results$prop_major_chpt_test[2]+1981,trimmed_wp_results$prop_major_chpt_test[2]+1981),
           `p-value` =c(ibtracs_wp_results$prop_major_chpt_test[3],hursat_wp_results$prop_major_chpt_test[3],trimmed_wp_results$prop_major_chpt_test[3]),
           `Regime 1 %` = c(sum(ibtracs_wp_results$data$Major_Storms[1:ibtracs_wp_results$prop_major_chpt_test[2]])/sum(ibtracs_wp_results$data$Total_Storms[1:ibtracs_wp_results$prop_major_chpt_test[2]])*100,
                            sum(hursat_wp_results$data$Major_Storms[1:hursat_wp_results$prop_major_chpt_test[2]])/sum(hursat_wp_results$data$Total_Storms[1:hursat_wp_results$prop_major_chpt_test[2]])*100,
                            sum(ibtracs_wp_results$data$Major_Storms[1:ibtracs_wp_results$prop_major_chpt_test[2]])/sum(ibtracs_wp_results$data$Total_Storms[1:ibtracs_wp_results$prop_major_chpt_test[2]])*100 ),
           `Regime 2 %` = c(sum(ibtracs_wp_results$data$Major_Storms[(ibtracs_wp_results$prop_major_chpt_test[2]+1):length(ibtracs_wp_results$data$Major_Storms)])/sum(ibtracs_wp_results$data$Total_Storms[(ibtracs_wp_results$prop_major_chpt_test[2]+1):length(ibtracs_wp_results$data$Major_Storms)])*100,
                            sum(hursat_wp_results$data$Major_Storms[(hursat_wp_results$prop_major_chpt_test[2]+1):length(hursat_wp_results$data$Major_Storms)])/sum(hursat_wp_results$data$Total_Storms[(hursat_wp_results$prop_major_chpt_test[2]+1):length(hursat_wp_results$data$Major_Storms)])*100,
                            sum(trimmed_wp_results$data$Major_Storms[(trimmed_wp_results$prop_major_chpt_test[2]+1):length(trimmed_wp_results$data$Major_Storms)])/sum(trimmed_wp_results$data$Total_Storms[(trimmed_wp_results$prop_major_chpt_test[2]+1):length(trimmed_wp_results$data$Major_Storms)])*100 ),
           `Change %` = c((sum(ibtracs_wp_results$data$Major_Storms[(ibtracs_wp_results$prop_major_chpt_test[2]+1):length(ibtracs_wp_results$data$Major_Storms)])/sum(ibtracs_wp_results$data$Total_Storms[(ibtracs_wp_results$prop_major_chpt_test[2]+1):length(ibtracs_wp_results$data$Major_Storms)]) - sum(ibtracs_wp_results$data$Major_Storms[1:ibtracs_wp_results$prop_major_chpt_test[2]])/sum(ibtracs_wp_results$data$Total_Storms[1:ibtracs_wp_results$prop_major_chpt_test[2]]) )*100,
                          (sum(hursat_wp_results$data$Major_Storms[(hursat_wp_results$prop_major_chpt_test[2]+1):length(hursat_wp_results$data$Major_Storms)])/sum(hursat_wp_results$data$Total_Storms[(hursat_wp_results$prop_major_chpt_test[2]+1):length(hursat_wp_results$data$Major_Storms)]) - sum(hursat_wp_results$data$Major_Storms[1:hursat_wp_results$prop_major_chpt_test[2]])/sum(hursat_wp_results$data$Total_Storms[1:hursat_wp_results$prop_major_chpt_test[2]]) )*100,
                          (sum(trimmed_wp_results$data$Major_Storms[(trimmed_wp_results$prop_major_chpt_test[2]+1):length(trimmed_wp_results$data$Major_Storms)])/sum(trimmed_wp_results$data$Total_Storms[(trimmed_wp_results$prop_major_chpt_test[2]+1):length(trimmed_wp_results$data$Major_Storms)]) - sum(trimmed_wp_results$data$Major_Storms[1:trimmed_wp_results$prop_major_chpt_test[2]])/sum(trimmed_wp_results$data$Total_Storms[1:trimmed_wp_results$prop_major_chpt_test[2]]) )*100 ) ) |>
  knitr::kable()
```



### Intense Storm Counts

**Poisson Regression**

```{r, echo=FALSE}
data.frame(Data=c("IBTracs 1980-2024", "HURSAT 1981-2017", "IBTracs 1981-2017"),
           Coefficent = c(coef(ibtracs_wp_results$intense_fit)[2],
                          coef(hursat_wp_results$intense_fit)[2],
                          coef(trimmed_wp_results$intense_fit)[2]),
           `p-value` = c(anova(ibtracs_wp_results$intense_fit)$`Pr(>Chi)`[2],
                         anova(hursat_wp_results$intense_fit)$`Pr(>Chi)`[2],
                         anova(trimmed_wp_results$intense_fit)$`Pr(>Chi)`[2]),
           `Start` = c(predict(ibtracs_wp_results$intense_fit, type="response")[1], 
                       predict(hursat_wp_results$intense_fit, type="response")[1], 
                       predict(trimmed_wp_results$intense_fit, type="response")[1] ),
           `End` = c(predict(ibtracs_wp_results$intense_fit,type="response")[length(predict(ibtracs_wp_results$intense_fit, type="response"))],
                     predict(hursat_wp_results$intense_fit, type="response")[length(predict(hursat_wp_results$intense_fit, type="response"))], 
                     predict(trimmed_wp_results$intense_fit, type="response")[length(predict(trimmed_wp_results$intense_fit, type="response"))] ),
           `Change per Decade` = c( (predict(ibtracs_wp_results$intense_fit, type="response")[length(predict(ibtracs_wp_results$intense_fit, type="response"))]-predict(ibtracs_wp_results$intense_fit, type="response")[1])/length(predict(ibtracs_wp_results$intense_fit, type="response"))*10,
                                    (predict(hursat_wp_results$intense_fit, type="response")[length(predict(hursat_wp_results$intense_fit, type="response"))]-predict(hursat_wp_results$intense_fit, type="response")[1])/length(predict(hursat_wp_results$intense_fit, type="response"))*10,
                                    (predict(trimmed_wp_results$intense_fit, type="response")[length(predict(trimmed_wp_results$intense_fit, type="response"))]-predict(trimmed_wp_results$intense_fit, type="response")[1])/length(predict(trimmed_wp_results$intense_fit, type="response"))*10 ) ) |>
  knitr::kable(digits=5)
```

**SCUSUM Change Point Test**

```{r, echo=FALSE}
data.frame(Data=c("IBTracs 1980-2024", "HURSAT 1981-2017", "IBTracs 1981-2017"),
           `Chpt Location` = c(ibtracs_wp_results$intense_pois_chpt_test[2]+1980,hursat_wp_results$intense_pois_chpt_test[2]+1981,trimmed_wp_results$intense_pois_chpt_test[2]+1981),
           `p-value` =c(ibtracs_wp_results$intense_pois_chpt_test[3],hursat_wp_results$intense_pois_chpt_test[3],trimmed_wp_results$intense_pois_chpt_test[3]),
           `Regime 1` = c(mean(ibtracs_wp_results$data$Intense_Storms[1:ibtracs_wp_results$intense_pois_chpt_test[2]]),
                          mean(hursat_wp_results$data$Intense_Storms[1:hursat_wp_results$intense_pois_chpt_test[2]]),
                          mean(trimmed_wp_results$data$Intense_Storms[1:trimmed_wp_results$intense_pois_chpt_test[2]]) ),
           `Regime 2` = c(mean(ibtracs_wp_results$data$Intense_Storms[(ibtracs_wp_results$intense_pois_chpt_test[2]+1):length(ibtracs_wp_results$data$Intense_Storms)]),
                          mean(hursat_wp_results$data$Intense_Storms[(hursat_wp_results$intense_pois_chpt_test[2]+1):length(hursat_wp_results$data$Intense_Storms)]),
                            mean(trimmed_wp_results$data$Intense_Storms[(trimmed_wp_results$intense_pois_chpt_test[2]+1):length(trimmed_wp_results$data$Intense_Storms)]) ),
           `Change` = c(mean(ibtracs_wp_results$data$Intense_Storms[(ibtracs_wp_results$intense_pois_chpt_test[2]+1):length(ibtracs_wp_results$data$Intense_Storms)]) - mean(ibtracs_wp_results$data$Intense_Storms[1:ibtracs_wp_results$intense_pois_chpt_test[2]]),
                        mean(hursat_wp_results$data$Intense_Storms[(hursat_wp_results$intense_pois_chpt_test[2]+1):length(hursat_wp_results$data$Intense_Storms)]) - mean(hursat_wp_results$data$Intense_Storms[1:hursat_wp_results$intense_pois_chpt_test[2]]),
                        mean(trimmed_wp_results$data$Intense_Storms[(trimmed_wp_results$intense_pois_chpt_test[2]+1):length(trimmed_wp_results$data$Intense_Storms)]) - mean(trimmed_wp_results$data$Intense_Storms[1:trimmed_wp_results$intense_pois_chpt_test[2]]) ) ) |>
  knitr::kable(digits=5)
```

### Proportion of Intense Storms

**Logistic Regression**

```{r, echo=FALSE}
data.frame(Data=c("IBTracs 1980-2024", "HURSAT 1981-2017", "IBTracs 1981-2017"),
           Coefficent = c(coef(ibtracs_wp_results$intense_logit_fit)[2],coef(hursat_wp_results$intense_logit_fit)[2],coef(trimmed_wp_results$intense_logit_fit)[2]),
           `p-value` = c(anova(ibtracs_wp_results$intense_logit_fit)$`Pr(>Chi)`[2], anova(hursat_wp_results$intense_logit_fit)$`Pr(>Chi)`[2], anova(trimmed_wp_results$intense_logit_fit)$`Pr(>Chi)`[2]),
                      `Start` = c(predict(ibtracs_wp_results$intense_logit_fit, type="response")[1]*100, 
                       predict(hursat_wp_results$intense_logit_fit, type="response")[1]*100, 
                       predict(trimmed_wp_results$intense_logit_fit, type="response")[1]*100 ),
           `End` = c(predict(ibtracs_wp_results$intense_logit_fit,type="response")[length(predict(ibtracs_wp_results$intense_fit, type="response"))]*100,
                     predict(hursat_wp_results$intense_logit_fit, type="response")[length(predict(hursat_wp_results$intense_logit_fit, type="response"))]*100, 
                     predict(trimmed_wp_results$intense_logit_fit, type="response")[length(predict(trimmed_wp_results$intense_logit_fit, type="response"))]*100 ),
           `Change per Decade` = c( (predict(ibtracs_wp_results$intense_logit_fit, type="response")[length(predict(ibtracs_wp_results$intense_logit_fit, type="response"))]-predict(ibtracs_wp_results$intense_logit_fit, type="response")[1])/length(predict(ibtracs_wp_results$intense_logit_fit, type="response"))*10*100,
                                    (predict(hursat_wp_results$intense_logit_fit, type="response")[length(predict(hursat_wp_results$intense_logit_fit, type="response"))]-predict(hursat_wp_results$intense_logit_fit, type="response")[1])/length(predict(hursat_wp_results$intense_logit_fit, type="response"))*10*100,
                                    (predict(trimmed_wp_results$intense_logit_fit, type="response")[length(predict(trimmed_wp_results$intense_logit_fit, type="response"))]-predict(trimmed_wp_results$intense_logit_fit, type="response")[1])/length(predict(trimmed_wp_results$intense_logit_fit, type="response"))*10*100 ) ) |>
  knitr::kable()
```

**SCUSUM Change Point Test**

```{r, echo=FALSE}
data.frame(Data=c("IBTracs 1980-2024", "HURSAT 1981-2017", "IBTracs 1981-2017"),
           `Chpt Location` = c(ibtracs_wp_results$prop_intense_chpt_test[2]+1980,hursat_wp_results$prop_intense_chpt_test[2]+1981,trimmed_wp_results$prop_intense_chpt_test[2]+1981),
           `p-value` =c(ibtracs_wp_results$prop_intense_chpt_test[3],hursat_wp_results$prop_intense_chpt_test[3],trimmed_wp_results$prop_intense_chpt_test[3]),
           `Regime 1 %` = c(sum(ibtracs_wp_results$data$Intense_Storms[1:ibtracs_wp_results$prop_intense_chpt_test[2]])/sum(ibtracs_wp_results$data$Total_Storms[1:ibtracs_wp_results$prop_intense_chpt_test[2]])*100,
                          sum(hursat_wp_results$data$Intense_Storms[1:hursat_wp_results$prop_intense_chpt_test[2]])/sum(hursat_wp_results$data$Total_Storms[1:hursat_wp_results$prop_intense_chpt_test[2]])*100,
                          sum(ibtracs_wp_results$data$Intense_Storms[1:ibtracs_wp_results$prop_intense_chpt_test[2]])/sum(ibtracs_wp_results$data$Total_Storms[1:ibtracs_wp_results$prop_intense_chpt_test[2]])*100 ),
           `Regime 2 %` = c(sum(ibtracs_wp_results$data$Intense_Storms[(ibtracs_wp_results$prop_intense_chpt_test[2]+1):length(ibtracs_wp_results$data$Intense_Storms)])/sum(ibtracs_wp_results$data$Total_Storms[(ibtracs_wp_results$prop_intense_chpt_test[2]+1):length(ibtracs_wp_results$data$Intense_Storms)])*100,
                          sum(hursat_wp_results$data$Intense_Storms[(hursat_wp_results$prop_intense_chpt_test[2]+1):length(hursat_wp_results$data$Intense_Storms)])/sum(hursat_wp_results$data$Total_Storms[(hursat_wp_results$prop_intense_chpt_test[2]+1):length(hursat_wp_results$data$Intense_Storms)])*100,
                          sum(trimmed_wp_results$data$Intense_Storms[(trimmed_wp_results$prop_intense_chpt_test[2]+1):length(trimmed_wp_results$data$Intense_Storms)])/sum(trimmed_wp_results$data$Total_Storms[(trimmed_wp_results$prop_intense_chpt_test[2]+1):length(trimmed_wp_results$data$Intense_Storms)])*100 ),
           `Change %` = c((sum(ibtracs_wp_results$data$Intense_Storms[(ibtracs_wp_results$prop_intense_chpt_test[2]+1):length(ibtracs_wp_results$data$Intense_Storms)])/sum(ibtracs_wp_results$data$Total_Storms[(ibtracs_wp_results$prop_intense_chpt_test[2]+1):length(ibtracs_wp_results$data$Intense_Storms)]) - sum(ibtracs_wp_results$data$Intense_Storms[1:ibtracs_wp_results$prop_intense_chpt_test[2]])/sum(ibtracs_wp_results$data$Total_Storms[1:ibtracs_wp_results$prop_intense_chpt_test[2]]) )*100,
                        (sum(hursat_wp_results$data$Intense_Storms[(hursat_wp_results$prop_intense_chpt_test[2]+1):length(hursat_wp_results$data$Intense_Storms)])/sum(hursat_wp_results$data$Total_Storms[(hursat_wp_results$prop_intense_chpt_test[2]+1):length(hursat_wp_results$data$Intense_Storms)]) - sum(hursat_wp_results$data$Intense_Storms[1:hursat_wp_results$prop_intense_chpt_test[2]])/sum(hursat_wp_results$data$Total_Storms[1:hursat_wp_results$prop_intense_chpt_test[2]]) )*100,
                        (sum(trimmed_wp_results$data$Intense_Storms[(trimmed_wp_results$prop_intense_chpt_test[2]+1):length(trimmed_wp_results$data$Intense_Storms)])/sum(trimmed_wp_results$data$Total_Storms[(trimmed_wp_results$prop_intense_chpt_test[2]+1):length(trimmed_wp_results$data$Intense_Storms)]) - sum(trimmed_wp_results$data$Intense_Storms[1:trimmed_wp_results$prop_intense_chpt_test[2]])/sum(trimmed_wp_results$data$Total_Storms[1:trimmed_wp_results$prop_intense_chpt_test[2]]) )*100 ) ) |>
  knitr::kable()
```

