---
title: "Tropical Cyclone Analysis"
subtitle: "Single changepoint & Regression Methods"
author: "Tom Fisher"
date: today
format:
  html:
    page-layout: full
    toc: true
    toc-depth: 3
    toc-location: left
    embed-resources: true
editor: visual
bibliography: sn-bibliography.bib
---

```{r, include=FALSE}
library(tidyverse)
library(patchwork)
library(ggfortify)
library(forecast)
load("./data/results_ibtracs.RData")
load("./data/results_hursat.RData")
```

# Methods

Before performing the analysis, we discuss the different methods used.

## Regression Based methods

The regression approaches use **Generalized Linear Models** and **Likelihood Ratio testing** [@mccullagh2019generalized].

### For Count Data

For the data consisting of counts (number of non-shorties per season and number of major storms per season) we perform a **Poisson regression with a log-link function**. That is, we model:

$$\log(Count) = \beta_0 + \beta_1\times Year$$

which in the original units looks like

$$Count = \exp(\beta_0 + \beta_1\times Year) = \exp(\beta_0)\times \exp(\beta_1\times Year)$$

So for the Poisson based method, all the effects we find are multiplicative.

It is worth noting that the Poisson regression is arguably more reliable than standard linear regression (particularly in the six basins) since we are working with moderately small counts. Inference is done using a Likelihood Ratio test on the $\beta$ coefficient using the `drop1()` function in R. Some residual diagnostics are provided for each fitted model.

I also report values from the predictions based on these models (confidence intervals are provided for these predictions). This provides some context of what the model is suggesting. We use the prediction at the first and last time points to find an *overall trend*, that trend is then scaled to the #/decade level for easier interpretation. 

### For Proportions

For the proportional data (proportion of major storms and proportion of intense storms), we are using a **Logistic Regression**. Let $p$=proportion of major (or intense) storms, we are modeling

$$\log\left(\frac{p}{1-p}\right) = \beta_0 + \beta_1\times Year$$

So again, when interpreting we need to transform back to the original units. It is worth noting that the associated hypothesis test for this regression must be interpreted in terms of log-odds. It is also worth noting that when performing this logistic regression we are implicitly assuming that the process controlling the total counts of storms, is not changing.

Again, a likelihood ratio test is used for significance of the $\beta_1$ coefficient. We provide predicted probability values at the first and last time point (with confidence intervals), and from those values find an *overall trend* (%/decade). Residual diagnostics are also provided.

## Changepoint Approaches

The regression approaches assume a fairly restrictive structure -- we are assuming some underlying linear model. That is, any changes we see in the record are following some underlying linear trend. In the changepoint approach, we are simply looking for shifts in the mean of the process. **Our primary results are based on the changepoint methods.**

At some level, both methods look for changes, or non-homegenieties, in the data, but the changepoint approach doesn't require that linearity assumption.

### At Most One Changepoint (AMOC)

We perform some iid **SCUSUM test** [@lund2023good] to look for at most one changepoint. The SCUSUM statistic is defined as $$\mbox{SCUSUM}=\frac{1}{n}\sum_{t=1}^n \mbox{CUSUM}(t)^2$$ for sample size $n$. The details of $\mbox{CUSUM}(t)$ for our two scenarios is outlined below.

The changepoint is found via $$k = \text{argmax} \left|\text{CUSUM}(k)\right|.$$ Under the null hypothesis of no changes, $\text{SCUSUM}$ converges to $\int_0^1 B^2(t)dt$, where $\{ B(t) \}_{t=0}^{t=1}$ is the classic Brownian bridge process of empirical statistics. We used the distribution approximations in [@Tolmatz2002] to compute $p$-values. This provides the most likely changepoint location (given only a single changepoint can occur) and its corresponding $p$-value.

It is worth noting that CUSUM based tests can be conservative in practice (reported $p$-values may be slightly elevated) -- we see this a little bit in the simulation study report. A boostrapping/simulation type approach could be performed but here we use the asymptotic approach partially because it is conservative.

### Changepoint for Frequencies

For the cyclone count series, we are assuming each observation, $X_t$, is a Poisson random variable where the mean of the series is determine by a rate parameter. Statistically speaking, when looking for changes in the series we are looking for a change in the rate, or mean, parameter $\lambda_t$. The $\text{CUSUM}(k)$ statistic can be calculated via $$\mbox{CUSUM}(k)= \frac{1}{\hat{\sigma} \sqrt{n}}\left[ \sum_{t=1}^k  X_t - \frac{k}{n} \sum_{t=1}^{n} X_t \right]$$ and the variance estimate is found via $\hat{\sigma}^2 = n^{-1} \sum_{t=1}^n X_t$ under the assumption of Poisson data [@robbins2011changepoints].

### Changepoint for Proportions

When working with the proportion of major (or intense) storms, we perform a CUSUM comparison on $\hat{p}_t$, the observed proportion of major (or intense) storms at time $t$. Let $Y_t$ be the number of major (or intense) storms in year $t$ and $X_t$ is the total number of storms in that year. Based on the Poisson-Binomial likelihood arguments we calculate $$\hat{p}_t = Y_t/X_t.$$ The $\mbox{CUSUM}(k)$ statistic is calculated on the $\hat{p}_t$ terms as $$\mbox{CUSUM}(k)= \frac{1}{\hat{\sigma} \sqrt{n}}\left[ \sum_{t=1}^k  \hat{p}_t - \frac{k}{n} \sum_{t=1}^{n} \hat{p}_t \right].$$ The standard deviation estimation in this case is unwieldy but can be found through conditional expectation arguments: $$\mbox{Var}(\hat{p}_t) = E\left[\mbox{Var}(\hat{p}_t | X_t)\right] + \mbox{Var}\left( E[\hat{p}_t|X_t]\right).$$ Detailed derivation reveals $$\mbox{Var}(\hat{p}_t) = p^2e^{-\lambda}(1-e^{-\lambda}) + p(1-p)\sum_{k=1}^\infty \frac{e^{-\lambda}\lambda^k}{k(k!)}.$$ We estimate this variance by using $$\hat{p} = \frac{\sum_{t=1}^n Y_t}{\sum_{t=1}^n X_t}~~~~~\mbox{and}~~~~~\hat{\lambda}=\bar{X} = \frac{1}{n}\sum_{t=1}^n X_t$$ in place of $p$ and $\lambda$ respectively. In practice, the term $e^{-\lambda}$ is small for large values of lambda, so the $p^2e^{-\lambda}(1-e^{-\lambda})$ terms goes to zero. The infinite summation is $E\left[\frac{1}{X}1_{X>0}\right]$ and is approximately equal to $1/\lambda$ for large values of $\lambda$. More importantly for our application, the sequence in the summation converges to zero fairly quickly, so the series can be well-approximated by replacing $\infty$ with a finite number.

### Modeling with Changepoints

Ultimately, a single changepoint model can be considered a special case of a generalized linear model. For example, suppose the method found a changepoint at time $c$ for tropical cyclone counts, we can build a model using a generalized linear regression with 
$$\log(Count) = \beta_0 + \beta_1\times I(t>c)$$
where $I(t>c)$ is an indicator function determining if the time point $t$ is greater than the changepoint location. In this parameterization, $\beta_1$ essentially determines the shift in the Poisson rate. After determining the changepoint time, we fit this model to easily extract predictions for the mean in each regime and to build rudimentary confidence intervals around those estimates. This also allows us to perform some model assessment through GLM diagnostics.

In a similar way, we do the same modeling with logistic regression for the proportion of major and intense storms. It is important to note these models are built after determining the changepoint location using the appropriate CUSUM statistic. Additional hypothesis testing should not be performed using these models. We use these fitted models for some follow-up procedures and are not the primary testing mechanism.

```{r, echo=FALSE}
##############################
## A function to extract all
##   residual plots provided
residual_plots <- function(fit) {
  if(length(hatvalues(fit)) != length(fitted.values(fit) ) ) {
    mat <- fit$model[which(fit$weights!=0),]
    names(mat) <- c("successfailure", "x")
    new_fit <- glm(successfailure ~ x, family=family(fit), data=mat)
    res_plots <- autoplot(new_fit, which=c(1,2,3,5,4,6) )
  } else {
    res_plots <- autoplot(fit, which=c(1,2,3,5,4,6))
  }
  
  df <- data.frame(y=residuals(fit, type="deviance") ) |>
    mutate(x=row_number() )
  p1 <- ggplot(df, aes(x=x,y=y)) + geom_line() + geom_point() + labs(x="Time Order", y="Deviance Residuals", title="Residuals in Time")
  p2 <- ggAcf(df$y) + labs(title="Series: Deviance Residuals")
  (res_plots[[1]] | res_plots[[2]]) /
    (res_plots[[3]] | res_plots[[4]]) /
    (res_plots[[5]] | res_plots[[6]]) /
    (p1 | p2)
}
```



# Global Cyclone Record

We begin with a study of the record across the entire globe -- the Global "basin". Our analysis looks at all non-shorties (storms that lasted at least 72 hours) and achieved tropical storm strength. We also look at those storms that achieved major hurricane strength.

{{< include ./report_analysis_code/cyclone_report_global_ibtracs.qmd >}}

{{< include ./report_analysis_code/cyclone_report_global_hursat.qmd >}}

# North Atlantic

{{< include ./report_analysis_code/cyclone_report_na_ibtracs.qmd >}}

{{< include ./report_analysis_code/cyclone_report_na_hursat.qmd >}}

# Eastern North Pacific

{{< include ./report_analysis_code/cyclone_report_ep_ibtracs.qmd >}}

{{< include ./report_analysis_code/cyclone_report_ep_hursat.qmd >}}

# Western North Pacific

{{< include ./report_analysis_code/cyclone_report_wp_ibtracs.qmd >}}

{{< include ./report_analysis_code/cyclone_report_wp_hursat.qmd >}}

# North Indian

{{< include ./report_analysis_code/cyclone_report_ni_ibtracs.qmd >}}

{{< include ./report_analysis_code/cyclone_report_ni_hursat.qmd >}}

# South Indian

{{< include ./report_analysis_code/cyclone_report_si_ibtracs.qmd >}}

{{< include ./report_analysis_code/cyclone_report_si_hursat.qmd >}}

# South Pacific

{{< include ./report_analysis_code/cyclone_report_sp_ibtracs.qmd >}}

{{< include ./report_analysis_code/cyclone_report_sp_hursat.qmd >}}

# References
