---
title: "Key Results"
subtitle: "Tables summarizing the key results"
author: "Tom Fisher"
date: today
format:
  html:
    page-layout: full
    toc: true
    toc-depth: 3
    toc-location: left
    embed-resources: true
editor: visual
bibliography: sn-bibliography.bib
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(knitr)
library(kableExtra)
load("./data/results_overview.RData")
load("data/ibtracs_storm_classifications.RData")
shorty_day_cutoff <- (min(filter(ibtracs_storms_fct, Duration_Class_Streak=="Non_Shorty")$Max_TC_Streak)-1)/4
```

## Premise and Notes

This report summarizes the big report down to just a few tables

In each table there are $7\times 6=42$ hypothesis test being performed. We perform the [False Discovery Rate](https://en.wikipedia.org/wiki/False_discovery_rate) method to adjust for multiple comparisons -- it is being reported in parenthesis.

A few caveats we have already established

- IBTrACS data goes from 1980 to 2025. HURSAT is from 1981 to 2024.
- When processing IBTrACS we are now using the USA Agency recorded wind speed.
- Storms with at least `r shorty_day_cutoff` consecutive days of tropical storm strength (non-shorty) are considered in this analysis.
- There are differences between IBTrACS and HURAT -- see the other report.

## Changepoint Approach

He we summarize the *at most one changepoint* (AMOC) method we describe in the other report. The "Shift of \## in YEAR" marks the change in mean (or proportion) and the year we found the change. It must be reiterated that by finding a change at a given year this does not mean something systematically happened, it just happens to be the point where we detecting a nonhomogeneity in the underlying mean process of the data.

### IBTrACS

```{r, echo=FALSE}
ibtracs_key_results |>
  dplyr::filter(Method=="Changepoint") |>
  mutate(adj_pvalue = p.adjust(p.value, method="fdr")) |>
  mutate(Output = paste0("Shift of ", round(Shift, 1), " (", round(Relative_Shift,2), "%) in ", Time, ". p-val=", round(p.value, 4), " (FDR: ", round(adj_pvalue, 4), ")" ) ) |>
  dplyr::select(Basin, Series, Output) |>
  pivot_wider(id_cols=Basin, names_from = Series, values_from = Output) |>
  knitr::kable(caption="Findings from the SCUSUM Changepoint method, reporting the magnitude of the shift (in terms of lambda, expected storms per year), the year of the shift, the corresponding p-value (and FDR adjusted p-value)")
```

### HURSAT

```{r, echo=FALSE}
hursat_key_results |>
  dplyr::filter(Method=="Changepoint") |>
  mutate(adj_pvalue = p.adjust(p.value, method="fdr")) |>
  mutate(Output = paste0("Shift of ", round(Shift, 1), " (", round(Relative_Shift,2), "%) in ", Time, ". p-val=", round(p.value, 4), " (FDR: ", round(adj_pvalue, 4), ")" ) ) |>
  dplyr::select(Basin, Series, Output) |>
  pivot_wider(id_cols=Basin, names_from = Series, values_from = Output) |>
  knitr::kable(caption="Findings from the SCUSUM Changepoint method, reporting the magnitude of the shift (in terms of lambda, expected storms per year), the year of the shift, the corresponding p-value (and FDR adjusted p-value)")
```

## Regression Results

Here, the coefficient is NOT a regression coefficient, but rather the decadal change based on a linear relation. Take the predicted value in Year 1, the predicted value in year $N$, find the difference and then scaled to the decade. This provide a contextual \# per decade increase (or decrease). These $p$-values are based on two-sided likelihood ratio test for the coefficient in the generalized linear model. As before, an FDR multiple comparisons is also provided.

### IBTrACS

```{r, echo=FALSE}
ibtracs_key_results |>
  dplyr::filter(Method=="Regression") |>
  mutate(adj_pvalue = p.adjust(p.value, method="fdr")) |>
  mutate(Output = paste0("Change ", round(Relative_Shift, 4)," per decade. p-val=", round(p.value, 4), " (FDR: ", round(adj_pvalue, 4), ")" ) ) |>
  dplyr::select(Basin, Series, Output) |>
  pivot_wider(id_cols=Basin, names_from = Series, values_from = Output) |>
  knitr::kable(caption="Generalized Linear Model Findings, reporting the regression coefficient (the estimated beta term), the p-value for two-sided LRT for a significant coefficient (and the FDR adjusted p-value)")
```

### HURSAT

```{r, echo=FALSE}
hursat_key_results |>
  dplyr::filter(Method=="Regression") |>
  mutate(adj_pvalue = p.adjust(p.value, method="fdr")) |>
  mutate(Output = paste0("Change ", round(Relative_Shift, 4)," per decade. p-val=", round(p.value, 4), " (FDR: ", round(adj_pvalue, 4), ")" ) ) |>
  dplyr::select(Basin, Series, Output) |>
  pivot_wider(id_cols=Basin, names_from = Series, values_from = Output) |>
  knitr::kable(caption="Generalized Linear Model Findings, reporting the regression coefficient (the estimated beta term), the p-value for two-sided LRT for a significant coefficient (and the FDR adjusted p-value)")
```
