---
title: "Shorty Sensitivity"
substitle: "A study on the definition of a shorty"
author: "Tom Fisher"
date: today
format:
  html:
    page-layout: full
    toc: true
    toc-depth: 3
    toc-location: left
    embed-resources: true
editor: visual
bibliography: sn-bibliography.bib
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(knitr)
library(kableExtra)
load("./data/shorty_sensitivity.RData")
```

## Definition of "Shorty"

This short report studies the effects of removing short duration storms from the record, or so called "shorties".

We calculate "shorties" based on the consecutive observations according to the USA agency record of 4 times per day (hours 00, 06, 12, 18). So if the cutoff for a shorty was 2-days, we need at least 9 observations at 35 knots to be counted as a non-shorty.

The HURSAT data does not provide results on regular time intervals, so we calculate streaks of wind speeds at 35+ knots, if that *streak* has a duration of at least \# days, then it is considered a non-shorty.

## Summary

### IBTrACS Record

```{r, echo=FALSE}
ibtracs_summaries |>
  dplyr::select(`Shorty Length`=Days, `Total Non-shorty`=Total, `Avg Cyclones`=`Total Storms (Avg)`, `Avg Hurricanes`=`Hurricanes (Avg)`, `Avg Major TC`=`Major Storms (Avg)`, `Avg Intense TC`=`Intense Storms (Avg)`) |>
  kable(caption="Change in Record after removing Shorties")
```

### HURSAT Record

```{r, echo=FALSE}
hursat_summaries |>
  dplyr::select(`Shorty Length`=Days, `Total Non-shorty`=Total, `Avg Cyclones`=`Total Storms (Avg)`, `Avg Hurricanes`=`Hurricanes (Avg)`, `Avg Major TC`=`Major Storms (Avg)`, `Avg Intense TC`=`Intense Storms (Avg)`) |>
  kable(caption="Change in Record after removing Shorties")
```

## Sensitivity in Changepoint Methods

### IBTrACS Record

```{r, echo=FALSE}
ibtracs_key_shorty |>
  dplyr::filter(Basin=="Global",
                Method=="Changepoint") |>
  mutate(Output = paste0("Shift of ", round(Shift, 1), " in ", Time, ". p-val=", round(p.value, 4) ) ) |>
  dplyr::select(Days, Series, Output) |>
  pivot_wider(id_cols=Days, names_from = Series, values_from = Output) |>
  rename(`Shorty Length`=Days) |>
  knitr::kable(caption="Findings from the SCUSUM Changepoint method, reporting the magnitude of the shift (in terms of lambda, expected storms per year), the year of the shift, the corresponding p-value based on Shorty definition length.")
```

### HURSAT Record

```{r, echo=FALSE}
hursat_key_shorty |>
  dplyr::filter(Basin=="Global",
                Method=="Changepoint") |>
  mutate(Output = paste0("Shift of ", round(Shift, 1), " in ", Time, ". p-val=", round(p.value, 4) ) ) |>
  dplyr::select(Days, Series, Output) |>
  pivot_wider(id_cols=Days, names_from = Series, values_from = Output) |>
  rename(`Shorty Length`=Days) |>
  knitr::kable(caption="Findings from the SCUSUM Changepoint method, reporting the magnitude of the shift (in terms of lambda, expected storms per year), the year of the shift, the corresponding p-value based on Shorty definition length.")
```

## Regression Results

### IBTrACS

```{r, echo=FALSE}
ibtracs_key_shorty |>
  dplyr::filter(Basin=="Global",
                Method=="Regression") |>
  mutate(Output = paste0("Change ", round(Relative_Shift, 4)," per decade. p-val=", round(p.value, 4) ) ) |>
  dplyr::select(Days, Series, Output) |>
  pivot_wider(id_cols=Days, names_from = Series, values_from = Output) |>
  rename(`Shorty Length`=Days) |>
  knitr::kable(caption="Generalized Linear Model Findings, reporting the regression coefficient (the estimated beta term), the p-value for two-sided LRT for a significant coefficient based on Shorty definition length.")
```

### HURSAT

```{r, echo=FALSE}
hursat_key_shorty |>
  dplyr::filter(Basin=="Global",
                Method=="Regression") |>
  mutate(Output = paste0("Change ", round(Relative_Shift, 4)," per decade. p-val=", round(p.value, 4) ) ) |>
  dplyr::select(Days, Series, Output) |>
  pivot_wider(id_cols=Days, names_from = Series, values_from = Output) |>
  rename(`Shorty Length`=Days) |>
  knitr::kable(caption="Generalized Linear Model Findings, reporting the regression coefficient (the estimated beta term), the p-value for two-sided LRT for a significant coefficient based on Shorty definition length.")
```
