---
title: "Tropical Cyclone Analysis"
subtitle: "Single changepoint & Regression Methods"
author: "Tom Fisher"
date: today
format:
  html:
    page-layout: full
    toc: true
    toc-depth: 3
    toc-location: left
    embed-resources: true
editor: visual
bibliography: sn-bibliography.bib
---

```{r, include=FALSE}
library(tidyverse)
set.seed(1)

load("./data/ibtracs_nonShortiesForProportions.RData")
ibtracs_data <- non_shorty_for_props
load("./data/hursat_nonShortiesForProportions.RData")
hursat_data <- non_shorty_for_props

# Load cusum test
source("cusumBasedChangePointsTest.R")

## Add the "Global" basin to each dataset
ibtracs_all_basins <- ibtracs_data |>
  group_by(SEASON) |>
  summarize(Major_Storms = sum(Major_Storms),
            Intense_Storms = sum(Intense_Storms),
            Total_Storms = sum(Total_Storms)) |>
  mutate(BASIN = "Global") |>
  bind_rows(ibtracs_data)

## Now with the HURSAT data
hursat_all_basins <- hursat_data |>
  group_by(SEASON) |>
  summarize(Major_Storms = sum(Major_Storms),
            Intense_Storms = sum(Intense_Storms),
            Total_Storms = sum(Total_Storms)) |>
  mutate(BASIN = "Global") |>
  bind_rows(hursat_data)

num_years_ibtracs <- length(unique(ibtracs_data$SEASON))
num_years_hursat <- length(unique(hursat_data$SEASON) )

## Trimmed version of IBTraCs matching HURSAT years
ibtracs_trimmed_all_basins <- ibtracs_all_basins |>
  dplyr::filter(SEASON %in% unique(hursat_all_basins$SEASON) )
```

```{r, echo=FALSE}
## Main analysis function
##
## Specify a Basin and dataset (IBTracs vs HURSAT vs Trimmed IBTracs)
## All analysis we need is reported.
get_report_basin <- function(basin="Global", all_basins=ibtracs_all_basins) {
  ## Some Data Processing for plots and summary statistics
  ##   and used in our analysis
  ## Note: YearsSinceStart is just a number counting the number of years
  ##   since the beginning of our record (1980 or 1981, depending on source)
  storm_basin <- all_basins |>
    dplyr::filter(BASIN==basin) |>
    mutate(YearsSinceStart = SEASON - min(SEASON)) |>
    mutate(Prop_Major = Major_Storms/Total_Storms,
           Prop_Intense = Intense_Storms/Total_Storms)
  
  ## Put data in a tall (or long) format for plotting
  storm_basin_long <- storm_basin |>
    dplyr::select(SEASON, Total_Storms, Major_Storms, Prop_Major, Intense_Storms, Prop_Intense, BASIN) |>
    pivot_longer(c(Total_Storms, Major_Storms, Prop_Major, Intense_Storms, Prop_Intense), names_to="Measure", values_to="values") |>
    mutate(Measure = factor(Measure, levels=c("Total_Storms", "Major_Storms", "Prop_Major", "Intense_Storms", "Prop_Intense"),
                            labels=c("Total Non-Shorties",
                                     "Major Storms",
                                     "Proportion of Major Storms",
                                     "Intense Storms",
                                     "Proportion of Intense Storms")))
  
  ## Basic Plots of the data
  plot_data <- ggplot(storm_basin_long, aes(x=SEASON, y=values) ) +
    geom_line(color="gray50") + 
    geom_point(color="gray40") +
    facet_wrap(~Measure, nrow=5, scales="free_y") +
    theme_bw() + 
    theme(axis.title = element_blank() )
  
  ## Summary statistics (Full data aggregate)
  summary_stat <- storm_basin |>
    summarise(`Total Storms (Avg)` = mean(Total_Storms),
              `Total Storms (stddev)` = sd(Total_Storms),
              `Major Storms (Avg)` = mean(Major_Storms),
              `Major Storms (stddev)` = sd(Major_Storms),
              `Prop Major Storms` = sum(Major_Storms)/sum(Total_Storms),
              `Intense Storms (Avg)` = mean(Intense_Storms),
              `Intense Storms (stddev)` = sd(Intense_Storms),
              `Prop Intense Storms` = sum(Intense_Storms)/sum(Total_Storms) )
  
  ## Poisson Regression on Total, Major & Intense
  
  count_pois_fit <- glm(Total_Storms ~ YearsSinceStart, data=storm_basin,
                        family=poisson(link="log"))
  
  major_pois_fit <- glm(Major_Storms ~ YearsSinceStart, data=storm_basin,
                        family=poisson(link="log") )

  intense_pois_fit <- glm(Intense_Storms ~ YearsSinceStart, data=storm_basin,
                          family=poisson(link="log") )

  ## Poisson SCUSUM AMOC Changepoint Test
  ##  on totals, major, and intense

  pois_chpt_test <- scusumTestPoisson(storm_basin$Total_Storms)
  
  major_pois_chpt_test <- scusumTestPoisson(storm_basin$Major_Storms)
  
  intense_pois_chpt_test <- scusumTestPoisson(storm_basin$Intense_Storms)
  
  ## Logistic Regression on Proportion of Major
  ##    and Proportion of Intense

  prop_data <- storm_basin |>
    mutate(Non_Major = Total_Storms - Major_Storms,
           Non_Intense = Total_Storms - Intense_Storms)
  
  prop_major_logit_fit <- glm(cbind(Major_Storms, Non_Major) ~ YearsSinceStart, 
                              data=prop_data, 
                              family=binomial)
  
  prop_intense_logit_fit <- glm(cbind(Intense_Storms, Non_Intense) ~ YearsSinceStart, 
                                data=prop_data, 
                                family=binomial)

  ## SCUSUM Test for Proportions
  ##    On Major & Intense
  
  prop_major_chpt_test <- scusumTestProportion(cbind(prop_data$Major_Storms, prop_data$Total_Storms) )
  
  prop_intense_chpt_test <- scusumTestProportion(cbind(prop_data$Intense_Storms, prop_data$Total_Storms) )
  
  ## Output all results in a useful list
  list(summary_stat = summary_stat,
       
       pois_fit = count_pois_fit,
       major_fit = major_pois_fit,
       intense_fit = intense_pois_fit,
       
       major_logit_fit = prop_major_logit_fit,
       intense_logit_fit = prop_intense_logit_fit,
       
       pois_chpt_test = pois_chpt_test,
       major_pois_chpt_test = major_pois_chpt_test,
       intense_pois_chpt_test = intense_pois_chpt_test,
       
       prop_major_chpt_test = prop_major_chpt_test,
       prop_intense_chpt_test = prop_intense_chpt_test,
       
       plot_data = plot_data,
       
       data=storm_basin)
}
```

# Methods

Before performing the analysis, we discuss the different methods used.

## Regression Based methods

The regression approaches use **Generalized Linear Models** and **Likelihood Ratio testing** [@mccullagh2019generalized]

### For Count Data

For the data consisting of counts (number of non-shorties per season and number of major storms per season) we perform a **Poisson regression with a log-link function**. That is, we model:

$$\log(Count) = \beta_0 + \beta_1\times Year$$

which in the original units looks like

$$Count = \exp(\beta_0 + \beta_1\times Year) = \exp(\beta_0)\times \exp(\beta_1\times Year)$$

So for the Poisson based method, all the effects we find are multiplicative.

It is worth noting that the Poisson regression is arguably more reliable than standard linear regression (particularly in the six basins) since we are working with moderately small counts. Inference is done using a Likelihood Ratio test on the $\beta$ coefficient using the `drop1()` function in R.

I also report values from the predictions based on these models. This provides some context of what the model is suggesting. We use the prediction at the first and last time points to find an *overall trend*, that trend is then scaled to the #/decade level.

### For Proportions

For the proportional data (proportion of major storms and proportion of intense storms), we are using a **Logistic Regression**. Let $p$=proportion of major (or intense) storms, we are modeling

$$\log\left(\frac{p}{1-p}\right) = \beta_0 + \beta_1\times Year$$

So again, when interpreting we need to transform back to the original units. It is worth noting that the associated hypothesis test for this regression must be interpreted in terms of log-odds. It is also worth noting that when performing this logistic regression we are implicitly assuming that the process controlling the total counts of storms, is not changing.

Again, a likelihood ratio test is used for significance of the $\beta_1$ coefficient. We provide predicted probability values at the first and last time point, and from those values find an *overall trend* (%/decade).

## Changepoint Approaches

The regression approaches assume a fairly restrictive structure -- we are assuming some underlying linear model. That is, any changes we see in the record are following some underlying linear trend. In the changepoint approach, we are simply looking for shifts in the mean of the process.

At some level, both methods looking for changes, or non-homegenieties in the data, but the changepoint approach doesn't require that linearity assumption.

### At Most One Changepoint (AMOC)

We perform some iid **SCUSUM test** [@lund2023good] to look for at most one changepoint. The SCUSUM statistic is defined as $$\mbox{SCUSUM}=\frac{1}{n}\sum_{t=1}^n \mbox{CUSUM}(t)^2$$ for sample size $n$, and where $$\mbox{CUSUM}(k) = \frac{k}{n}\left(1-\frac{k}{n}\right)\left[\frac{\sqrt{n}}{\hat{\sigma}_{\hat{\theta}}}\left(\hat{\theta}_k - \hat{\theta}_k^*\right)\right].$$ The changepoint is found via $$k = \text{argmax} \left|\text{CUSUM}(k)\right|$$ Under the null hypothesis of no changes, $\text{SCUSUM}$ converges to $\int_0^1 B^2(t)dt$, where $\{ B(t) \}_{t=0}^{t=1}$ is the classical Brownian bridge process of empirical statistics. We used the distribution approximations in [@Tolmatz2002] to compute $p$-values. This provides the most likely changepoint location (given only a single changepoint can occur) and its corresponding $p$-value.

It is worth noting that CUSUM based tests can be conservative in practice (reported $p$-values may be slightly elevated) -- we see this a little bit in the simulation study report. If you want something more robust, a boostrapping/simulation type approach could be performed. For this report we use the asymptotic approach partially because it is conservative.

### Changepoint for Frequencies

When working with Poisson data (count data), we are looking for a change in the rate, or mean, parameter $\lambda_t$. The $\text{CUSUM}(k)$ statistic can be calculated via $$\mbox{CUSUM}(k)= \frac{1}{\hat{\sigma} \sqrt{n}}
\left[ \sum_{t=1}^k  X_t - \frac{k}{n} \sum_{t=1}^{n} X_t \right]$$ and the variance estimate is found via $\hat{\sigma}^2 = n^{-1} \sum_{t=1}^n X_t$ under the assumption of Poisson data [@robbins2011changepoints]

### Changepoint for Proportions

When working with the proportion of major (or intense) storms, we assume an underlying Binomial likelihood and the parameter of interested in $p$, the proportion of major (or intense) storms. Let $X_t$ be the number of major (or intense) storms in year $t$ and $T_t$ is the total number of storms in that year. Based on Binimial likelihood arguments we calculate $$\hat{\theta}_k = \hat{p}_k = \frac{\sum_{i=1}^k X_i}{\sum_{i=1}^k T_i}~~~~~\mbox{and}~~~~~\hat{\theta}_k^* = \hat{p}_k^* = \frac{\sum_{i=k+1}^n X_i}{\sum_{i=k+1}^n T_i}.$$ For the variance estimate, $\sigma_{\hat{\theta}}$, we use the Wald estimate (covered in Intro Stat courses), $$\sigma_{\hat{p}} = \sqrt{\frac{\hat{p}_n(1-\hat{p}_n)}{\sum_{i=1}^n T_i}}$$

# Global Cyclone Record

We begin with a study of the record across the entire globe -- the Global "basin". Our analysis looks at all non-shorties (storms that lasted at least 48 hours) and achieved tropical storm strength. We also look at those storms that achieved major hurricane strength.

{{< include ./report_analysis_code/cyclone_report_global_ibtracs.qmd >}} 
{{< include ./report_analysis_code/cyclone_report_global_hursat.qmd >}} 
{{< include ./report_analysis_code/cyclone_report_global_trimmed.qmd >}}

# North Atlantic

{{< include ./report_analysis_code/cyclone_report_north_atlantic_ibtracs.qmd >}} 
{{< include ./report_analysis_code/cyclone_report_north_atlantic_hursat.qmd >}} 
{{< include ./report_analysis_code/cyclone_report_north_atlantic_trimmed.qmd >}}

# Eastern North Pacific

{{< include ./report_analysis_code/cyclone_report_eastern_pacific_ibtracs.qmd >}} 
{{< include ./report_analysis_code/cyclone_report_eastern_pacific_hursat.qmd >}} 
{{< include ./report_analysis_code/cyclone_report_eastern_pacific_trimmed.qmd >}}

# Western North Pacific

{{< include ./report_analysis_code/cyclone_report_western_pacific_ibtracs.qmd >}} 
{{< include ./report_analysis_code/cyclone_report_western_pacific_hursat.qmd >}} 
{{< include ./report_analysis_code/cyclone_report_western_pacific_trimmed.qmd >}}

# North Indian

{{< include ./report_analysis_code/cyclone_report_north_indian_ibtracs.qmd >}} 
{{< include ./report_analysis_code/cyclone_report_north_indian_hursat.qmd >}} 
{{< include ./report_analysis_code/cyclone_report_north_indian_trimmed.qmd >}}

# South Indian

{{< include ./report_analysis_code/cyclone_report_southern_indian_ibtracs.qmd >}} 
{{< include ./report_analysis_code/cyclone_report_southern_indian_hursat.qmd >}} 
{{< include ./report_analysis_code/cyclone_report_southern_indian_trimmed.qmd >}}

# South Pacific

{{< include ./report_analysis_code/cyclone_report_southern_pacific_ibtracs.qmd >}} 
{{< include ./report_analysis_code/cyclone_report_southern_pacific_hursat.qmd >}} 
{{< include ./report_analysis_code/cyclone_report_southern_pacific_trimmed.qmd >}}



```{r, echo=FALSE}
save(ibtracs_global_results, ibtracs_na_results, ibtracs_ep_results,
     ibtracs_wp_results, ibtracs_ni_results, ibtracs_si_results, ibtracs_sp_results,
     file="./data/results_ibtracs.RData")
save(hursat_global_results, hursat_na_results, hursat_ep_results,
     hursat_wp_results, hursat_ni_results, hursat_si_results, hursat_sp_results,
     file="./data/results_hursat.RData")
save(ibtracs_trimmed_global_results, ibtracs_trimmed_na_results, ibtracs_trimmed_ep_results,
     ibtracs_trimmed_wp_results, ibtracs_trimmed_ni_results, ibtracs_trimmed_si_results, ibtracs_trimmed_sp_results,
     file="./data/results_trimmed.RData")
```

# References
