---
title: "Tropical Cyclone Analysis"
subtitle: "Single Change Point & Regression Methods"
author: "Tom Fisher"
date: today
format:
  html:
    toc: true
    toc-depth: 3
    toc-location: left
    embed-resources: true
editor: visual
---

```{r, include=FALSE}
library(tidyverse)
set.seed(1)

load("./data/ibtracs_nonShortiesForProportions.RData")
ibtracs_data <- non_shorty_for_props
load("./data/hursat_nonShortiesForProportions.RData")
hursat_data <- non_shorty_for_props

# Load cusum test
source("cusumBasedChangePointsTest.R")

## Add the "Global" basin to each dataset
ibtracs_all_basins <- ibtracs_data |>
  group_by(SEASON) |>
  summarize(Major_Storms = sum(Major_Storms),
            Intense_Storms = sum(Intense_Storms),
            Total_Storms = sum(Total_Storms)) |>
  mutate(BASIN = "Global") |>
  bind_rows(ibtracs_data)

## Now with the HURSAT data
hursat_all_basins <- hursat_data |>
  group_by(SEASON) |>
  summarize(Major_Storms = sum(Major_Storms),
            Intense_Storms = sum(Intense_Storms),
            Total_Storms = sum(Total_Storms)) |>
  mutate(BASIN = "Global") |>
  bind_rows(hursat_data)

num_years_ibtracs <- length(unique(ibtracs_data$SEASON))
num_years_hursat <- length(unique(hursat_data$SEASON) )

## Trimmed version of IBTraCs matching HURSAT years
ibtracs_trimmed_all_basins <- ibtracs_all_basins |>
  dplyr::filter(SEASON %in% unique(hursat_all_basins$SEASON) )
```

```{r, echo=FALSE}
## Main analysis function
##
## Specify a Basin and dataset (IBTracs vs HURSAT vs Trimmed IBTracs)
## All analysis we need is reported.
get_report_basin <- function(basin="Global", all_basins=ibtracs_all_basins) {
  ## Some Data Processing for plots and summary statistics
  ##   and used in our analysis
  ## Note: YearsSinceStart is just a number counting the number of years
  ##   since the beginning of our record (1980 or 1981, depending on source)
  storm_basin <- all_basins |>
    dplyr::filter(BASIN==basin) |>
    mutate(YearsSinceStart = SEASON - min(SEASON)) |>
    mutate(Prop_Major = Major_Storms/Total_Storms,
           Prop_Intense = Intense_Storms/Total_Storms)
  
  ## Put data in a tall (or long) format for plotting
  storm_basin_long <- storm_basin |>
    dplyr::select(SEASON, Total_Storms, Major_Storms, Prop_Major, Intense_Storms, Prop_Intense, BASIN) |>
    pivot_longer(c(Total_Storms, Major_Storms, Prop_Major, Intense_Storms, Prop_Intense), names_to="Measure", values_to="values") |>
    mutate(Measure = factor(Measure, levels=c("Total_Storms", "Major_Storms", "Prop_Major", "Intense_Storms", "Prop_Intense"),
                            labels=c("Total Non-Shorties",
                                     "Major Storms",
                                     "Proportion of Major Storms",
                                     "Intense Storms",
                                     "Proportion of Intense Storms")))
  
  ## Basic Plots of the data
  plot_data <- ggplot(storm_basin_long, aes(x=SEASON, y=values) ) +
    geom_line(color="gray50") + 
    geom_point(color="gray40") +
    facet_wrap(~Measure, nrow=5, scales="free_y") +
    theme_bw() + 
    theme(axis.title = element_blank() )
  
  ## Summary statistics (Full data aggregate)
  summary_stat <- storm_basin |>
    summarise(`Total Storms (Avg)` = mean(Total_Storms),
              `Total Storms (stddev)` = sd(Total_Storms),
              `Major Storms (Avg)` = mean(Major_Storms),
              `Major Storms (stddev)` = sd(Major_Storms),
              `Prop Major Storms` = sum(Major_Storms)/sum(Total_Storms),
              `Intense Storms (Avg)` = mean(Intense_Storms),
              `Intense Storms (stddev)` = sd(Intense_Storms),
              `Prop Intense Storms` = sum(Intense_Storms)/sum(Total_Storms) )
  
  ## Poisson Regression on Total, Major & Intense
  
  count_pois_fit <- glm(Total_Storms ~ YearsSinceStart, data=storm_basin,
                        family=poisson(link="log"))
  
  major_pois_fit <- glm(Major_Storms ~ YearsSinceStart, data=storm_basin,
                        family=poisson(link="log") )

  intense_pois_fit <- glm(Intense_Storms ~ YearsSinceStart, data=storm_basin,
                          family=poisson(link="log") )

  ## Poisson SCUSUM AMOC Change Point Test
  ##  on totals, major, and intense

  pois_chpt_test <- scusumTestPoisson(storm_basin$Total_Storms)
  
  major_pois_chpt_test <- scusumTestPoisson(storm_basin$Major_Storms)
  
  intense_pois_chpt_test <- scusumTestPoisson(storm_basin$Intense_Storms)
  
  ## Logistic Regression on Proportion of Major
  ##    and Proportion of Intense

  prop_data <- storm_basin |>
    mutate(Non_Major = Total_Storms - Major_Storms,
           Non_Intense = Total_Storms - Intense_Storms)
  
  prop_major_logit_fit <- glm(cbind(Major_Storms, Non_Major) ~ YearsSinceStart, 
                              data=prop_data, 
                              family=binomial)
  
  prop_intense_logit_fit <- glm(cbind(Intense_Storms, Non_Intense) ~ YearsSinceStart, 
                                data=prop_data, 
                                family=binomial)

  ## SCUSUM Test for Proportions
  ##    On Major & Intense
  
  prop_major_chpt_test <- scusumTestProportion(cbind(prop_data$Major_Storms, prop_data$Total_Storms) )
  
  prop_intense_chpt_test <- scusumTestProportion(cbind(prop_data$Intense_Storms, prop_data$Total_Storms) )
  
  ## Output all results in a useful list
  list(summary_stat = summary_stat,
       
       pois_fit = count_pois_fit,
       major_fit = major_pois_fit,
       intense_fit = intense_pois_fit,
       
       major_logit_fit = prop_major_logit_fit,
       intense_logit_fit = prop_intense_logit_fit,
       
       pois_chpt_test = pois_chpt_test,
       major_pois_chpt_test = major_pois_chpt_test,
       intense_pois_chpt_test = intense_pois_chpt_test,
       
       prop_major_chpt_test = prop_major_chpt_test,
       prop_intense_chpt_test = prop_intense_chpt_test,
       
       plot_data = plot_data,
       
       data=storm_basin)
}
```

# Methods

Before performing the analysis, we discuss the different methods used.

## Regression Based methods

### For Count Data

For the data consisting of counts (number of non-shorties per season and number of major storms per season) we perform a Poisson regression with a log-link function. That is, we model:

$$\log(Count) = \beta_0 + \beta_1\times Year$$

which in the original units looks like

$$Count = \exp(\beta_0 + \beta_1\times Year) = \exp(\beta_0)\times \exp(\beta_1\times Year)$$

So for the Poisson based method, all the effects we find are multiplicative.

It is worth noting that the Poisson regression is arguably more reliable than standard linear regression (particularly in the six basins) since we are working with moderately small counts.

I also report values from the predictions based on these models. This provides some context of what the model is suggesting.

### For Proportions

For the proportional data (proportion of major storms), we are using a logistic regression. Let $p$=proportion of major storms, we are modeling

$$\log\left(\frac{p}{1-p}\right) = \beta_0 + \beta_1\times Year$$

So again, when interpreting we need to transform back to the original units. But it is worth noting that the associated hypothesis test for this regression must be interpreted in terms of log-odds. It is also worth noting that when performing this logistic regression we are implicitly assuming that the process controlling the total counts of storms, is not changing. We use this same method on the proportion of intense storms.

## Change Point Approaches

The regression approaches assume a fairly restrictive structure -- we are assuming some linear model. That is, any changes we see in the record are following some underlying linear trend. In the change point approach, we are simply looking for shifts in the mean of the process.

At some level, both methods looking for changes, or non-homegenieties in the data, but the change point approach doesn't require that linearity assumption.

### At Most One Change Point (AMOC)

We perform some iid SCUSUM based-test to look for at most one change point. This provides the most likely change point location (given only a single change point can occur) and its corresponding $p$-value.

It is worth noting that CUSUM based tests can be conservative in practice (reported $p$-values may be slightly elevated). If we want something more robust, we can perform a boostrapping/simulation type approach. For this report I use the asymptotic approach.

### Change point for Proportions

We coded up a SCUSUM test for the proportion using standard Intro Stat binomial distribution theory. This is not all that different than the MDL search performed earlier. However, unlike the counts (or total number of storms), now the parameter of interest is $p$, the proportion of major storms or proportion of intense storms.

# Global Cyclone Record

We begin with a study of the record across the entire globe -- the Global "basin". Our analysis looks at all non-shorties (storms that lasted at least 48 hours) and achieved tropical storm strength. We also look at those storms that achieved major hurricane strength.

{{< include ./report_analysis_code/cyclone_report_global_ibtracs.qmd >}}
{{< include ./report_analysis_code/cyclone_report_global_hursat.qmd >}}
{{< include ./report_analysis_code/cyclone_report_global_trimmed.qmd >}}

# North Atlantic

{{< include ./report_analysis_code/cyclone_report_north_atlantic_ibtracs.qmd >}}
{{< include ./report_analysis_code/cyclone_report_north_atlantic_hursat.qmd >}}
{{< include ./report_analysis_code/cyclone_report_north_atlantic_trimmed.qmd >}}

# Eastern North Pacific

{{< include ./report_analysis_code/cyclone_report_eastern_pacific_ibtracs.qmd >}}
{{< include ./report_analysis_code/cyclone_report_eastern_pacific_hursat.qmd >}}
{{< include ./report_analysis_code/cyclone_report_eastern_pacific_trimmed.qmd >}}

# Western North Pacific

{{< include ./report_analysis_code/cyclone_report_western_pacific_ibtracs.qmd >}}
{{< include ./report_analysis_code/cyclone_report_western_pacific_hursat.qmd >}}
{{< include ./report_analysis_code/cyclone_report_western_pacific_trimmed.qmd >}}

# North Indian

{{< include ./report_analysis_code/cyclone_report_north_indian_ibtracs.qmd >}}
{{< include ./report_analysis_code/cyclone_report_north_indian_hursat.qmd >}}
{{< include ./report_analysis_code/cyclone_report_north_indian_trimmed.qmd >}}

# South Indian

{{< include ./report_analysis_code/cyclone_report_southern_indian_ibtracs.qmd >}}
{{< include ./report_analysis_code/cyclone_report_southern_indian_hursat.qmd >}}
{{< include ./report_analysis_code/cyclone_report_southern_indian_trimmed.qmd >}}

# South Pacific

{{< include ./report_analysis_code/cyclone_report_southern_pacific_ibtracs.qmd >}}
{{< include ./report_analysis_code/cyclone_report_southern_pacific_hursat.qmd >}}
{{< include ./report_analysis_code/cyclone_report_southern_pacific_trimmed.qmd >}}


```{r, echo=FALSE}
# save(ibtracs_global_results, ibtracs_na_results, ibtracs_ep_results, 
#      ibtracs_wp_results, ibtracs_ni_results, ibtracs_si_results, ibtracs_sp_results,
#      file="./data/results_ibtracs.RData")
# save(hursat_global_results, hursat_na_results, hursat_ep_results, 
#      hursat_wp_results, hursat_ni_results, hursat_si_results, hursat_sp_results,
#      file="./data/results_hursat.RData")
# save(trimmed_global_results, trimmed_na_results, trimmed_ep_results, 
#      trimmed_wp_results, trimmed_ni_results, trimmed_si_results, trimmed_sp_results,
#      file="./data/results_trimmed.RData")
```
